\section*{Introduction}

A macroscopic ion-channel current is the summed signal of a population of channels, and its mean carries the gating kinetics only through the deterministic relaxation, while its fluctuations carry, in addition, the number of channels, the unitary conductance, and, through the temporal structure of the noise, the rates themselves. The standard way to fit a kinetic scheme to such a recording is least squares on the mean current in a deterministic framework: the four approaches surveyed by \cite{clerx2019four} are all deterministic fits with a root-mean-square objective, and the community benchmark IonBench \cite{owen2025ionbench} states that it does not cover optimisation approaches for stochastic models. The cost of fitting only the mean is one the field has named, that standard methods for deterministic models do not distinguish stochastic channel gating from measurement noise and return biased estimates \cite{delcore2025parameter}. This is not an argument that every recording should be fit by a likelihood, only that the fluctuations carry information the mean-fit discards. One established method does use the fluctuations, non-stationary fluctuation analysis, but it is a two-moment regression that returns the unitary current, the channel number and the peak open probability rather than a kinetic scheme, and its own practitioners note that the unitary current is close to the only parameter it recovers reliably and that kinetic rates have essentially never been estimated this way for receptors in their native setting \cite{stepanyuk2014maximum}.

Using the fluctuations to identify a scheme requires a likelihood for the recorded current, and the likelihoods that exist for macroscopic recordings are a small lineage, each an approximation: the full-covariance method of \cite{celentano2004use}, exact but costing tens of hours per fit; the diagonal likelihood of \cite{milescu2005maximum}; the recursive filter of \cite{moffatt2007estimation}; the semiseparable-covariance methods of \cite{stepanyuk2011efficient}; and the Bayesian Kalman filter of \cite{munch2022bayesian}. Every one of them makes the same two Gaussian approximations and differs along two axes. The first is recursion, whether the hidden occupancy distribution is conditioned on the data as they arrive, as in a filter, or propagated open-loop from the initial condition. The second is the acquisition window, whether each sample is treated as an instantaneous observation or as what it physically is, an average of the current over a finite window. On the first axis the recursive filters are established and in use; on the second, the published macroscopic-current likelihoods treat each sample as instantaneous, and a filter that conditions the interval-averaged observable on the window was introduced only recently (MacroIR). This paper holds the recursion where the prior work sits, on, and varies the window conditioning through the ladder R, MR, VR, IR, so that the family reads as a coordinate system rather than a set of competitors.

The field has written the open questions down in its own words. On recursion, \cite{milescu2005maximum} left open whether estimates obtained by any method that is not a Bayesian filtering algorithm are intrinsically biased, and named the dominant source of error in his own method as the local time correlation of the current, which is exactly the cross-interval term of the distortion decomposition this paper measures. On averaging, \cite{moffatt2007estimation} noted that electrophysiological measurements always result from time-averaging while the recursive algorithms treat them as instantaneous, and that a formulation able to handle time-averaged signals is therefore fundamental to their use.

These questions stayed open for a subtler reason than hard algebra. An approximate likelihood is a misspecified model, and there was no accepted way to ask, of a given approximation, how badly it misrepresents the process. Goodness of fit does not answer it, because an approximation can fit the mean current closely and still report an information matrix wrong by more than an order of magnitude; comparing two approximations to each other does not answer it, because both may be wrong. What makes the question answerable is an asymmetry: the forward process can be simulated exactly, while its likelihood cannot be computed exactly, so the simulation supplies the ground truth that the likelihood is an approximation of, and any candidate likelihood can be interrogated against data drawn from the very process it claims to describe. The statistical apparatus this unlocks is classical: the score, the gradient of the log-likelihood, must average to zero at the true parameters if the likelihood is correct, and the covariance of the score must equal the Fisher information the likelihood reports, the information matrix equality \cite{white1982maximum,huber1967behavior}. Under misspecification both fail, and how they fail is measurable. The apparatus is standard in statistics; what is new is applying it to macroscopic-current likelihoods, which the exact simulator is what makes possible.

This paper does three things, in the order it delivers them. It provides a test, at the likelihood level: the standardized residuals white with unit variance, the score vanishing at the true parameters, and the covariance of the score matching the Fisher information the likelihood reports. It turns the mismatch between the last two into a measurement, an information distortion matrix that says, in units of parameter variance the reader already understands, by how much and in which directions an approximation misreports what the data determine. And it draws a map, across channel number, acquisition interval and instrumental noise, for the recursive ladder, in a minimal two-state channel model, in simulation, in the regime where gating fluctuations dominate the instrumental noise. The minimal model is a choice: it isolates the error the approximation itself commits, with no confounding from scheme misspecification, and it keeps the exact simulation cheap enough to run thousands of replicates. The few-channel boundary, richer schemes, the stationary regime and experimental data are the subjects of companion work.

What the map buys is a decision the field is actively debating. \cite{delcore2025parameter} argue against filtering on cost, because the moment integration and the correction step run between every pair of samples, and they are right about the cost. A distortion map is what that decision needs: it says which approximation misreports the uncertainty, by how much, and in which corner of the operating regime, so a reader can see where the cheaper method is safe and where it is not. The same matrix is, near the maximum, the correction that keeps Bayesian model comparison valid under an approximate likelihood, which is where this work goes next; that derivation is deferred to a later component.
