\section*{Diagnostics: testing a likelihood against its simulation}

An approximate likelihood is a misspecified model, and misspecification leaves a classical signature. At the true parameters the score no longer has zero mean, so the estimator the likelihood defines is biased; and the covariance of the score no longer equals the negative expected curvature, so the uncertainty the likelihood reports is wrong. These are the two Bartlett identities, and their failure is the content of the information-matrix-equality diagnostic of \cite{white1982maximum} and of the sandwich covariance of \cite{huber1967behavior,godambe1960optimum}. In the classical setting those identities can only be tested, through a statistic with a known finite-sample defect, because the data-generating law is unknown. Our setting removes that limitation. The forward process is an exact continuous-time Markov chain (CTMC) and we simulate it exactly, so we hold unlimited data drawn from the very law the likelihood claims to describe, at parameters we set. The classical test therefore becomes a direct measurement: the anchor curvature is computed analytically and the score covariance is estimated by Monte Carlo over independent simulated recordings, with no null hypothesis to reject. The apparatus is standard; what is new here is applying it in this domain and reading its output as a measured property of each approximation rather than as a hypothesis to accept or reject.

We read three quantities off the simulation, in increasing strength. The predictive residual and the information anchor are evaluated at the estimate; the expectation behind the score bias is taken at the simulation truth. For a misspecified likelihood the optimum and the truth differ, and that optimum-versus-truth gap is what makes the bias observable.

The first is the standardized residual. With $\mu_t$ and $\sigma_t^2$ the likelihood's own predictive mean and variance for interval $t$, the residual $r_t=(y_t-\mu_t)/\sigma_t$ must have mean zero, unit variance, and no autocorrelation across intervals if the one-step predictive distribution is correct (Figure 3). This probes the predictive one interval at a time. It is the cheapest of the three and the only one an experimentalist can run on a real recording, since it needs the predictive moments but not the true parameters. Residual whiteness in particular is not optional: residual autocorrelation is the same temporal signal that the strongest check quantifies below.

The second is the score. Writing $\ell(\theta)$ for the total log-likelihood and $S(\theta)=\nabla_\theta\ell(\theta)$ for its gradient, correct specification requires $\mathbb{E}[S(\theta^\ast)]=0$ at the true parameters $\theta^\ast$. A nonzero mean score is what biases the estimate, and projected through the parameter covariance it becomes a displacement in parameter units, $b=\mathbf{H}^{-1}\,\mathbb{E}[S]$, the reader-facing statement of how far the estimate is pushed (Figure 3). This projection is a first-order approximation, from the expansion of the score at the optimum given below. Unlike the residual, the check needs a known truth, so it requires the simulator, and its confidence interval narrows as the number of simulated recordings grows.

The third and strongest is the information itself. Correct specification requires $\mathrm{Cov}[S]=\mathbf{H}$, the information the likelihood reports. An approximation can pass the residual and score checks and still violate this one by an order of magnitude. The two failures are independent: whether the estimate is biased is governed by how the conductance is modelled over the interval, while whether the reported information is inflated is governed by the recursion. The non-recursive members of the wider family (Theory) show the second failure in its gross form, which is why this paper holds the recursion fixed and turns the three checks on the recursive ladder, where the distortions the interval conditioning produces are subtle enough that a diagnostic has to be sharp to catch them. Because the failure is directional, we record it as a matrix,
\[ \mathbf{C}=\mathbf{H}^{-1/2}\,\mathbf{J}\,\mathbf{H}^{-1/2}, \]
with $\mathbf{J}=\mathrm{Cov}(S)$ the covariance of the total score, cross-interval terms included. This definition is exact algebra given $\mathbf{H}$ and $\mathbf{J}$, and under correct specification $\mathbf{J}=\mathbf{H}$ so $\mathbf{C}=\mathbf{I}$ exactly. We fix the direction by convention: $C_{ii}>1$ means the likelihood under-reports the uncertainty in parameter $i$ and is over-confident, $C_{ii}<1$ means it over-reports. The eigen-directions of $\mathbf{C}$ say which combinations of parameters are misreported, which a scalar summary cannot recover.

The anchor $\mathbf{H}$ is the likelihood's own Gaussian Fisher information, $\mathbf{H}=\sum_t\mathbb{E}\!\left[(\nabla\mu_t)(\nabla\mu_t)^\top/\sigma_t^2+(\nabla\sigma_t^2)(\nabla\sigma_t^2)^\top/(2\sigma_t^4)\right]$, formed from the same predictive moments and their parameter derivatives that the score already requires, at no extra cost. For a macroscopic algorithm the likelihood is Gaussian by construction, so this is the model's own Fisher and $\mathbf{C}$ is the information-matrix-equality test carried out in the model's own frame. Two properties follow. The Gaussian Fisher is positive semidefinite by construction, so the diagnostic never has to be rescued from a numerically indefinite anchor. And it is analytic, so it does not inherit the finite-difference noise of a numerical Hessian, which in this regime is severe: the finite-difference Fisher is widely indefinite per replicate, and only pooling across replicates recovers a definite matrix. By design the anchor is the Gaussian Fisher; the numerical Fisher we compute alongside it serves to gauge how faithful that Gaussian anchor is. Comparing the approximation to its own Fisher is not circular. $\mathbf{H}$ is the information the likelihood claims, $\mathbf{J}$ is the information it delivers against data from the true process, the two are computed from different objects, and their disagreement is the definition of misspecification.

The same matrix repairs what it measures. The sandwich covariance
\[ \Sigma=\mathbf{H}^{-1}\,\mathbf{J}\,\mathbf{H}^{-1}=\mathbf{H}^{-1/2}\,\mathbf{C}\,\mathbf{H}^{-1/2} \]
is the corrected parameter covariance: it substitutes the information the likelihood actually carries for the information it reported. This step, and the bias vector $b$ above, are first-order approximations, from a Taylor expansion of the score at the optimum with the empirical Hessian replaced by its expectation $-\mathbf{H}$ (see Supplementary Information). We verify the correction directly, without leaning on the expansion, by comparing the sandwich-corrected ellipse against the empirical covariance of the maximum-likelihood estimates (MLEs) over the simulated recordings (Figure 2). A diagnostic that also corrects the error bars it flags is a stronger result than one that only scores algorithms.

The distortion splits into two physically distinct parts (Figure 4--figure supplement 3). The sample distortion $\mathbf{C}_s=\mathbf{H}^{-1/2}\,\mathbf{J}_s\,\mathbf{H}^{-1/2}$, built from the within-interval score covariance $\mathbf{J}_s=\sum_t\mathrm{Cov}(s_t)$ with $s_t=\nabla_\theta\ell_t$ the score of interval $t$, measures per-sample departure from Gaussianity; to the leading corrections it is the third and fourth cumulants of the interval current, the non-Gaussian shape the Gaussian predictive cannot carry. The correlation distortion $\mathbf{R}=\mathbf{J}_s^{-1/2}\,\mathbf{J}\,\mathbf{J}_s^{-1/2}$ measures what the within-interval covariance misses, the cross-interval correlation of the score, which is exactly the local time correlation of the current that \cite{milescu2005maximum} named as the dominant error of non-recursive fitting; $\mathbf{R}=\mathbf{I}$ when the score carries no cross-interval dependence. These compose exactly. The exact identity is
\[ \mathbf{C}=\mathbf{K}\,\mathbf{R}\,\mathbf{K}^\top,\qquad \mathbf{K}=\mathbf{H}^{-1/2}\,\mathbf{J}_s^{1/2}, \]
which holds as algebra for any $\mathbf{H}$ and $\mathbf{J}_s$, since substituting the definitions collapses the chain. The factor $\mathbf{K}$ satisfies $\mathbf{K}\mathbf{K}^\top=\mathbf{C}_s$, so it is a legitimate square-root factor of the sample distortion, but it is not the symmetric principal square root, and the form $\mathbf{C}=\mathbf{C}_s^{1/2}\,\mathbf{R}\,\mathbf{C}_s^{1/2}$ holds only where $\mathbf{H}$ and $\mathbf{J}_s$ commute. What composes with no assumption at all is the information volume,
\[ \log\det\mathbf{C}=\log\det\mathbf{C}_s+\log\det\mathbf{R}, \]
an exact identity because the determinant is multiplicative. The per-parameter diagonal, the trace, the eigenvalues and the Frobenius norm do not compose across the split; only the volume does, and only the volume is decomposed in the maps.

The correlation distortion has a reader-facing translation. For any additive statistic $X_t$ with sum $L=\sum_t X_t$, the ratio $\kappa=\mathrm{Var}(L)\big/\sum_t\mathrm{Var}(X_t)$ is unity when the terms are independent and larger when they are positively correlated, and it converts to an effective sample size $T_\mathrm{eff}=T/\kappa$: a non-recursive likelihood believes it has $T$ independent intervals when it effectively has $T/\kappa$. This is the number from the diagnostic most worth quoting, and it is exact by definition. When $\mathbf{R}\approx\kappa\mathbf{I}$ the temporal error is isotropic and this scalar captures it; when $\mathbf{R}$ is anisotropic, or when $\mathbf{C}_s\neq\mathbf{I}$, the distortion cannot be reduced to a sample-size correction and the full matrix is needed.

Two conventions govern the edge cases. The direction is fixed as above, $C_{ii}>1$ being over-confidence, stated once so every verdict reads the same way. And the anchor can be near-singular where a parameter is unidentifiable, which by design happens once the open population relaxes (Figure 3--figure supplement 1); there we work in the retained eigenspace of $\mathbf{H}$ above a numerical cutoff and leave the quantity undefined, not finite, when that eigenspace is empty. Undefined is not zero, and the maps render those cells in grey to keep the two apart.
