\section*{Data and code availability}

All code needed to reproduce the study is openly available. The analysis pipeline that generates the figures, together with the simulation configurations that define every run, is deposited in a public repository (\texttt{macroir-validity}) and archived at Zenodo, with the DOI issued on acceptance. The simulation and inference engine is the \texttt{macro\_dr} software \cite{moffatt2025macroir}, referenced at the pinned version that produced the deposited data; each output file records the engine commit hash that generated it. No experimental data were used: every dataset in the paper is simulated, and is regenerable from the deposited configurations. % NOTE (LM): confirm the public repo name/URL and mint the Zenodo DOI at submission.

\section*{Acknowledgements}

The author thanks colleagues at INQUIMAE and the University of Buenos Aires for discussion. % NOTE (LM): add named acknowledgements (readers of drafts, compute support) before submission.

\section*{Author contributions}

L.M. conceived the study, developed the theory, implemented the algorithms, designed and ran the simulations, analysed the data, and wrote the manuscript.

\section*{Competing interests}

The author declares no competing interests.

\section*{Funding}

This work was supported by the Consejo Nacional de Investigaciones Cient\'ificas y T\'ecnicas (CONICET) and the Universidad de Buenos Aires. % NOTE (LM): add the specific grant agencies and award numbers (e.g. ANPCyT/PICT, UBACyT) before submission.
