---
title: "Figure 1 (merged paper) — one filter step over the window × recursion lattice: NR, INR, R, IR"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

Figure 1 for the merged paper. **Four columns**, the 2×2 of the two axes: `NR` neither, `INR` the
acquisition window only, `R` the recursive update only, `IR` both. `NR` and `INR` are non-recursive,
so their "Bayes & posterior" row reads "no update (open loop)", which is where the figure says what
open loop means. `MR` and `VR` are the partial interval corrections, off this lattice, and stay in
Figures 2, 3 and 4 where the claim that the ladder stops being an ordering needs them.
`figure_1_all.Rmd` (in `../archive/`) still renders every algorithm for exploration.

**THE TWO LEAST-SQUARES COLUMNS CAME OUT ON 2026-08-12 (Luciano), and the reason is measured, not
aesthetic.** On this recording the predicted mean of `LSE` is that of `NR` to `8.9e-16` and the
predicted mean of `ILSE` is that of `INR` to `2.2e-16` (`06_methods.tex`, measured off these same
dumps), so rows A and B of each least-squares column were a copy of its neighbour's; row C was blank;
and the whole content of the pair was one constant predictive sd, `3.536` pA for `LSE` and `3.559`
for `ILSE`, against `NR`'s `0.224` to `2.010`. Two columns for one fact, and a fact the caption now
states in one sentence ("`LSE` is the `NR` column with every error bar of the same width, `ILSE` the
`INR` column with a band of constant height"). What the removal buys is width: the panels go from
about 1.0 in to about 1.55 in, and `IR`'s two occupancy rows, their dashed covariance tie and the
two-tone boundary disc are the part of the figure that needed it. Predecessor kept for rewind at
`../archive/figure_1_superseded_20260812.Rmd` and `../archive/Figure_1_superseded_20260812.pdf`
(six columns, `LSE ILSE NR INR R IR`); the least-squares dumps stay on disk and the roster below
reaches them by name if the call is ever reversed.

Panels live in `figure_1_panels.R` (shared: it still defines all eight columns, and `build_figure`
sizes `IR`'s geometry from `length(cols)`).

**Two changes to the panels on 2026-08-12, both from looking at the four-column render (Luciano).**
The Markov arc's bow is set in inches by `geom_curve`, so widening the panels sent it off the bottom
of row A in the open-loop columns, taking the arrowhead with it: fixed by `MK_CURV` and an asymmetric
bottom pad on the row-A range, which are one fix in two places. And the x axis is no longer the
clock: ticks label the **acquisition interval** at its middle, dotted verticals mark the interval
boundaries in every panel, and the interval duration (2 ms) is now a caption fact. Both are
commented at their site in `figure_1_panels.R`; re-check the first if the roster length or the figure
width ever changes.

**No supplement.** The declared Figure 1—figure supplement 1 was withdrawn on 2026-08-11/12
(`papers/1_method/decisions.md`, the supplement table): it drew samples 0–4 where the body draws 3–4,
so it contained the body figure. Its `build_figure(0:4, ...)` call is gone from the chunk below and
`Figure_S1.pdf` is deleted. Render the wider crop into `../../../../tmp/` if you want to look at it.

(Re-running `ops/local/figure_1_plus_lse.macroir` redraws the whole recording — new random ensemble,
seed = 0 — so every panel changes trace-by-trace; figure 1 is illustrative, not quantitative, and no
number is quoted from it anywhere in the paper.)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r panels, include=FALSE}
source("figure_1_panels.R")
```

```{r roster}
# ROSTER OF FOUR, 2026-08-12: the window x recursion lattice. A column auto-drops if its
# per-interval diagnostic dump is not on disk.
COLS_TARGET <- c("NR", "INR", "R", "IR")
# Data-file token per column. Kept as a map because the least-squares tokens are crossed on disk
# (..._LSE is averaging = 1 and therefore ILSE; ..._LSE_av0 is LSE) and that crossing must stay
# resolved in one place for whoever puts those columns back.
.tok <- c(LSE = "LSE_av0", ILSE = "LSE", NR = "NR", INR = "INR", R = "R", IR = "IR")
COLS <- COLS_TARGET[file.exists(paste0("../data/figure_1_likelihood_diagnostic_", .tok[COLS_TARGET], ".csv"))]
message("figure_1 roster present: ", paste(COLS, collapse = ", "))
```

```{r figures}
# 7.0 in wide: the body figure is declared inside a fullwidth environment. 6.05 in tall so the float
# plus its legend fits the page (at 7.5 in LaTeX dropped the last lines of the caption).
build_figure(c(3, 4), "Figure_1.pdf", COLS, hgt = 6.05)   # acquisition intervals 3 and 4
```
