---
title: "Figure 1 (all algorithms) — one filter step across the whole roster"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

The exploratory version of Figure 1: **six columns**, every algorithm on disk, including the two
non-recursive members that paper 1 names but does not measure. Its job is to show the recursion axis
next to the conductance axis in one picture; the paper's figure is the four-column recursive ladder in
`figure_1.Rmd`.

It writes to its **own** files (`Figure_1_all.pdf`, `Figure_S1_all.pdf`) so it can never overwrite the
paper's figure. Everything below the roster lives in `figure_1_panels.R`, shared with `figure_1.Rmd`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r panels, include=FALSE}
source("figure_1_panels.R")
```

```{r roster}
# DATA keys, not printed names: the mean-non-recursive algorithm is NMR here and prints as MNR
# (the .DISP map in figure_1_panels.R does the translation).
COLS <- c("NR", "NMR", "R", "MR", "VR", "IR")
```

```{r figures}
build_figure(0:4,     "Figure_S1_all.pdf", COLS)   # samples 0-4 (0 to 10 ms of the 12 ms record)
build_figure(c(3, 4), "Figure_1_all.pdf",  COLS)   # the 6 to 10 ms zoom
```
