---
title: "Figure 3—figure supplement 1 — per-step information against score variance, all four identified parameters"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

<!--
REVIVED, NOT REWRITTEN (Luciano, 2026-07-31). This is the archived
`figure_3_supplements_superseded_20260731/figure_3_S3.Rmd` brought back on its own layout,
which the archive README already named as the candidate to start from. A first attempt at this
supplement put the four parameters as four COLOURS inside the parent's panels and was rejected on
sight: four curves and four ribbons in a 1.3 in panel is unreadable, and the two amplitude
parameters overlie during the pulse so one of them is simply invisible.

THE LAYOUT RULE THE WHOLE ARCHIVED SET OBEYS, and the reason it works: PARAMETER IS A ROW,
ALGORITHM IS A COLUMN, and no panel ever carries more than two series. All three archived
four-parameter figures are built that way -- this one, `Figure_S_score_acf_allparams.pdf` and
`Figure_S_score_mean.pdf` -- and none of them is crowded. The parent figure can afford colour for
the parameter because it draws two; a supplement that draws four cannot, and has to spend rows
instead. Rows are the cheap axis here: eLife sets no height limit, so 11 in of page is free while
panel width is not.

FOUR PARAMETER BANDS, each two rows, over all seven members: LSE / NR / INR / R / MR / VR / IR.
  INFO row (log10 y, no null): the per-step Gaussian Fisher information F_t = E[I_t] in blue and the
    score variance J_t = Var(s_t) in orange. They coincide wherever the member is calibrated per
    step, so the row is read as "do the two curves lie on each other", and it doubles as the map of
    WHERE in the recording each parameter is measured at all.
  LOG J/F row (linear, null 0, half height): log10(Var(s_t)/E[I_t]) with its bootstrap interval. A
    per-step-calibrated member sits on 0 at every step.
The pairing is deliberate and is why the info row has no null line: the ratio row is the test, the
info row is the magnitude the test is made of. Margins couple each pair (small gap inside a band,
larger gap between bands), inherited from the archived file.

WHAT THE FIGURE IS FOR, one objection: the body draws its score rows for k_off and N_ch, and three
sentences in the manuscript are about the other two. All three now land here.
  - Results and Discussion quote the per-step median log10(J_t/F_t) on k_on for all four members.
    Until this figure existed those numbers came from a notebook archived on 2026-07-31 and no
    longer a display item: numbers in the body with no figure behind them. The ratio rows are that
    statistic, and they reproduce the archived values to five decimals from the digests.
  - Results and Discussion claim the information about three parameters dies at the washout. The
    info rows are that claim, and they do not support it as written -- see MEASURED, below.
  - Diagnostics leaves a quantity undefined where the anchor is near-singular and points here for
    where that happens. That is the k_on and N_ch bands after the washout.

MEASURED 2026-07-31, AND IT DOES NOT MATCH THE SENTENCE THAT CITES THIS FIGURE. Mean F_t over the
washout window as a fraction of its value during the pulse, and over the last ten intervals (t
90-100 ms, three to four decay constants after the agonist is removed):

  member  parameter   F_wash/F_pulse   F over the last ten intervals
  IR      k_on            0.0039              7.9e-44
  IR      N_ch            0.0050              1.4e-21
  IR      i               0.028               0.66
  IR      k_off           0.655               1.13

k_on and N_ch reach the numerical floor; the unitary current does not. It falls by a factor of
thirty-six and levels at the same order as k_off, and the two survivals are of different kinds: 81%
of what k_off keeps comes from the mean, the shape of the decay it governs, while 71% of what i
keeps comes from the variance, the gating fluctuation of the few channels still open. N_ch's is
99.6% mean-borne and it is that mean derivative that vanishes, because conditioned on the past the
predicted current is the open count the filter already carries times i. The manuscript names the
unitary current with the dying pair in two places, flagged there as TODO and not edited here.

THE INFO ROW RUNS DOWN TO THE MASK, where the archived version stopped at 1e-3 and let the cliff
leave the panel. For IR the channel-number information runs 55.4, 10.5, 0.357, 1.67e-2, 6.9e-4,
2.8e-5, 1.05e-6 over six intervals after the washout, so its last point above the mask sits ON the
floor: with the axis ending there, the fall is seen to terminate rather than to be cut, and a reader
can tell a masked point from a clipped one.

LEAST SQUARES CARRIES NO N_ch BAND, by exclusion and not by absence of data, which is the parent's
rule (`figure_3.Rmd`, and `EXCLUDE_ROWS` in `figure_4_common.R`). With the gating fluctuations
unmodelled the channel count and the unitary current enter only through their product, and this
figure MEASURES that rather than assuming it: F_t over the pulse is 1076.55 for both, to every digit.
The panel is not left empty, which would read as a rendering failure; it carries the one-line reason.

A STRUCTURAL FACT THE FOUR-PARAMETER VIEW MAKES VISIBLE. In the least-squares column the ratio row
is the SAME curve in all four bands (median log10 = 0.183889 in every one). With one constant
variance the per-interval score is (y - mu) dmu/dtheta / sigma^2 and its information is
(dmu/dtheta)^2 / sigma^2, so the parameter cancels out of the ratio and what is left is the true
residual variance over the assumed one. Least squares therefore fails at the per-step level, and in
a way that measures its homoscedastic premise and nothing about any parameter -- a different failure
mode from the recursive family's, which is cross-time and only appears once the intervals are
summed. This is the data-level heteroscedasticity of the parent's row B, read at the score level.

CHANGES FROM THE ARCHIVED FILE, all four for consistency with the parent rather than preference:
  - reads the digests written by figure_3_digest.R (37 MB) instead of four ~1 GB dumps;
  - B = 400 percentile bootstrap replicates and 95% intervals, the figure_3 family convention,
    against the archived 200 and 90%;
  - the info row ends at the mask, as above;
  - the N_ch exclusion for least squares, which postdates the archived file (2026-07-30).

DATA: figures/data/digest/figure_3_digest_{LSE,NR,INR,R,MR,VR,IR}.rds. Cell N_ch = 100, Current_Noise = 1e-4,
interval = 0.1 tau, 1000 recordings, seed 20260722. Every number the caption quotes is printed by
the caption-numbers chunk on each knit and is never read off the plot.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(data.table)
library(patchwork)
```

```{r read}
# ALL SEVEN, widened from the body's four on 2026-07-31 once macro_INR had run. The reason is
# INR itself: it passes every per-interval check in the paper and fails only the accumulated one,
# so the rows of THIS figure are where it belongs. MR and VR come with it because they are what
# makes its number mean something -- both are recursive, both cost more, and both are WORSE here
# (per-step median log10(J_t/F_t) of -0.229 and -0.135 against INR's +0.007). Per-interval
# calibration is not monotone along the cost ladder, and with those two columns missing a reader
# would assume it is. NMR, the build that lost the interval-variance term, is not drawn anywhere.
ALGOS <- c("LSE", "NR", "INR", "R", "MR", "VR", "IR")
ALAB  <- c(LSE = "LSE", NR = "NR", INR = "INR", R = "R", MR = "MR", VR = "VR", IR = "IR")
# Wrapped at seven columns. A one-line subtitle is fine at four columns and 1.5 in per panel; at
# 0.9 in "Interval Non-Recursive" is wider than its column and runs into its neighbours.
FULL  <- c(LSE = "Least\nSquares", NR = "Non-\nRecursive", INR = "Interval Non-\nRecursive",
           R = "Recursive", MR = "Mean\nRecursive", VR = "Variance\nRecursive",
           IR = "Interval\nRecursive")
LEAD  <- ALGOS[1]                            # leftmost column carries the y titles and the tags
XTTL  <- ALGOS[ceiling(length(ALGOS) / 2)]   # the one column that carries the x title: the middle

PIDX  <- c(0L, 1L, 2L, 5L)                   # k_on, k_off, unitary_current, Num_ch_mean
PLAB  <- c("k[on]", "k[off]", "i", "N[ch]")
PTTL  <- list(bquote(k[on] * "  (opening rate)"), bquote(k[off] * "  (closing rate)"),
              bquote(italic(i) * "  (unitary current)"), bquote(N[ch] * "  (channel number)"))

FLOOR <- 1e-6      # absolute information mask; below it the derivatives are floating-point noise
B     <- 400       # bootstrap replicates over recordings, the figure_3 family convention
YLIM  <- c(-1, 1)  # log-ratio zoom; LSE breathes to about +0.4 and down past -1 in its tail
YCURVE <- c(FLOOR, 1e4)   # info row: from the mask to the top, so the washout cliff terminates

# THE EXCLUSION, and it is the parent's (figure_3.Rmd:145). Least squares does not model the gating
# fluctuations, the only thing separating the channel count from the unitary current, so its N_ch
# curve is its i curve. Measured in the caption chunk, not assumed.
EXCL <- data.table(algo = "LSE", param = "N[ch]")

dg <- setNames(lapply(ALGOS, function(a) {
  f <- sprintf("../data/digest/figure_3_digest_%s.rds", a)
  if (!file.exists(f)) stop("missing digest for ", a, ": run figure_3_digest.R ", a)
  readRDS(f)
}), ALGOS)

SI <- rbindlist(lapply(ALGOS, function(a)
  as.data.table(dg[[a]]$param)[param_index %in% PIDX,
    .(algo = a, param_index, simulation_index, sample_index, s, dm, dv, I)]))
tmap <- unique(as.data.table(dg[[ALGOS[1]]]$scalar)[, .(sample_index, t, agonist)])[order(sample_index)]
SI[, param := factor(PLAB[match(param_index, PIDX)], levels = PLAB)]
SI[, algo  := factor(algo, levels = ALGOS)]
SI_ALL <- copy(SI)                            # kept whole: the degeneracy check needs the dropped band
SI <- SI[!EXCL, on = .(algo, param)]
```

```{r compute-bootstrap}
set.seed(20260731)
cMean <- function(M) colMeans(M, na.rm = TRUE)
cVar  <- function(M) { n <- colSums(!is.na(M)); m <- colMeans(M, na.rm = TRUE)
                       (colMeans(M * M, na.rm = TRUE) - m^2) * n / (n - 1) }

# One resampling loop per (member, parameter). The ratio is resampled WHOLE, cVar(S)/cMean(I), so it
# carries the correlation between the score and its own information rather than treating F as known.
facet_ratio <- function(d) {
  Sm  <- as.matrix(dcast(d, simulation_index ~ sample_index, value.var = "s")[, -1])
  Im  <- as.matrix(dcast(d, simulation_index ~ sample_index, value.var = "I")[, -1])
  smp <- as.integer(colnames(Sm)); n <- nrow(Sm)
  Ft  <- cMean(Im); ratio <- cVar(Sm) / Ft
  reps <- matrix(NA_real_, B, ncol(Sm))
  for (b in seq_len(B)) {
    idx <- sample.int(n, n, replace = TRUE)
    reps[b, ] <- cVar(Sm[idx, , drop = FALSE]) / cMean(Im[idx, , drop = FALSE])
  }
  lrep <- log10(reps); lrep[!is.finite(lrep)] <- NA
  data.table(sample_index = smp, F = Ft, lr = log10(ratio),
             lo = apply(lrep, 2, quantile, 0.025, na.rm = TRUE),
             hi = apply(lrep, 2, quantile, 0.975, na.rm = TRUE))
}

R <- merge(SI[, facet_ratio(.SD), by = .(algo, param)], tmap, by = "sample_index")
R <- R[is.finite(lr) & F > FLOOR]              # drop only genuine zeros (the 0/0 steps)

# The two raw magnitude curves of the info row. Same mask, so the two rows of a band stop together.
curves <- melt(SI[, .(FIM = mean(I), score_var = var(s)), by = .(algo, param, sample_index)],
               measure.vars = c("FIM", "score_var"), variable.name = "quantity",
               value.name = "val")
curves <- merge(curves[is.finite(val) & val > FLOOR], tmap[, .(sample_index, t)], by = "sample_index")
curves[, quantity := factor(quantity, levels = c("FIM", "score_var"))]
```

```{r caption-numbers}
# --- the numbers the caption and the manuscript quote (LINT-SRC), computed from the digests in this
#     session; nothing is carried forward from the archived notebook.
cat("\n=== Figure 3 supplement 1 caption numbers (n_simulations =",
    uniqueN(SI$simulation_index), "| members:", paste(ALGOS, collapse = ", "), ") ===\n")
on  <- tmap[agonist > 0, sample_index]
off <- tmap[agonist == 0 & sample_index > max(on), sample_index]
cat(sprintf("agonist on over samples %d-%d (t %.1f-%.1f ms); washout window %d samples\n",
            min(on), max(on), tmap[sample_index == min(on), t],
            tmap[sample_index == max(on), t], length(off)))

# THE RATIO ROWS: the statistic Results and Discussion quote, and how often the interval covers 0.
tab <- R[, .(n_steps = .N, median_log10_JF = median(lr), frac_CI_covers_0 = mean(lo <= 0 & hi >= 0)),
         by = .(algo, param)]
setorder(tab, param, algo)
cat("\nratio rows -- per-step calibration over the informative steps:\n")
print(as.data.frame(tab), digits = 3, row.names = FALSE)
cat(sprintf("overall fraction of (step, member, parameter) whose interval covers 0: %.3f\n",
            R[, mean(lo <= 0 & hi >= 0)]))

# THE INFO ROWS: the washout claim, as a number. The mask is not applied here, since the whole point
# is how far below it the two dying parameters go.
WASH <- SI[, .(F = mean(I)), by = .(algo, param, sample_index)][
  , .(F_pulse = mean(F[sample_index %in% on]), F_wash = mean(F[sample_index %in% off]),
      F_last10 = mean(F[sample_index %in% tail(off, 10)])), by = .(algo, param)]
WASH[, ratio_wash_over_pulse := F_wash / F_pulse]
setorder(WASH, algo, param)
cat("\ninfo rows -- per-step information across the washout:\n")
print(as.data.frame(WASH), digits = 3, row.names = FALSE)

# WHY the two survivors survive differently: k_off through the mean, i through the variance. The
# decomposition F_t = (dmu)^2/v + (dv)^2/2v^2 needs y_var, which lives on the scalar side of the
# digest, so it is joined here rather than carried through the bootstrap.
YV <- rbindlist(lapply(ALGOS, function(a)
  as.data.table(dg[[a]]$scalar)[, .(algo = a, simulation_index, sample_index, y_var)]))
DEC <- merge(SI, YV, by = c("algo", "simulation_index", "sample_index"))[
  sample_index %in% tail(off, 10) & algo %in% c("R", "IR"),
  .(F = mean(dm^2 / y_var + 0.5 * dv^2 / y_var^2),
    share_from_mean = mean(dm^2 / y_var) / mean(dm^2 / y_var + 0.5 * dv^2 / y_var^2)),
  by = .(algo, param)]
setorder(DEC, algo, param)
cat("\n  where the surviving information comes from, over the last ten intervals:\n")
print(as.data.frame(DEC), digits = 3, row.names = FALSE)

# THE TWO CROSS-CHECKS THIS FIGURE OWES THE MANUSCRIPT.
lse <- SI_ALL[algo == "LSE", .(F_pulse = mean(I[sample_index %in% on])), by = param]
cat(sprintf("\nLSE degeneracy: F_pulse for i = %.6g and for N_ch = %.6g (identical = %s)\n",
            lse[param == "i", F_pulse], lse[param == "N[ch]", F_pulse],
            isTRUE(all.equal(lse[param == "i", F_pulse], lse[param == "N[ch]", F_pulse]))))
cat("LSE per-step median log10(J_t/F_t) by parameter (the parameter cancels, so these agree):\n  ",
    paste(sprintf("%s %.6f", tab[algo == "LSE", param], tab[algo == "LSE", median_log10_JF]),
          collapse = " | "), "\n")
cat("\nquoted in the manuscript (04_results.tex:77, 05_discussion.tex:31), for comparison:\n",
    " per-step median log10(J_t/F_t) on k_on: LSE +0.184, NR -0.038, R -0.181, IR +0.006\n",
    " same on k_off: R -0.145, IR +0.033\n")
```

```{r layout, fig.width = 7.0, fig.height = 11}
FIMc <- "#0072B2"; SCVc <- "#D55E00"; NUL <- "#009E73"; RATc <- "#D55E00"
tr   <- range(tmap$t)
agon <- tmap[agonist > 0]; x0 <- min(agon$t); x1 <- max(agon$t)
# Full-panel agonist band. On a log10 axis a negative floor is rejected (log10(<=0) = NaN drops the
# rect), so the info rows need a tiny positive floor; the linear ratio rows go negative.
# AGONIST BAND, -Inf/Inf and not +-1e10 (fixed 2026-07-31). A rect 2e10 tall clipped to a window of
# order one rasterises to NOTHING on a linear axis, so every LINEAR row of this figure was rendering
# unshaded while the log rows were fine -- inconsistent shading inside one figure, which reads as
# deliberate and was not. -Inf/Inf means "to the panel edge"; on a log axis it is NaN and the layer
# is dropped, hence the tiny positive lower bound there instead.
mkshade <- function(logr = FALSE) annotate("rect", xmin = x0, xmax = x1,
                  ymin = if (logr) 1e-300 else -Inf, ymax = Inf, fill = "grey85", alpha = .5)

colhead <- function(k) ggplot() + theme_void() + ggtitle(ALAB[[k]], subtitle = FULL[[k]]) +
  theme(plot.title = element_text(hjust = .5, size = 9),
        plot.subtitle = element_text(hjust = .5, size = 7))

# axis.text raised to 7 pt on 2026-08-05. theme_bw(base_size = 8) leaves it at rel(0.8) = 6.4 pt,
# and the old 7.0 x 11 in art was then scaled by 0.81 into the 5.65 in supplement column, so the
# numbers landed on page 45 at a measured 4.82 pt against eLife's 7 pt floor for base text. Drawing
# at 5.6 in removes the scaling; this puts the text where the floor needs it.
base <- function() theme_bw(base_size = 8) + theme(axis.text = element_text(size = 7))
# the shared legend is base text as well; theme_bw leaves it at rel(0.8) = 6.4 pt
LEGEND_TXT <- theme(legend.text = element_text(size = 7))

# mode "info" -> log10 magnitude axis, no null; "linear" -> signed axis with 0 forced onto the ticks.
# y title only on LEAD; x numbers on every bottom-of-block panel, x TITLE once, on XTTL.
deco <- function(p, col, xr, ylim, yq, ttl = NULL, tag = NULL, bottom = FALSE, xlab = NULL,
                 mode = "linear") {
  if (mode == "info") {
    p <- p + scale_y_log10(breaks = c(1e-4, 1, 1e4),
                           labels = expression(10^-4, 1, 10^4)) +
      coord_cartesian(xlim = xr, ylim = ylim)
  } else {
    brk <- pretty(ylim); brk <- sort(unique(c(0, brk))); brk <- brk[brk >= ylim[1] & brk <= ylim[2]]
    p <- p + scale_y_continuous(breaks = brk) + coord_cartesian(xlim = xr, ylim = ylim)
  }
  if (col == LEAD) p <- p + ylab(yq) + theme(axis.title.y = element_text(size = 8))
  else p <- p + theme(axis.title.y = element_blank(), axis.text.y = element_blank(),
                      axis.ticks.y = element_blank())
  if (bottom) {
    if (col == XTTL) p <- p + xlab(xlab) + theme(axis.title.x = element_text(size = 8))
    else p <- p + theme(axis.title.x = element_blank())
  } else p <- p + theme(axis.title.x = element_blank(), axis.text.x = element_blank())
  if (!is.null(tag)) p <- p + labs(tag = tag)
  if (!is.null(ttl)) p <- p + ggtitle(ttl) +
    theme(plot.title = element_text(hjust = 0, size = 8.5, face = "plain", margin = margin(b = 1)))
  # Tight coupling of each info/ratio pair: the info panel takes a small bottom margin and the ratio
  # panel a small top margin, so the pair reads as one band and the larger gap falls between bands.
  mrg <- if (mode == "info") margin(1.5, 3, 0.3, 3) else margin(0.3, 3, 1.5, 3)
  p + theme(panel.grid.minor = element_blank(), plot.margin = mrg)
}

# THE EXCLUDED CELL IS A PANEL, NOT A HOLE, and it goes through the SAME `deco` as every other one.
# Two reasons, the second learned the hard way. First, an empty box in a grid of thirty-two reads as
# a rendering failure, which is the misreading the parent's note says the legend has to prevent; the
# panel carries its reason instead. Second, the excluded cell is in the LEAD column, and in this
# layout the LEAD panel is what carries the band's title, its letter tag, the y title and the y
# numbers for the whole row. Replacing it with a theme_void() placeholder silently stripped all four
# from the channel-number band, and the three columns beside it lost their y axis with it. Building
# it as an ordinary panel with no data layers keeps the furniture and the grid regular.
excluded <- function(a, p) any(EXCL$algo == a & EXCL$param == p)
# Shortened for seven columns, where the four-column wording overran the panel and lost its first
# line off the left edge. The panel says WHICH fact; the caption says why.
EXCL_TXT <- "not separable\nfrom the\nunitary current"

mk_info <- function(i, a, tag = NULL, bottom = FALSE) {
  ex <- excluded(a, PLAB[i])
  g <- if (ex) {
    ggplot() + mkshade(TRUE) +
      annotate("text", x = mean(tr), y = 1, size = 6 / .pt, colour = "grey40", lineheight = .95,
               label = EXCL_TXT) + base()
  } else {
    ggplot(curves[param == PLAB[i] & algo == a], aes(t, val, colour = quantity)) + mkshade(TRUE) +
      geom_line(linewidth = .4) +
      scale_colour_manual(values = c(FIM = FIMc, score_var = SCVc), name = NULL,
                          labels = expression(F[t] == E * "[" * I[t] * "]" ~ "(Fisher information)",
                                              J[t] == Var(s[t]) ~ "(score variance)")) +
      base()
  }
  deco(g, a, tr, YCURVE, "information", ttl = if (a == LEAD) PTTL[[i]] else NULL,
       tag = tag, bottom = bottom, xlab = "time (ms)", mode = "info")
}
mk_ratio <- function(i, a, tag = NULL, bottom = FALSE) {
  ex <- excluded(a, PLAB[i])
  g <- ggplot(if (ex) data.table(t = numeric(), lr = numeric())
              else R[param == PLAB[i] & algo == a], aes(t, lr)) + mkshade(FALSE) +
    geom_hline(yintercept = 0, colour = NUL, linewidth = .4)
  if (!ex) g <- g + geom_ribbon(aes(ymin = lo, ymax = hi), fill = RATc, alpha = .25) +
    geom_line(colour = RATc, linewidth = .4)
  deco(g + base(), a, tr, YLIM, expression(log[10] ~ J[t] / F[t]), tag = tag, bottom = bottom,
       xlab = "time (ms)", mode = "linear")
}

# SPLIT INTO TWO SUPPLEMENTS, 2026-08-05. As one 7.0 x 11 in figure this could not be published:
# a figure supplement is placed in the 5.65 in text column (elife.cls:409-413), so the art arrived
# scaled by 0.81 and occupied 639 pt of the ~648 pt page, which left no room for the 273-word legend
# (154 of its words were dropped off the bottom of page 45) and shrank the axis numbers to 4.82 pt.
# Two halves, two parameters each, drawn at 5.6 in wide so nothing is scaled. The per-column panel
# width on the page is unchanged (7.0 x 0.81 = 5.67), so this costs no panel area; it buys the
# legend its room and the axis text its 7 pt. Letter tags run A-D in each half, restarting, because
# each half is now its own display item.
HALVES <- list(list(idx = 1:2, out = "Figure_3_S1.pdf",
                    what = "the opening rate and the closing rate"),
               list(idx = 3:4, out = "Figure_3_S2.pdf",
                    what = "the unitary current and the channel number"))
info_tags  <- c("A", "C")
ratio_tags <- c("B", "D")
for (hh in HALVES) {
  panels <- list()
  for (k in seq_along(hh$idx)) {
    i <- hh$idx[k]
    for (a in ALGOS) panels <- c(panels, list(mk_info (i, a, tag = if (a == LEAD) info_tags[k] else NULL)))
    for (a in ALGOS) panels <- c(panels, list(mk_ratio(i, a, tag = if (a == LEAD) ratio_tags[k] else NULL,
                                                       bottom = (k == length(hh$idx)))))
  }
  gh <- wrap_plots(c(lapply(ALGOS, colhead), panels), ncol = length(ALGOS),
                   heights = c(0.10, rep(c(1, 0.5), length(hh$idx))), guides = "collect") &
        theme(legend.position = "bottom", legend.text = element_text(size = 7))
  # 5.60 x 5.60 in: the width is the supplement column exactly, and the height is the old 11 in
  # scaled by this half's share of the row budget (3.10 of 6.10 units), so the panels keep their shape.
  ggsave(hh$out, gh, width = 5.6, height = 5.6)
  cat(sprintf("wrote %s: %d parameter bands x 2 rows, %d members, 5.6 x 5.6 in (%s)\n",
              hh$out, length(hh$idx), length(ALGOS), hh$what))
}
