---
title: "Figure 7 — the usage map: what a macroscopic recording can be asked for, over the design space"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

<!--
RENUMBERED 2026-08-10: this figure is now FIGURE 7. It was Figure 6 until the tau_int decision
panel was promoted into the body as Figure 5 and the corner figure became Figure 6. The archive
directory name (archive/figure_6_superseded_20260731) keeps its old number on purpose: it is dated.

WHAT THIS IS. One plane, N_ch x instrumental noise, partitioned by boundaries into the regimes that
decide what a macroscopic recording can support. Unlike every other figure in this set it is NOT a
heatmap: the content is the boundaries, so the cells are the input and the lines are the output.

THE ENCODING (Luciano, 2026-07-26) — one visual channel per meaning, no exceptions:

    COLOUR    = the PARAMETER the line is about:  k_on, k_off, i (unitary conductance), N_ch
    LINE TYPE = WHAT is being asked about it:
        solid   "can it be measured at all"  -> distortion-corrected SE crosses a factor 2
        dotted  "is ILSE's error bar honest"  -> ILSE's information distortion crosses 1.15
        dashed  "is IR's own error bar honest" -> IR's information distortion crosses 1.15
                 (this one is the macro/micro floor: below it the Gaussian closure is misspecified)
    RIBBON    = the same boundary swept over the sampling interval, Delta.k_off = 0.01 to 1.
                The heavy line is the reference Delta.k_off = 0.1. So the ribbon is not an error
                bar, it is "where your own sampling rate puts the boundary".

WHERE THE SE FAMILY COMES FROM, and why it is mixed: the RATES from ILSE, the AMPLITUDE from IR.
"Where does this parameter die" is a property of the experiment, so in principle it should be judged
by the best available method — but measured, IR and ILSE rate SEs agree to a few percent at N_ch >= 100
(the only gap is N_ch 10, where ILSE loses a factor 1.9 on k_on, and there both are already past the
threshold), while ILSE is the one with the high-noise coverage: at N_ch 1e4 IR stops at noise 1e5 and
ILSE runs to 1e7, so an IR-sourced rate boundary simply vanishes off the right of the plot. The
amplitude pair must come from IR, because ILSE holds unitary_current Fixed and never estimates i at
all — which is the structural point of the figure, not a coverage accident.

THE REGIONS, bottom to top at fixed N_ch:
    0  below the dashed line: the Gaussian closure itself fails. Labelled "closure fails", NOT
       "use micro": micro is not measured in this paper.
    1  i and N_ch both resolved AND ILSE's error bar wrong  -> IR buys two things
    2  amplitude gone, rates alive, ILSE still wrong        -> IR buys one thing (an honest CI)
    3  ILSE's error bar honest and rates alive              -> IR buys nothing, ILSE is par
    4  above the solid k_on line: nothing useful left
  The 2x2 cell "amplitude alive AND ILSE honest" is EMPTY at every measured cell, which is what makes
  these an ordered ladder rather than a table.

WHY k_on IS THE BINDING RATE: SE(k_on) ~ 3.1 x SE(k_off) everywhere measured, for both methods. One
concentration jump, so k_off is read twice (pulse rate and washout) while k_on comes only from their
difference. Region 4 therefore opens on k_on, ~1 decade of noise below where k_off alone would die,
and both lines are drawn so the reader can pick the criterion that matches what they need.

TRAPS, none of them re-derivable from the plot:
 - param_index is a POSITION IN THE FREE PARAMETER VECTOR, so the map differs by algorithm:
     macro: 0 on, 1 off, 2 unitary_current, 3 Current_Noise, 4 Current_Baseline, 5 Num_ch_mean
     ILSE  : 0 on, 1 off, 2 Current_Baseline, 3 Num_ch_mean   (the two Fixed ones are absent)
   A single global PIDX silently drops ILSE from anything keyed on index 5 and mislabels index 3.
 - the y axis is the RAW noise label, and current_noise = label/1000. Never divide it by N_ch: the
   diagonal boundaries are the finding, and dividing flattens them by construction.
 - N_ch is NOT capped to four columns. IR has runs at 5/20/50/200/500/2000/5000 and the closure floor
   is the best-sampled boundary in N_ch; with four columns its crossing (N_ch ~ 70-150) is invisible.
   Those extra columns are LOW-NOISE runs, so they help the floor and not the high-noise boundaries.
 - everything comes from battery_pool_G (theta_pool anchor), so the figure is internally consistent
   on one anchor. Bias (battery_sim_G) is not on this figure.

MEASURED, at Delta.k_off = 0.1 (the render log prints all of it):
  slopes, d log10 noise / d log10 N_ch:  ILSE-honest 1.01 (k_off) and 1.03 (N_ch), i.e. N_ch-independent
  in r = noise/N_ch;  SE(i) 1.11;  SE(N_ch) 1.61;  SE(k_off) 2.24 and SE(k_on) 2.21;  closure floor
  -0.56, NEGATIVE — instrumental noise gaussianises the emission, so it HELPS the closure while
  hurting everything else. That sign contrast is the second non-obvious thing in the figure.
  region 3 (ILSE is par) OPENS at N_ch 120 to 1368, the range being the strict-vs-loose criterion.
  the map CLOSES from below at N_ch 3 to 9, EXTRAPOLATED — exactly where the planned N_ch 2 and 5
  runs would land, which is the cheapest run this figure asks for.
  the loose criterion moves the SE boundaries up by a factor 9 (i, N_ch) to 19-21 (the rates) in
  noise, and the distortion boundaries down by a factor 2-3. That is why the vertices are ranges.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
suppressMessages({ library(tidyverse) })
# The three superseded drawings of this same measurement are still built, because two of them are
# the versions for CHECKING and the third carries the region vocabulary the text still uses, but
# they are written to the archive rather than into the figure directory: paper_both holds the
# figure set, and as of 2026-07-31 the figure is the `map` chunk at the foot of this notebook.
ARCH <- "../archive/figure_6_superseded_20260731"
dir.create(ARCH, showWarnings = FALSE, recursive = TRUE)
arch <- function(f) file.path(ARCH, f)
```

```{r data}
DATA_DIRS <- c("../data/1c2ae6f", "../data/87889e6", "../data/0ffbda7")
DIST <- "Probit_statistics_Likelihood_Gaussian_Information_Distortion"
SEC  <- "Probit_statistics_Gaussian_Distortion_Corrected_Covariance"
DREF <- 0.1                       # reference interval, in units of tau = 1/k_off
DRNG <- c(0.01, 1)                # interval range spanned by the ribbons

read_algo <- function(pat, algo, idx) {
  fs <- list.files(DATA_DIRS, pattern = pat, full.names = TRUE)
  if (!length(fs)) stop("figure_7: no files for ", algo)
  map_dfr(fs, function(p) read.csv(p, skip = 1, stringsAsFactors = FALSE) %>%
    filter(component_path %in% c(DIST, SEC), statistic == "value", probit == "mean",
           param_index == param_col, param_index %in% idx) %>%
    transmute(algo = algo, nch = Num_ch, noise = noise_in_conductance_tau, D = interval_in_tau,
              what = ifelse(component_path == DIST, "dist", "var"),
              par = names(idx)[match(param_index, idx)], v = value)) %>% distinct()
}
IR  <- read_algo("^figure_3_G_nch_[0-9]+_nsim_10000_macro_IR_noise_[0-9.eE+]+_battery_pool_G\\.csv$",
                 "IR",  c(kon = 0, koff = 1, i = 2, Nch = 5))
# RELABELLED 2026-08-11 (Luciano): this pattern matches `nonlinearsqr`, which is the
# INTERVAL-AVERAGED least-squares arm, ILSE (figure_4.Rmd:329-330; the data-file tokens are crossed
# on purpose, see figure_1.Rmd). The un-averaged arm is `nonlinearsqr_g` and is NOT read here. The
# label said ILSE and the classical arm of this map is therefore ILSE throughout.
ILSE <- read_algo("^figure_3_LSE_nch_[0-9]+_nsim_10000_nonlinearsqr_noise_[0-9.eE+]+_battery_pool_G\\.csv$",
                 "ILSE", c(kon = 0, koff = 1, Nch = 3))

# THE SECOND CLASSICAL ARM, added 2026-08-26. Until now the record said nothing about why this map
# uses ILSE rather than the un-averaged arm, and "the boundaries would not move" was asserted
# nowhere and established nowhere (04_results.tex, note retracted 2026-08-12). It is now measured.
# The un-averaged arm is `nonlinearsqr_g` (averaging = 0, the mean is the bare conductance half an
# interval in, NR's mean model) and it landed in ../data/a202e03 after this map was built: 30 cells
# on the Gaussian lane, N_ch 10 to 10^4 and noise 0.1 to 1e7, the same coverage the map needs, with
# the same param_index layout. It is read from its own directory list so that the DRAWN map keeps
# reading exactly the three frozen directories and cannot move: a202e03 holds zero files matching
# either pattern above, checked, so this is additive by construction.
DATA_DIRS_XARM <- c(DATA_DIRS, "../data/a202e03")
LSE <- local({
  fs <- list.files(DATA_DIRS_XARM,
        pattern = "^figure_3_LSE_nch_[0-9]+_nsim_10000_nonlinearsqr_g_noise_[0-9.eE+]+_battery_pool_G\\.csv$",
        full.names = TRUE)
  if (!length(fs)) NULL else map_dfr(fs, function(p) read.csv(p, skip = 1, stringsAsFactors = FALSE) %>%
    filter(component_path %in% c(DIST, SEC), statistic == "value", probit == "mean",
           param_index == param_col, param_index %in% c(kon = 0, koff = 1, Nch = 3)) %>%
    transmute(algo = "LSE", nch = Num_ch, noise = noise_in_conductance_tau, D = interval_in_tau,
              what = ifelse(component_path == DIST, "dist", "var"),
              par = c("kon", "koff", NA, "Nch")[param_index + 1], v = value)) %>%
    filter(!is.na(par)) %>% distinct()
})

cat("figure_7: noise levels on disk, per algorithm and N_ch\n")
print(as.data.frame(bind_rows(IR, ILSE) %>% distinct(algo, nch, noise) %>% count(algo, nch) %>%
      pivot_wider(names_from = algo, values_from = n, values_fill = 0)), row.names = FALSE)

# one row per (question, parameter, nch, D, noise). SEs are already in log10 units; the distortion is
# two-sided (IR under-disperses N_ch where it over-disperses k_off), hence abs(log()).
MET <- bind_rows(
  ILSE %>% filter(what == "var", par %in% c("kon", "koff")) %>%
          transmute(q = "se",    par, nch, D, noise, m = sqrt(pmax(v, 0))),
  IR  %>% filter(what == "var", par %in% c("i", "Nch")) %>%
          transmute(q = "se",    par, nch, D, noise, m = sqrt(pmax(v, 0))),
  ILSE %>% filter(what == "dist") %>% transmute(q = "lse",   par, nch, D, noise, m = abs(log(v))),
  IR  %>% filter(what == "dist") %>% transmute(q = "floor", par, nch, D, noise, m = abs(log(v))))

# thresholds and drawing rules, per question
QQ <- tribble(
  ~q,      ~t1,       ~t2,      ~rising, ~lty,   ~qlab,
  "se",    0.3,       1.0,      TRUE,    "solid","solid: parameter no longer measurable (SE > factor 2)",
  "lse",   log(1.15), log(1.3), FALSE,   "12",   "dotted: ILSE's reported error bar becomes honest (distortion < 1.15)",
  "floor", log(1.15), log(1.3), FALSE,   "42",   "dashed: IR's OWN error bar is wrong -> Gaussian closure fails (distortion > 1.15)")
PARS <- c(kon = "k_on", koff = "k_off", i = "i  (unitary conductance)", Nch = "N_ch")
# which (question, parameter) pairs are drawn. k_on/k_off carry region 4; i and N_ch carry the
# amplitude boundary; the honesty questions are asked of k_off and N_ch (i is Fixed for ILSE).
DRAW <- tribble(~q, ~par,
  "se","kon", "se","koff", "se","i", "se","Nch",
  "lse","koff", "lse","Nch",
  "floor","koff", "floor","Nch")

cross1 <- function(x, y, t, rising) {          # interpolate in log10(noise); NA when off-grid
  o <- order(x); x <- x[o]; y <- y[o]; ok <- is.finite(x) & is.finite(y); x <- x[ok]; y <- y[ok]
  if (length(x) < 2) return(NA_real_)
  s <- y - t
  if (all(s > 0) || all(s < 0)) return(NA_real_)
  k <- which(s[-length(s)] * s[-1] <= 0)
  if (!length(k)) return(NA_real_)
  i <- if (rising) k[1] else k[length(k)]
  x[i] + (t - y[i]) * (x[i + 1] - x[i]) / (y[i + 1] - y[i])
}
CR <- MET %>% inner_join(DRAW, by = c("q", "par")) %>% inner_join(QQ, by = "q") %>%
  group_by(q, par, qlab, lty, nch, D) %>%
  summarise(c1 = cross1(log10(noise), m, first(t1), first(rising)),
            c2 = cross1(log10(noise), m, first(t2), first(rising)), .groups = "drop") %>%
  mutate(lx = log10(nch), pf = factor(unname(PARS[par]), levels = unname(PARS)))

cat("\nfigure_7: boundary position (noise) at Delta.k_off =", DREF, ", strict criterion\n")
print(as.data.frame(CR %>% filter(D == DREF) %>% transmute(q, par, nch, noise = signif(10^c1, 3)) %>%
      pivot_wider(names_from = nch, values_from = noise, names_prefix = "N=")), row.names = FALSE)
cat("\nfigure_7: fitted log-log slopes (d log10 noise / d log10 N_ch) at Delta.k_off =", DREF, "\n")
print(as.data.frame(CR %>% filter(D == DREF, is.finite(c1)) %>% group_by(q, par) %>% filter(n() >= 3) %>%
      summarise(slope = round(coef(lm(c1 ~ lx))[2], 2), n = n(), .groups = "drop")), row.names = FALSE)
# BEST ACHIEVABLE SE, minimised over the whole noise sweep. A parameter whose best is already past
# the threshold is NEVER measurable at that N_ch, at ANY noise -- which is why some boundary lines
# terminate. That is a physical statement, not missing data, so it is printed and annotated.
BEST <- MET %>% filter(q == "se", D == DREF) %>% group_by(par, nch) %>%
  summarise(best = min(m), .groups = "drop")
cat("\nfigure_7: best achievable SE over the whole noise sweep (> 0.3 = never measurable there)\n")
print(as.data.frame(BEST %>% pivot_wider(names_from = par, values_from = best) %>%
      mutate(across(where(is.numeric), ~signif(.x, 3)))), row.names = FALSE)
NEVER <- BEST %>% group_by(par) %>% arrange(nch) %>%
  summarise(nch_min = if (min(best) > 0.3 || max(best) < 0.3) NA_real_ else
              10^approx(log10(best), log10(nch), xout = log10(0.3))$y, .groups = "drop")
cat("\nfigure_7: N_ch below which the parameter is UNREACHABLE at any noise\n")
print(as.data.frame(NEVER %>% mutate(nch_min = round(nch_min))), row.names = FALSE)

cat("\nfigure_7: how far the loose criterion moves each boundary (median factor in noise)\n")
print(as.data.frame(CR %>% filter(D == DREF, is.finite(c1), is.finite(c2)) %>% group_by(q, par) %>%
      summarise(factor = round(median(10^(c2 - c1)), 2), n = n(), .groups = "drop")), row.names = FALSE)

# DOES THE CLASSICAL BOUNDARY MOVE IF THE OTHER LEAST-SQUARES ARM IS USED? Measured, not drawn.
# The dotted boundary is recomputed on the un-averaged arm with the same threshold and the same
# interpolation, and the two are reported side by side. It is deliberately NOT added to the plot:
# at the reference interval the two land within a factor of 1.4 of each other on a boundary that
# sweeps five decades, so a second dotted line would sit on top of the first and the figure would
# carry a fourth line code for no information. The number belongs in the text instead.
if (!is.null(LSE)) {
  XA <- LSE %>% filter(what == "dist", par %in% c("koff", "Nch")) %>%
        transmute(algo = "LSE", par, nch, D, noise, m = abs(log(v))) %>%
    bind_rows(ILSE %>% filter(what == "dist", par %in% c("koff", "Nch")) %>%
              transmute(algo = "ILSE", par, nch, D, noise, m = abs(log(v)))) %>%
    group_by(algo, par, nch, D) %>%
    summarise(c1 = cross1(log10(noise), m, log(1.15), FALSE), .groups = "drop")
  cat("\nfigure_7: the honesty boundary on BOTH least-squares arms (noise), Delta.k_off =", DREF, "\n")
  print(as.data.frame(XA %>% filter(D == DREF) %>%
        transmute(algo, par, nch, noise = signif(10^c1, 3)) %>%
        pivot_wider(names_from = nch, values_from = noise, names_prefix = "N=")), row.names = FALSE)
  cat("\nfigure_7: LSE / ILSE ratio of that boundary, and the fitted slopes\n")
  print(as.data.frame(XA %>% filter(D == DREF, is.finite(c1)) %>%
        pivot_wider(names_from = algo, values_from = c1) %>%
        mutate(ratio = signif(10^(LSE - ILSE), 3)) %>% select(par, nch, ratio)), row.names = FALSE)
  print(as.data.frame(XA %>% filter(D == DREF, is.finite(c1)) %>% mutate(lx = log10(nch)) %>%
        group_by(algo, par) %>% filter(n() >= 3) %>%
        summarise(slope = round(coef(lm(c1 ~ lx))[2], 3), n = n(), .groups = "drop")), row.names = FALSE)
  cat("\nfigure_7: the same over the ribbon range of Delta (", DRNG[1], "to", DRNG[2], "), in noise\n")
  print(as.data.frame(XA %>% filter(D >= DRNG[1], D <= DRNG[2], is.finite(c1)) %>%
        group_by(algo, par, nch) %>%
        summarise(lo = signif(10^min(c1), 3), hi = signif(10^max(c1), 3), .groups = "drop")),
        row.names = FALSE)
} else cat("\nfigure_7: the un-averaged least-squares arm is not on disk in", DATA_DIRS_XARM, "\n")
```

```{r geometry}
# the two vertices, each a RANGE bracketed by the strict and loose criteria on the SE(k_on) line
vx <- function(qa, pa, qb, pb, cb) {
  a <- CR %>% filter(q == qa, par == pa, D == DREF, is.finite(c1))
  b <- CR %>% filter(q == qb, par == pb, D == DREF, is.finite(.data[[cb]]))
  if (nrow(a) < 2 || nrow(b) < 2) return(NULL)
  fa <- coef(lm(c1 ~ lx, a)); fb <- coef(lm(as.formula(paste(cb, "~ lx")), b))
  x <- unname((fb[1] - fa[1]) / (fa[2] - fb[2]))
  tibble(lx = x, ly = unname(fa[1] + fa[2] * x))
}
v_open  <- list(vx("lse", "koff", "se", "kon", "c2"), vx("lse", "koff", "se", "kon", "c1"))
v_close <- list(vx("floor", "koff", "se", "kon", "c2"), vx("floor", "koff", "se", "kon", "c1"))
rng <- function(v) if (any(vapply(v, is.null, TRUE))) "NA" else
  sprintf("%.0f-%.0f", 10^min(v[[1]]$lx, v[[2]]$lx), 10^max(v[[1]]$lx, v[[2]]$lx))
cat("\nregion 3 (ILSE is par) OPENS at N_ch =", rng(v_open), "(range = the two criteria)\n")
cat("the map CLOSES from below at N_ch =", rng(v_close),
    "-- EXTRAPOLATED, this is what the N_ch 2 and 5 runs would pin\n")

PCOL <- c("k_on" = "#E69F00", "k_off" = "#009E73",
          "i  (unitary conductance)" = "#0072B2", "N_ch" = "#CC79A7")
# x reaches 3e7 so oocyte-scale preparations fit; the simulated span is marked on the plot
XLIM <- log10(c(2.5, 4e9)); YLIM <- log10(c(0.008, 5e7))
NOISE_MIN <- 0.05            # the lowest noise level anywhere in the sweep; below it we know nothing
xbrk <- log10(10^(1:9))
xlab <- c("10", expression(10^2), expression(10^3), expression(10^4), expression(10^5),
          expression(10^6), expression(10^7), expression(10^8), expression(10^9))
ybrk <- log10(10^(-2:7))
ylab <- c("0.01", "0.1", "1", "10", expression(10^2), expression(10^3), expression(10^4),
          expression(10^5), expression(10^6), expression(10^7))
```

```{r render, fig.width = 7, fig.height = 6.6}
# ribbon = the boundary swept over the interval range; line = the reference interval
RIB <- CR %>% filter(D >= DRNG[1], D <= DRNG[2], is.finite(c1)) %>%
  group_by(q, par, pf, lty, nch, lx) %>%
  summarise(lo = min(c1), hi = max(c1), n = n(), .groups = "drop") %>% filter(n >= 2)
LIN <- CR %>% filter(D == DREF, is.finite(c1)) %>% arrange(lx)
# A vertex is an N_ch RANGE, so it is drawn as a vertical strip. Deliberately NOT a line: line type
# already encodes the question, and a black line here read as a fifth boundary.
strip <- function(v, col, alpha) if (any(vapply(v, is.null, TRUE))) NULL else
  annotate("rect", xmin = min(v[[1]]$lx, v[[2]]$lx), xmax = max(v[[1]]$lx, v[[2]]$lx),
           ymin = YLIM[1], ymax = YLIM[2], fill = col, alpha = alpha)
vpts <- function(v, fill) if (any(vapply(v, is.null, TRUE))) NULL else
  geom_point(data = bind_rows(v), aes(lx, ly), shape = 21, fill = fill, colour = "grey20",
             size = 1.9, inherit.aes = FALSE)

g <- ggplot() +
  geom_ribbon(data = RIB, aes(x = lx, ymin = lo, ymax = hi, fill = pf, group = paste(q, par)),
              alpha = 0.16, show.legend = FALSE) +
  geom_line(data = LIN, aes(x = lx, y = c1, colour = pf, linetype = qlab, group = paste(q, par)),
            linewidth = 0.75) +
  geom_point(data = LIN, aes(x = lx, y = c1, colour = pf), size = 0.8, show.legend = FALSE) +
  scale_colour_manual(values = PCOL, name = NULL, drop = FALSE) +
  scale_fill_manual(values = PCOL, drop = FALSE) +
  scale_linetype_manual(values = setNames(QQ$lty, QQ$qlab), name = NULL) +
  guides(colour = guide_legend(order = 1, ncol = 2, keywidth = unit(15, "pt")),
         linetype = guide_legend(order = 2, ncol = 1, keywidth = unit(24, "pt"))) +
  strip(v_open, "grey35", 0.13) + vpts(v_open, "black") +
  strip(v_close, "grey55", 0.10) + vpts(v_close, "white") +
  { if (!any(vapply(v_open, is.null, TRUE)))
      annotate("text", x = mean(c(v_open[[1]]$lx, v_open[[2]]$lx)), y = YLIM[2] - 0.28,
               hjust = 0.5, size = 2.2, lineheight = .95,
               label = paste0("region 3 opens somewhere\nin this strip: N_ch ", rng(v_open))) } +
  { if (!any(vapply(v_close, is.null, TRUE)))
      annotate("text", x = log10(4.3), y = log10(0.042), hjust = 0, size = 2.2, colour = "grey35",
               lineheight = .95, label = paste0("map closes here, EXTRAPOLATED (N_ch ", rng(v_close),
                                                "):\nthis is what the N_ch 2 and 5 runs would pin")) } +
  annotate("text", x = log10(230),  y = log10(0.031), label = "0   Gaussian closure fails",
           size = 2.4, hjust = 0.5, colour = "grey25") +
  annotate("text", x = log10(2200), y = log10(2.5),   label = "1   i + N_ch + honest CI",
           size = 2.8, hjust = 0.5, fontface = "bold") +
  annotate("text", x = log10(38),   y = log10(3.2),   label = "i but\nnot N_ch", size = 2.1,
           hjust = 0.5, colour = "grey25", lineheight = .95) +
  annotate("text", x = log10(3000), y = log10(6e3),   label = "2   rates + honest CI\n(amplitude needs a prior)",
           size = 2.5, hjust = 0.5, lineheight = .95) +
  annotate("text", x = log10(6000), y = log10(2.4e5), label = "3   ILSE\nis par", size = 2.4,
           hjust = 0.5, lineheight = .95) +
  { if (is.finite(NEVER$nch_min[NEVER$par == "kon"]))
      annotate("text", x = log10(12), y = log10(3e3),
               hjust = 0, size = 2.1, colour = PCOL[["k_on"]], lineheight = .95,
               label = sprintf("k_on unreachable at ANY noise below N_ch %.0f:\nthe line ENDS here, it does not leave the grid",
                               NEVER$nch_min[NEVER$par == "kon"])) } +
  annotate("text", x = log10(90),   y = log10(4e6),   label = "4   nothing estimable",
           size = 2.6, hjust = 0.5, colour = "grey25") +
  scale_x_continuous(breaks = xbrk, labels = xlab, limits = XLIM, expand = c(0, 0)) +
  scale_y_continuous(breaks = ybrk, labels = ylab, limits = YLIM, expand = c(0, 0)) +
  labs(x = expression(N[ch]), y = "dimensionless instrumental noise") +
  theme_bw(base_size = 8, base_family = "Helvetica") +
  theme(panel.grid.minor = element_blank(), axis.title = element_text(size = 8.5),
        legend.position = "bottom", legend.justification = "left", legend.box = "vertical",
        legend.box.just = "left", legend.text = element_text(size = 6.9),
        legend.margin = margin(0, 0, 0, 0), legend.spacing.y = unit(1, "pt"),
        legend.box.margin = margin(-4, 0, 0, -6), legend.key.height = unit(9, "pt"),
        plot.margin = margin(4, 6, 2, 4))
g
# Was Figure_6.pdf until 2026-07-31. The figure is now the `map` chunk at the foot of this notebook;
# this one stays as the version for CHECKING, since the boundaries are here as lines with their
# ribbons and their vertices rather than as the edges of filled regions.
ggsave(arch("Figure_6_lines.pdf"), g, width = 7.0, height = 6.6, device = cairo_pdf)  # cairo: keeps the Delta glyph
ggsave(arch("Figure_6_lines.png"), g, width = 7.0, height = 6.6, dpi = 140)
```

```{r regions, fig.width = 7.6, fig.height = 5.2}
# ---- experimental preparations, from sourced numbers ----------------------------------------------
# Provenance and every figure below: docs/bibliography/recording_configurations/SOURCES.md, with the
# Axon Guide chapter saved beside it. THE CONVERSION IS VERIFIED, from qmodel.h:3740,
#     e = Current_Noise * fs / n_samples   ->  variance per recorded point = Current_Noise / Delta,
# so Current_Noise is a noise POWER SPECTRAL DENSITY and, with label = 1000 * Current_Noise,
#     label = 10 * k_off * sigma_raw^2 / (fs * i^2),
# which at Delta = 0.1 tau = 1 ms and k_off = 100 /s becomes label = (sigma per 1 ms point / i)^2.
# White noise of RMS sigma_B in a bandwidth B gives sigma_Delta^2 = sigma_B^2 / (2 B Delta).
# THE SCALING LEDGER (Luciano, 2026-07-27: which quantities scale linearly and which quadratically).
# Every exponent below is load-bearing, so they are written out rather than left implicit:
#   LINEAR     the macroscopic current in N_ch and in i, separately:  I = N_ch * i * P_open
#   LINEAR     the gating variance in the macroscopic current, and in i as well, because
#              Var_gate = N_ch i^2 p(1-p) = I * i * (1-p)
#   LINEAR     label in the noise power spectral density S, and in k_off at fixed bandwidth
#   QUADRATIC  label in 1/i: the signal is measured in units of i and the rig is not
#   QUADRATIC  S in the rms noise:                        S = sigma^2 / B
#   QUADRATIC  the noise PSD in membrane capacitance for the series-resistance mechanism, because the
#              current noise density is e_n * 2 pi f * C_m, so S grows as C_m^2 (Axon Guide ch. 12)
# The last two together are the reason a bigger preparation loses. N_ch grows LINEARLY with membrane
# area while the instrumental variance grows QUADRATICALLY with it, so r = label/N_ch worsens about
# linearly in area. The excised patch escapes that scaling because its noise is set by the pipette
# dielectric and the headstage rather than by C_m, which is why going from a patch to a whole cell buys
# channels almost for free, and why going from a whole cell to an oocyte does not.
#
# IN OBSERVABLES the same statement is shorter. Substituting N_ch = I/(i P_open),
#     r = label / N_ch = 10 * S * P_open / (tau * i * I)
# so r falls LINEARLY in the macroscopic current and LINEARLY in the unitary current. Once the clamp
# ceiling pins I at I_max, raising expression can only lower i, and r ~ N_ch: every preparation has an
# OPTIMAL channel number N* = I_max / (i_max * P_open), and expressing past it makes things worse.
# Printed below per preparation. The oocyte at 5e7 is an order of magnitude past its own N*.
# THREE PREPARATIONS, AND THE CHANNEL NOW LIVES INSIDE EACH BOX rather than outside all three (scheme
# by Luciano, 2026-07-27). The channel contributes two ranges, shared by every preparation:
#     i   = 1-10 pA, i.e. about 10-170 pS at 60-100 mV of driving force. Covers AMPA at the bottom and
#           glycine or NMDA at the top; it clips CFTR below and BK (25-30 pA at +100 mV) above.
#     tau = 1-30 ms. Covers nAChR, AMPA, P2X2 and glycine; it clips Na_V inactivation (0.3-0.5 ms)
#           below and NMDA or GABA-A deactivation (50-500 ms) above.
# Both extreme corners are real receptors, glycine at high driving force for (10 pA, 30 ms) and AMPA
# for (1 pA, 1 ms), so the rectangle is not empty where it matters.
# tau IS CLIPPED PER PREPARATION, because it is not a free choice: it is the fastest relaxation the
# configuration can resolve. Fast perfusion of an excised patch reaches 0.1 ms, so the whole channel
# range is open to it; whole cell and oocyte are held to about 10 ms by solution exchange and by clamp
# settling, so their window is 10-30 ms. OPEN QUESTION, and it is not cosmetic: for a BATH concentration
# jump the oocyte's exchange is 1-10 s, which would put it outside the 1-30 ms window altogether. The
# 10 ms floor used here is the voltage-step or fast-perfusion case.
#
# TWO CONSTRAINTS CUT CORNERS OFF THE RECTANGLES, both straight out of the ledger above.
#  1. THE CLAMP WINDOW, both ends of it. N_ch and i are independent as biology and coupled as an
#     experiment, because I = N_ch i P_open has to land inside what the configuration can record.
#     THE CEILING is what the clamp can hold: about 2 nA in a patch, 10 nA in whole cell (at
#     Rs = 10 MOhm that is already 100 mV of error), a few tens of uA in TEVC. Because label goes as
#     1/i^2, substituting i = I_max/(N_ch P_open) gives
#         label >= 10 * S * (N_ch P_open)^2 / (tau * I_max^2),
#     so the FLOOR of each box rises as N_ch^2 once the ceiling binds. N* below is where that starts,
#     and it is where each preparation is at its best.
#     THE MINIMUM is the other half of the same statement (Luciano, 2026-07-27): a record has to be
#     big enough to be a record at all, about 0.1 uA in an oocyte, where the endogenous conductances
#     of an uninjected cell are already tens of nA. So FEW CHANNELS HAVE TO BE BIG ONES,
#         i >= I_min / (N_ch P_open),
#     and by the same substitution the CEILING of each box FALLS as N_ch^2 below
#     N_min* = I_min/(i_lo P_open). Two slope-2 diagonals, mirror images, one cutting the bottom right
#     corner and one the top left. Below n_feas = I_min/(i_hi P_open) the configuration is infeasible
#     at any i: even the largest channel available cannot reach I_min. That is a DERIVED left edge for
#     the box, and for the oocyte it lands at 2e4 channels, well below the expression floor used here,
#     so the two statements do not conflict. Both are printed.
#     WHAT THIS COSTS THE ARGUMENT: it removes exactly the corners where the map was least favourable.
#     The patch's worst r falls from 2.5 to 0.63 and the oocyte's from 90 to 22, so the minimum-current
#     cut is not a decoration, it is the difference between crossing two boundaries and one.
#  2. THE SPECTRAL EXPONENT. The y axis is a WHITE noise level, so placing a real rig on it means asking
#     which white level reproduces the variance of a Delta-averaged point. For a rig whose PSD goes as
#     f^p, S_eq = S(B0) * (B/B0)^p with B = 1/Delta = 10/tau, hence
#         label = 10 * S(B0) * (10/(tau B0))^p / (tau i^2),
#     which is 1/tau for white noise, 1/tau^2 at p = 1, 1/tau^3 at p = 2. Taken as p = 1 for the patch
#     (pipette dielectric noise, PSD linear in f) and p = 2 for whole cell and oocyte (series-resistance
#     noise, PSD quadratic in f), both from Axon Guide ch. 12. All three forms AGREE at tau = 10 ms,
#     which is exactly where the sigma values were quoted (B0 = 1 kHz), so the correction pivots on the
#     anchor instead of moving it. It bites at the edges: the patch's fast edge rises 10x and every slow
#     edge drops by 3^p. The white-noise floor is printed alongside so the swing stays visible.
#     LIMITATION: a variance match is not an autocorrelation match. A rising spectrum is high-pass, so a
#     real rig's consecutive Delta-points are slightly anticorrelated where the simulation's are
#     independent. Nothing here claims more than the variance.
#
# NOISE, sigma at B0 = 1 kHz, all from SOURCES.md:
#   patch       0.1-0.5 pA. Brackets the one hard measurement in that file (S = 1.04e-4 pA^2/Hz gives
#               sqrt(1.04e-4 * 1000) = 0.32 pA at 1 kHz), and the true floor may be 2-3x lower.
#   whole cell  1 pA, the author's value; the measured spread in SOURCES.md is 0.4-3 pA at 1 kHz.
#   oocyte      0.5-3 nA. The two ends are the two oocyte methods: 0.5 nA at 1 kHz is S = 250 pA^2/Hz,
#               which is the cut-open figure (1 nA at 3-5 kHz), and 3 nA is conventional TEVC measured
#               off a published trace (Nagel 2025 Nat Commun 16:10367 Fig 8b, S = 1.09e4 pA^2/Hz).
# EXPRESSION LEVELS as set in the table below: patch 10-300, whole cell 20-1e4, oocyte 1e5-5e7 channels
# (Luciano, 2026-07-27). The literature sweep behind them is now written up in
# docs/bibliography/recording_configurations/EXPRESSION_LEVELS.md (23 agents, 159 claims, 134 confirmed).
# THREE OF THESE EDGES HAVE A SOURCED ALTERNATIVE and none is applied here, because they are decisions:
#   patch Imin 10 -> 5 pA        Khanra 2026 Nat Commun 17:3789, a stated Methods acceptance criterion.
#                                Costs the argument: the upper-left diagonal disappears, i returns to
#                                1 pA at the left edge and the patch's worst r rises from 0.63 to 2.5.
#   whole cell Imin 100 -> 500 pA  Li 2017 PLoS ONE, measured over 42,272 wells. Helps: forces i = 3.3 pA
#                                at n_lo and drops that corner about 11x.
#   oocyte n_lo 1e5 -> 8e5       Fujiwara & Kubo 2004 J Physiol 558:31. The 1e5 has no source and is
#                                unreachable at i = 1 pA anyway (50 nA, under the oocyte's own Imin).
# THE PROJECT'S OWN P2X2 STUDY, precisely: it supplies no number on the COUNT axis. On the NOISE axis the
# patch row's 0.1-0.5 pA was originally chosen to bracket that record's S = 1.04e-4 pA^2/Hz, so the
# exclusion was incomplete. That is now closed independently: six laboratories with the seal attached and
# the bandwidth stated give 0.05-0.35 pA at 1 kHz (Smith/Wang/Howe 2000, Prieto & Wollmuth 2010, Rycroft
# & Gibb 2002, Ding & Sachs 1999, Premkumar & Auerbach 1997, McGee 2025; EXPRESSION_LEVELS.md section 2g),
# a band that contains the 0.1-0.5 used here. The patch row no longer rests on the excluded record.
# THE NOISE AXIS IS REFERRED TO k_off, NOT TO THE SAMPLING INTERVAL (Luciano, 2026-07-27, correcting a
# derivation of mine that went round in circles). The definition, and the only thing here that is not
# open, is
#     label = 10 * Current_Noise * k_off / i^2
# with Current_Noise the noise power spectral density S. There is NO Delta in it. Writing instead
# sigma_Delta^2 = S/Delta and substituting Delta = 0.1 tau gives label = (sigma_Delta/i)^2, which is
# algebraically the same thing but only because 10*k_off = 1/Delta at that convention, AND it needs the
# noise to be white to write sigma_Delta^2 = S/Delta at all. The k_off form needs no such assumption, so
# it is the one used here. With S = sigma_B^2/B evaluated at B = k_off = 1/tau,
#     label = 10 * (sigma_{1/tau} / i)^2
# so tau = 1 ms means a 1 kHz band, not 10 kHz.
# CONSEQUENCE WORTH MORE THAN THE CORRECTION: the y axis does not depend on Delta at all. The sampling
# interval moves the BOUNDARIES, which is what DRNG already sweeps, and it does not move a preparation's
# vertical position. The two axes are independent rather than coupled through a convention.
#
# WHAT IS NOT KNOWN, stated plainly so it is not re-derived a fourth time. label needs S at that row's
# own k_off. That is the rig's noise SPECTRUM, and it is not measured for any of the three preparations:
# each has ONE rms figure at ONE band. How many decades of label a decade of tau buys is therefore NOT
# DERIVABLE from anything in this file. Three attempts at it by assuming a spectrum shape gave 9.9, 4.05
# and 1.39 decades, all arithmetic on an invented spectrum and none of them evidence. Until two measured
# sigma values per preparation exist, tau contributes NOTHING to the height of a box here, and the boxes
# below are drawn from the i range and the rig-to-rig spread alone. The gap is left visible on purpose.
# HOW TO MEASURE IT, when it is done: TRACE_WIDTH_METHOD.md, on a segment with NO channel activity. At
# long tau the gating noise takes over, and this file's own excised-patch record has 99 % of the variance
# of a 13.7 ms point coming from unliganded gating, a factor 93 over the instrument. Measured carelessly,
# "the noise at the tau of the kinetics" returns the quantity being estimated.
# ONE MORE MISSPECIFICATION, unrelated to the closure: the simulation adds Gaussian WHITE noise. A real
# series-resistance-limited rig is Gaussian and coloured, hence high-pass, so its consecutive samples are
# anticorrelated where the simulation's are independent. Matching S at one band does not match that, and
# nothing here models it.
B0    <- 1000     # Hz, retained only to convert the sigma values that were quoted at 1 kHz
I_LO  <- 1        # pA, smallest unitary current considered
I_HI  <- 10       # pA, largest
POPEN <- 0.5      # the model's open probability at saturation
lbl <- function(S, tau, i) 10 * S / (tau * i^2)   # = 10 * S_tilde, the sweep's own label. See note above
psd <- function(sig, B) sig^2 / B                 # rms in a bandwidth B -> power spectral density
# Imin for the patch (10 pA) and whole cell (100 pA) are placeholders of the same kind as the oocyte's
# 0.1 uA and have NOT been signed off; the oocyte value is the author's. Whole cell's N_min* = 200 falls
# below its own n_lo, so that row is unaffected either way.
# sig_lo and sig_hi are the RIG-TO-RIG spread at the band named in sig_band, which is a different thing
# from the band-to-band change of sigma across each row's own tau range. Only the first is known, so tau
# contributes NOTHING to the height of a box here. That is the honest current state and it is deliberate:
# the missing measurement shows up as a box that is too short rather than as an invented number.
# WHOLE CELL, tau = 0.5-1000 ms and N_ch = 20-1e4, settled with Luciano 2026-07-27. The N_ch edges are
# DERIVED, not assumed: 20 = 100 pA/(10 pA x 0.5) and 1e4 = 5 nA/(1 pA x 0.5), so they encode Imin and
# Imax rather than an expression guess. Imax = 5 nA is his (at Rs compensated to 2 MOhm that is already
# 20 mV of error); Imin = 100 pA is still a placeholder. STILL OPEN on this row: the i floor, since Cav
# with physiological Ca carries 0.3-0.5 pA and label goes as 1/i^2, so dropping I_LO to 0.3 would raise
# every box top 11x; and whether the row is HEK or native, since a 70-150 pF neuron or myocyte is 3-10x
# noisier than the 20-30 pF HEK cell the 1 pA came from, and the fast half of that literature is native.
PREP <- tribble(~name,           ~n_lo, ~n_hi, ~sig_lo, ~sig_hi, ~sig_band, ~tau_lo, ~tau_hi, ~Imin, ~Imax,
  "excised outside-out patch",      10,   300,     0.1,     0.5,      1000,    1e-3,    3e-2,   1e1,   2e3,
  "whole cell",                     20,   1e4,     1.0,     1.0,      1000,    5e-4,       1,   1e2,   5e3,
  "oocyte, two-electrode",         1e5,   5e7,     5e2,     3e3,      1000,    1e-2,    3e-2,   1e5,   3e7) %>%
  mutate(xlo = log10(n_lo), xhi = log10(n_hi),
         S_lo = psd(sig_lo, sig_band), S_hi = psd(sig_hi, sig_band),
         nstar = pmin(pmax(Imax / (I_HI * POPEN), n_lo), n_hi),   # ceiling starts to bind: floor turns up
         nmin_star = pmin(pmax(Imin / (I_LO * POPEN), n_lo), n_hi), # minimum binds: ceiling turns down
         nfeas = Imin / (I_HI * POPEN),                           # derived left edge: below it, no i works
         yhi = log10(lbl(S_hi, tau_lo, pmax(I_LO, Imin / (n_hi * POPEN)))),   # top at the right edge
         yhi_l = log10(lbl(S_hi, tau_lo, pmax(I_LO, Imin / (n_lo * POPEN)))), # top at the left edge
         ylo = log10(lbl(S_lo, tau_hi, I_HI)),                                # flat part of the floor
         ylo_r = log10(lbl(S_lo, tau_hi, pmin(I_HI, Imax / (n_hi * POPEN)))))
# The footprint, both diagonals: the top follows i >= Imin/(N P_open) and the floor i <= Imax/(N P_open).
# Where the two meet the box PINCHES to zero height: there is no freedom left in i, so nothing to draw.
POLY <- bind_rows(lapply(seq_len(nrow(PREP)), function(k) {
  p  <- PREP[k, ]
  xs <- sort(unique(c(seq(p$xlo, p$xhi, length.out = 120), log10(p$nstar), log10(p$nmin_star))))
  xs <- xs[xs >= p$xlo & xs <= p$xhi]
  N  <- 10^xs
  itop <- pmax(I_LO, p$Imin / (N * POPEN))   # few channels have to be big ones
  ibot <- pmin(I_HI, p$Imax / (N * POPEN))   # many channels have to be small ones
  tibble(name = p$name, x = c(xs, rev(xs)),
         y = c(log10(lbl(p$S_hi, p$tau_lo, itop)), rev(log10(lbl(p$S_lo, p$tau_hi, ibot)))))
}))
# r over the whole footprint, not just its corners: the extrema sit where the diagonals start to bind.
RR <- POLY %>% mutate(r = y - x) %>% group_by(name) %>% summarise(r_lo = min(r), r_hi = max(r))
cat("\nfigure_7: three preparations, in the sweep's own label = 10 * S * k_off / i^2 = 10 * S_tilde\n")
print(as.data.frame(PREP %>% left_join(RR, by = "name") %>% transmute(name,
      N_ch    = sprintf("%.3g - %.3g", n_lo, n_hi),
      tau_ms  = sprintf("%.3g - %.3g", tau_lo * 1e3, tau_hi * 1e3),
      sigma_pA_at = sprintf("%.3g - %.3g @ %.0f Hz", sig_lo, sig_hi, sig_band),
      label   = sprintf("%.3g - %.3g", 10^pmin(ylo, ylo_r), 10^pmax(yhi, yhi_l)),
      S_tilde = sprintf("%.3g - %.3g", 10^pmin(ylo, ylo_r)/10, 10^pmax(yhi, yhi_l)/10),
      r       = sprintf("%.2g - %.2g", 10^r_lo, 10^r_hi),
      N_opt   = sprintf("%.3g", nstar),
      N_feasible_above = sprintf("%.3g", nfeas))), row.names = FALSE)

# ---- the one real experiment: WITHDRAWN 2026-07-27 --------------------------------------------------
# The box drawn here used to be Moffatt & Hume 2007's own P2X2 outside-out record, measured from
# projects/p2x2/data/experiments/, at N_ch = 1777-2961 with an arrow for a x93 unliganded-gating floor.
# It is withdrawn on the author's instruction, for two reasons that both bear on the numbers: the
# records SUM three traces per response, so the 1777 pA peak is not one patch's current (it would be
# about 590 pA, and N_ch about 590-990, with sigma smaller by sqrt(3) so that S falls 3x too, which
# moves the box diagonally without changing r), and the preparation shows unliganded gating, so its
# x93 noise floor is a property of that preparation and not a generic feature of an excised patch.
# The excised-patch box above is therefore carried entirely by other laboratories.
# TO RESTORE: set SHOW_EXP to TRUE and rebuild EXP here. Left as a switch rather than deleted, because
# whether the figure keeps one worked example is still open.
SHOW_EXP <- FALSE

# THE SIMPLIFIED SIBLING — the version for reading; the boundary figure above is the version for
# checking. Same measurements, read as areas. Four rules it follows, each fixing something the first
# draft got wrong:
#  1. ORDERED fill, not categorical. The regions are a ladder (the likelihood buys two things, then
#     one, then none, then the experiment is empty), so the ramp goes dark -> pale in that order and
#     the reader sees the ordering without reading a word. Region 0 is a DIFFERENT kind of failure
#     (the model, not the experiment), so it gets its own hue rather than a rung of the ramp.
#  2. Labels say WHAT TO DO, not what the diagnostic did. This is a usage map. The four boundary
#     labels are deliberately PARALLEL, because they are the 2x2 written out: two of them say
#     "<method> becomes honest" (the likelihood at the bottom, least squares higher up) and two say
#     "<parameter> no longer measurable" (the amplitude pair, then k_off). Phrasing one of them
#     differently -- "the closure fails", "i and N_ch stop separating" -- hid that structure.
#  3. MEASURED vs EXTRAPOLATED is visible: heavy line inside the simulated N_ch span, faint dotted
#     outside, dots on the actual crossings. The first draft drew fitted power laws to the plot edges
#     and so claimed confidence it does not have — the floor, in particular, is measured only over
#     N_ch 10-100.
#  4. Everything clamped to the panel, so no ribbon can vanish against the axis limit.
fitb <- function(qq, pp) {
  d <- CR %>% filter(q == qq, par == pp, D == DREF, is.finite(c1))
  if (nrow(d) < 2) return(NULL)
  list(co = coef(lm(c1 ~ lx, d)), lo = min(d$lx), hi = max(d$lx), pts = d)
}
FB <- list(floor = fitb("floor", "koff"), ilost = fitb("se", "i"),
           hon = fitb("lse", "koff"),     dead = fitb("se", "koff"))
ev  <- function(f, x) f$co[1] + f$co[2] * x
cl  <- function(y) pmin(pmax(y, YLIM[1]), YLIM[2])
gx  <- seq(XLIM[1], XLIM[2], length.out = 500)

# the preparation footprints, clipped to the panel. N* is marked only where it is VISIBLE: for whole
# cell the whole floor sits below the panel, so its upturn cannot be shown and marking it would put a
# dot on the axis that means nothing.
POLYC <- POLY %>% mutate(y = cl(y))
NSTAR <- PREP %>% filter(n_hi > nstar * 1.001, ylo > YLIM[1]) %>%
  transmute(name, x = log10(nstar), y = ylo)
OFFB  <- PREP %>% filter(ylo < YLIM[1])
cat("\nfigure_7: preparation footprints, clipped to the panel\n")
print(as.data.frame(PREP %>% transmute(name,
      floor_true = sprintf("%.3g", 10^ylo), floor_at_right_edge = sprintf("%.3g", 10^ylo_r),
      top = sprintf("%.3g", 10^yhi),
      off_panel_bottom = ifelse(ylo < YLIM[1], "yes", "no"))), row.names = FALSE)

# the floor is measured to be ABSENT above N_ch 200 (IR is calibrated at every noise there), which is
# a measurement and not an extrapolation: pin it to the panel floor rather than extending the fit.
y_floor <- ifelse(gx > log10(150), log10(NOISE_MIN), cl(ev(FB$floor, gx)))
y_i     <- cl(ev(FB$ilost, gx)); y_hon <- cl(ev(FB$hon, gx)); y_dead <- cl(ev(FB$dead, gx))
R <- tibble(x = gx,
  r0_lo = YLIM[1],                r0_hi = pmax(YLIM[1], pmin(y_floor, y_i)),
  r1_lo = pmin(y_floor, y_i),     r1_hi = pmax(pmin(y_floor, y_i), y_i),
  r2_lo = y_i,                    r2_hi = pmax(y_i, pmin(y_hon, y_dead)),
  r3_lo = pmin(y_hon, y_dead),    r3_hi = pmax(pmin(y_hon, y_dead), y_dead),
  r4_lo = y_dead,                 r4_hi = YLIM[2]) %>%
  mutate(across(starts_with("r"), cl))

# 0 = its own hue; 1..4 = one ramp. Regions 1 and 2 lightened 2026-07-31: #3E9B62 was dark enough
# that the grey15 measured points and boundary lines drawn on top of it stopped reading.
RC   <- c("#B9A7D0", "#8FCBA6", "#C8E6D2", "#FBE08A", "#EFEFEF")
RLAB <- c("0  the likelihood itself fails",
          "1  amplitude and rates, with an honest interval",
          "2  rates only, with an honest interval",
          "3  least squares is at par",
          "4  nothing measurable")
# One mapped layer instead of five hardcoded fills, so the regions get a key. The in-panel prose
# says what to do; the key says which numbered region the text and the caption are talking about.
RLONG <- dplyr::bind_rows(lapply(0:4, function(k) tibble::tibble(
  x   = R$x,
  lo  = R[[paste0("r", k, "_lo")]],
  hi  = R[[paste0("r", k, "_hi")]],
  reg = factor(RLAB[k + 1], levels = RLAB))))
MEAS <- range(c(FB$ilost$pts$lx, FB$hon$pts$lx, FB$dead$pts$lx, FB$floor$pts$lx))

lineseg <- function(f, cap = NULL) {                      # heavy where measured, faint outside
  hi <- if (is.null(cap)) f$hi else min(f$hi, cap)
  xin <- seq(f$lo, hi, length.out = 120)
  xout <- gx[(gx < f$lo & gx > f$lo - 0.5) | (gx > hi & gx < hi + 0.5)]
  list(geom_line(data = tibble(x = xout, y = cl(ev(f, xout))), aes(x, y),
                 linewidth = 0.3, linetype = "12", colour = "grey30"),
       geom_line(data = tibble(x = xin, y = cl(ev(f, xin))), aes(x, y),
                 linewidth = 0.7, colour = "grey15"),
       geom_point(data = f$pts %>% mutate(y = cl(c1)), aes(lx, y), size = 1.1, colour = "grey15"))
}
lab <- function(x, y, t, sz = 3.0, fc = "plain", col = "grey10", hj = 0.5)
  annotate("text", x = log10(x), y = log10(y), label = t, size = sz, fontface = fc,
           colour = col, hjust = hj, lineheight = 1.0)
# boundary labels set ALONG each line. The angle is the fitted slope carried into device space:
# x spans diff(XLIM) decades over ~6.3 in and y spans diff(YLIM) over ~4.3 in. Approximate on
# purpose -- this is a concept map, and these labels are what make it readable without the caption.
ASPECT <- (diff(XLIM) / diff(YLIM)) * (4.1 / 6.4)   # panel is ~6.3 x 4.1 in once the caption left
blab <- function(x, f, t, dy = 0.3, sz = 2.35) {
  lx <- log10(x)
  annotate("text", x = lx, y = ev(f, lx) + dy, label = t, size = sz, colour = "grey15",
           angle = atan(f$co[2] * ASPECT) * 180 / pi, hjust = 0.5)
}

gr <- ggplot() +
  geom_ribbon(data = RLONG, aes(x = x, ymin = lo, ymax = hi, fill = reg), alpha = 0.9) +
  scale_fill_manual(values = stats::setNames(RC, RLAB), name = NULL,
                    # the region-4 fill is near-white, so the key needs an outline to exist against the page
                    guide = guide_legend(nrow = 2, byrow = TRUE, order = 1,
                                         override.aes = list(colour = "grey55", linewidth = .2))) +
  lineseg(FB$floor, cap = log10(150)) + lineseg(FB$ilost) +
  lineseg(FB$hon) + lineseg(FB$dead) +
  annotate("rect", xmin = XLIM[1], xmax = XLIM[2], ymin = YLIM[1], ymax = log10(NOISE_MIN),
           fill = "white", alpha = 0.72) +
  annotate("segment", x = XLIM[1], xend = XLIM[2], y = log10(NOISE_MIN), yend = log10(NOISE_MIN),
           colour = "grey45", linewidth = 0.3, linetype = "22") +
  annotate("text", x = XLIM[2] - 0.05, y = log10(0.0155), hjust = 1, size = 2.2, colour = "grey40",
           label = "no data below here: 0.05 is the lowest noise simulated") +
  annotate("segment", x = MEAS, xend = MEAS, y = YLIM[1], yend = YLIM[2],
           colour = "grey35", linewidth = 0.3, linetype = "22") +
  annotate("text", x = mean(MEAS), y = YLIM[2] - 0.22, label = "simulated range of N_ch",
           size = 2.2, colour = "grey35") +
  blab(30,   FB$floor, "the likelihood becomes honest",  0.45) +
  blab(600,  FB$ilost, "i and N_ch no longer measurable", 0.32) +
  blab(1600, FB$hon,   "least squares becomes honest",  0.34) +
  blab(55,   FB$dead,  "k_off no longer measurable",    0.36) +
  geom_polygon(data = POLYC, aes(x, y, group = name),
               fill = "white", alpha = 0.35, inherit.aes = FALSE) +
  geom_polygon(data = POLYC, aes(x, y, group = name),
               fill = NA, colour = "grey10", linewidth = 0.45, inherit.aes = FALSE) +
  geom_text(data = PREP %>% mutate(ly = pmin(yhi, YLIM[2] - 0.15) + 0.30,
                                   lx = c((xlo[1]+xhi[1])/2, xlo[2] - 0.10, (xlo[3]+xhi[3])/2),
                                   hj = c(0.5, 1, 0.5)),
            aes(x = lx, y = ly, label = name, hjust = hj),
            size = 2.3, fontface = "bold", colour = "grey5", inherit.aes = FALSE) +
  # N*: past it the clamp ceiling forces i down and the floor turns upward as N_ch^2
  { if (nrow(NSTAR)) geom_point(data = NSTAR, aes(x, y), shape = 21, fill = "white",
                                colour = "grey10", size = 1.5, inherit.aes = FALSE) } +
  { if (nrow(NSTAR)) geom_text(data = NSTAR %>% slice(1), aes(x, y),
      label = "N*: past here the clamp ceiling\nforces i down and the floor\nturns upward as N_ch^2",
      hjust = 1, nudge_x = -0.12, nudge_y = -0.40, size = 2.05, colour = "grey20",
      lineheight = .95, inherit.aes = FALSE) } +
  # boxes whose floor leaves the panel: quieter than anything the sweep simulated
  { if (nrow(OFFB)) geom_segment(data = OFFB, aes(x = (xlo + xhi) / 2, xend = (xlo + xhi) / 2,
                                                 y = YLIM[1] + 0.42, yend = YLIM[1] + 0.06),
                                 colour = "grey20", linewidth = .4, inherit.aes = FALSE,
                                 arrow = arrow(length = unit(3, "pt"), type = "closed")) } +
  { if (nrow(OFFB)) annotate("text", x = log10(3.2), y = YLIM[1] + 0.62, hjust = 0, size = 2.05,
      colour = "grey20", lineheight = .95, label = sprintf(
        "at i = 10 pA and tau = 30 ms these floors leave the panel (label %s):\nquieter than the 0.05 the sweep reaches, so the map cannot answer there",
        paste(sprintf("%.1g", 10^OFFB$ylo), collapse = " and "))) } +
  annotate("text", x = log10(1.7e9), y = log10(30), hjust = 1, size = 2.1, colour = "grey15",
           lineheight = .95,
           label = "the channel is INSIDE each box now:\ni = 1-10 pA and tau = 1-30 ms, clipped\nto what the configuration resolves.\nHeight is 1/(tau^p+1 i^2), p = 1 patch, 2 the rest") +
  lab(1.2e7, 0.9, "USE THE LIKELIHOOD\nyou get the conductance and the channel count,\nand the error bar is right", 2.8, "bold") +
  lab(4e6, 3e6, "USE THE LIKELIHOOD, WITH A PRIOR\non the amplitude pair. Least squares\nis still overconfident here", 2.5) +
  lab(1500, 3e6, "LEAST SQUARES\nIS ENOUGH", 2.5, "bold") +
  lab(40, 1e6, "REDESIGN THE EXPERIMENT\nnothing useful is left", 2.7) +
  lab(17, 0.105, "THE LIKELIHOOD FAILS", 2.2, "bold", "grey10") +
  scale_x_continuous(breaks = xbrk, labels = xlab, limits = XLIM, expand = c(0, 0)) +
  scale_y_continuous(breaks = ybrk, labels = ylab, limits = YLIM, expand = c(0, 0),
    sec.axis = sec_axis(~ . / 2, name = "baseline noise, RMS per 1 ms point (pA, for i = 1 pA)",
                        breaks = log10(c(0.1, 1, 10, 100, 1000)),
                        labels = c("0.1", "1", "10", "100", "1000"))) +
  labs(x = expression(N[ch]), y = "dimensionless instrumental noise") +
  theme_bw(base_size = 8, base_family = "Helvetica") +
  theme(panel.grid.major = element_line(colour = "white", linewidth = .18),
        panel.grid.minor = element_blank(), axis.title = element_text(size = 8.5),
        legend.position = "bottom", legend.text = element_text(size = 7),
        legend.key.size = unit(9, "pt"), legend.margin = margin(0, 0, 0, 0),
        legend.box.spacing = unit(3, "pt"),
        plot.margin = margin(4, 6, 2, 4))
gr
ggsave(arch("Figure_6_regions.pdf"), gr, width = 7.6, height = 5.6, device = cairo_pdf)
ggsave(arch("Figure_6_regions.png"), gr, width = 7.6, height = 5.6, dpi = 140)
```

```{r frontiers, fig.width = 7, fig.height = 5.6}
# THE FRONTIER VERSION (Luciano, 2026-07-26). The criterion is the author's convention and the reader
# cannot resolve it, so instead of picking one and drawing a line, BOTH criteria are drawn and the
# strip between them is left blank: a transition zone where the answer depends on how strict you are.
# What survives as solid colour is what holds under BOTH criteria. The weakness of the sharp-line
# version (the region-3 vertex moves from N_ch 120 to 1368 with the criterion) becomes the content:
# region 3 simply has no core until the two bands separate.
#
# ONE thing is deliberately NOT banded: the sampling interval. Delta is a design choice the READER
# makes and knows, so it stays fixed at the reference and is stated. Banding it too would widen every
# frontier by another 0.6-1 decade and swallow regions 2 and 3 entirely -- which is itself a measured
# fact, and is why the interval gets its own treatment in the companion figure instead.
#
# REGION 4 IS DEFINED BY k_off, THE LAST PARAMETER STANDING, and that is not a convenience. Defining
# it as "not every rate is measurable" (i.e. by k_on) makes the whole plane below N_ch ~52 read
# "nothing useful left", which is plainly false: at 10 channels you still get k_off, the conductance
# and the channel count. Losing k_on is its own rung, not the end of the ladder, so it is drawn as an
# internal boundary with its own label, and the fact that k_on is out of reach at ANY noise below
# N_ch ~52 is marked on the axis rather than swallowing four regions.
fit2 <- function(qq, pp) {
  d <- CR %>% filter(q == qq, par == pp, D == DREF)
  f1 <- d %>% filter(is.finite(c1)); f2 <- d %>% filter(is.finite(c2))
  if (nrow(f1) < 2) return(NULL)
  list(a = coef(lm(c1 ~ lx, f1)),
       b = if (nrow(f2) >= 2) coef(lm(c2 ~ lx, f2)) else coef(lm(c1 ~ lx, f1)),
       pts = f1)
}
FF <- list(floor = fit2("floor", "koff"), ilost = fit2("se", "i"),
           hon = fit2("lse", "koff"),     dead = fit2("se", "koff"),
           kon = fit2("se", "kon"))
gx <- seq(XLIM[1], XLIM[2], length.out = 500)
bnd <- function(f) { y1 <- f$a[1] + f$a[2] * gx; y2 <- f$b[1] + f$b[2] * gx
                     list(lo = cl(pmin(y1, y2)), hi = cl(pmax(y1, y2))) }
Bf <- bnd(FF$floor); Bi <- bnd(FF$ilost); Bh <- bnd(FF$hon); Bd <- bnd(FF$dead)
Bk <- bnd(FF$kon)
off <- gx > log10(150)                       # closure measured ABSENT above N_ch ~150
Bf$lo[off] <- log10(NOISE_MIN); Bf$hi[off] <- log10(NOISE_MIN)
# NOTE, and it cost a wrong claim (caught by Luciano, 2026-07-26): NEVER$nch_min is computed at the
# REFERENCE interval only, so "k_on is unreachable below N_ch 52" was an artefact of fixing Delta.
# Minimising over Delta as well, IR reaches SE(k_on) = 0.267 at N_ch 10 with Delta.k_off = 0.01
# (0.375 at 0.1, 0.511 at 1). ILSE does NOT: its best is 0.708 at every Delta and every noise. So at
# few channels the rate boundary is genuinely method-dependent, a factor 2.7, and must not be sourced
# from ILSE there. No "unreachable" marker is drawn.

FR <- tibble(x = gx,
  r0_lo = YLIM[1],  r0_hi = Bf$lo,
  r1_lo = Bf$hi,    r1_hi = pmax(Bf$hi, Bi$lo),
  r2_lo = Bi$hi,    r2_hi = pmax(Bi$hi, pmin(Bh$lo, Bd$lo)),
  r3_lo = Bh$hi,    r3_hi = pmax(Bh$hi, Bd$lo),
  r4_lo = Bd$hi,    r4_hi = YLIM[2]) %>% mutate(across(starts_with("r"), cl))
CORE <- c("#B9A7D0", "#3E9B62", "#9ED4A8", "#F9D77A", "#D6D6D6")
fillc <- function(k) geom_ribbon(data = FR, aes(x = x, ymin = .data[[paste0("r", k, "_lo")]],
                                                ymax = .data[[paste0("r", k, "_hi")]]),
                                 fill = CORE[k + 1], alpha = 0.95)
edgeb <- function(B) list(
  geom_line(data = tibble(x = gx, y = B$lo), aes(x, y), linewidth = .26, linetype = "22", colour = "grey35"),
  geom_line(data = tibble(x = gx, y = B$hi), aes(x, y), linewidth = .26, linetype = "22", colour = "grey35"))

gf <- ggplot() +
  lapply(0:4, fillc) + edgeb(Bf) + edgeb(Bi) + edgeb(Bh) + edgeb(Bd) +
  # k_on is NOT drawn here, and that is a scope decision, not an omission. Its boundary CROSSES the
  # others: at N_ch 1e3 the amplitude pair is lost first (noise ~3e2) and k_on much later (~1.4e4),
  # but at N_ch 10 the order flips. So k_on is not a rung of this ladder. It is not a method choice
  # either -- IR and ILSE lose it within a few percent of each other except at few channels -- it is a
  # property of the PROTOCOL: one concentration jump determines k_on only through the difference
  # between the pulse rate and the washout rate. If you need k_on, the fix is a second concentration,
  # not a better likelihood. That belongs in the text, not on a usage map.
  lapply(list(FF$floor, FF$ilost, FF$hon, FF$dead), function(f)
    geom_point(data = f$pts %>% mutate(y = cl(c1)), aes(lx, y), size = 1.0, colour = "grey20")) +
  annotate("rect", xmin = XLIM[1], xmax = XLIM[2], ymin = YLIM[1], ymax = log10(NOISE_MIN),
           fill = "white", alpha = 0.75) +
  annotate("segment", x = XLIM[1], xend = XLIM[2], y = log10(NOISE_MIN), yend = log10(NOISE_MIN),
           colour = "grey45", linewidth = .3, linetype = "22") +
  annotate("text", x = XLIM[2] - 0.05, y = log10(0.0125), hjust = 1, size = 2.1, colour = "grey40",
           label = "no data below here: 0.05 is the lowest noise simulated") +
  annotate("segment", x = MEAS, xend = MEAS, y = YLIM[1], yend = YLIM[2],
           colour = "grey40", linewidth = .3, linetype = "22") +
  annotate("text", x = mean(MEAS), y = YLIM[2] - 0.20, label = "simulated range of N_ch",
           size = 2.1, colour = "grey40") +
  geom_polygon(data = POLYC, aes(x, y, group = name),
               fill = "white", alpha = 0.35, inherit.aes = FALSE) +
  geom_polygon(data = POLYC, aes(x, y, group = name),
               fill = NA, colour = "grey10", linewidth = 0.45, inherit.aes = FALSE) +
  geom_text(data = PREP %>% mutate(ly = pmin(yhi, YLIM[2] - 0.15) + 0.30,
                                   lx = c((xlo[1]+xhi[1])/2, xlo[2] - 0.10, (xlo[3]+xhi[3])/2),
                                   hj = c(0.5, 1, 0.5)),
            aes(x = lx, y = ly, label = name, hjust = hj),
            size = 2.3, fontface = "bold", colour = "grey5", inherit.aes = FALSE) +
  lab(3000, 1.6, "USE THE LIKELIHOOD\nconductance, channel count, and\nan error bar you can trust", 2.9, "bold") +
  lab(5000, 1.6e4, "USE THE LIKELIHOOD,\nWITH A PRIOR on the amplitude", 2.5) +
  lab(13000, 2e6, "LEAST SQUARES\nIS ENOUGH", 2.5, "bold", hj = 1) +
  lab(55, 3e6, "REDESIGN THE EXPERIMENT", 2.6) +
  lab(17, 0.105, "THE LIKELIHOOD FAILS", 2.2, "bold") +
  lab(24, 2.2e4, "blank strips are the transitions:\nthere the answer depends on\nhow strict a criterion you use",
      2.15, "italic", "grey25", 0) +
  scale_x_continuous(breaks = xbrk, labels = xlab, limits = XLIM, expand = c(0, 0)) +
  scale_y_continuous(breaks = ybrk, labels = ylab, limits = YLIM, expand = c(0, 0)) +
  labs(x = expression(N[ch]), y = "dimensionless instrumental noise") +
  theme_bw(base_size = 8, base_family = "Helvetica") +
  theme(panel.grid.major = element_line(colour = "white", linewidth = .18),
        panel.grid.minor = element_blank(), axis.title = element_text(size = 8.5),
        plot.margin = margin(4, 6, 2, 4))
gf
ggsave(arch("Figure_6_frontiers.pdf"), gf, width = 7.0, height = 5.6, device = cairo_pdf)
ggsave(arch("Figure_6_frontiers.png"), gf, width = 7.0, height = 5.6, dpi = 140)
```

## The figure

THE MAP. Same plane, same cells, read as areas. Every cell is filled with WHICH parameters survive,
not with how many: the count conflates the two cases the figure exists to separate. Over the 80
cells at the reference interval the surviving subsets are

    all four   {k_on, k_off, i, N_ch}   51        two   {k_off, i}      11   <- amplitude present
    three      {k_off, i, N_ch}          1        two   {k_on, k_off}    5   <- amplitude gone
    one        {k_off}                   6        none  {}               6

and the two rows labelled "two" are opposite answers to the only question that decides which tool is
needed. LEAST SQUARES HOLDS THE UNITARY CURRENT FIXED AND CANNOT SEPARATE THE CHANNEL COUNT FROM IT,
so where the amplitude is recoverable it is not a candidate at all, and not because its interval is
wrong: it does not estimate those parameters. Where it IS a candidate, the rates, the only question
left is whose interval is calibrated. Five regions:

    violet      the likelihood's own interval is not calibrated, whatever the yield; it overrides
    green       i and k_off at least, usually all four: only the likelihood can deliver this
    pale green  the rates only, and least squares still reports an interval that is too narrow
    amber       the rates only, and least squares is calibrated too
    near-white  nothing measurable

COUNTING WAS ADOPTED (revisions AF to AL) BECAUSE THE BOUNDARIES ARE NOT NESTED: slopes run +1.1 to
+2.2, so they cross and a ladder of count levels inverts. At THIS resolution they are nested, with no
counterexample in the 80 cells: amplitude implies rates, N_ch never survives without i, no subset
carries k_on without k_off. So the partition needs exactly two measurability boundaries, i and
k_off, and k_on and N_ch stay as thin interior lines that subdivide the green region without
partitioning it.

DOES THE DRAWING AGREE WITH THE CELLS? Every earlier revision drew fitted lines over measured cells
and never asked. Classified twice, once from each cell's own measurements and once from the four
fitted lines, 78 of the 80 agree. Both failures are outside a fit's own support: N_ch 5 for the i
boundary, which is fitted from 10, and N_ch 200 for the calibration boundary, which is fitted to 100.
The chunk prints the confusion table on every render.

WHAT IS MEASURED AND WHAT IS NOT. The cells stop at N_ch 1e4 and the panel runs to 1.2e8, so 56% of
the width is fitted power law carried past its own support, the whole oocyte box included. Three
marks say so, none of them a colour: the fills are veiled with sparse white stripes past the last
simulated column, each boundary is solid only over the N_ch span where its own crossings were found
and dashed outside it, and the crossings are drawn as open dots. Per-boundary support: i 10-1e4,
k_off 10-1e3, k_on 5-1e4, N_ch 5-1e4, IR calibration 5-100, ILSE calibration 10-1e4. The last is
fitted over a decade and drawn over seven, which is why it thins where it is extended.

THREE CLASSES OF MARK, three ways of setting them, so that no two kinds of statement look alike:

    dark on plain white       a PARAMETER      i, k_off, k_on, N_ch     names a thin black line
    white barred in dark      an ALGORITHM     ILSE, IR                  names a white cased line
    white on solid            a PREPARATION    the three footprints     names a real experiment
    italic grey on white      an editorial note                         the key's own asides

CALIBRATED, NOT "HONEST", as of 2026-07-31. The property is that the reported Fisher covariance
equals the actual sampling covariance, which is what calibration names and what the rest of the
manuscript already calls it. "Honest" is also taken: an honest confidence interval is one whose
coverage holds uniformly over a class of parameters, a stronger and different property.

<!-- DEPENDS ON THE `regions` CHUNK, which is where PREP, POLY and POLYC are built. If that chunk is
     ever retired, the preparation table has to move up into `geometry` first.
     NUDGE moves two footprint EDGES off the simulated columns for legibility and is OFF: whole
     cell's 20 and 1e4 are DERIVED from Imin and Imax, so the drawn box would stop stating them. The
     collision it was for is solved instead by the white ring on the cells, which costs no number.
     HATCH_BAD writes the sibling in which the region where the likelihood's interval is not
     calibrated is hatched OVER the yield colour instead of replaced, so that corner keeps saying
     that the amplitude is recoverable there and only the interval is wrong. The purple override is
     the figure, decided 2026-07-31; the sibling is kept because the choice is still live.
     CHOICE CLOSED 2026-08-26: the manuscript's Figure 7 float now includes Figure_7_hatch, so THE
     SIBLING IS THE BODY FIGURE and the purple one is the alternative. Both are still written on
     every knit and nothing here changes. Why: the key's own row already says "use MacroIR, but not
     its reported interval", which is the hatch's claim and not the purple one's; Figure 6 resolves
     that same corner at about a factor of 1.5, correctable, while the purple fill erases what is
     recoverable there; and the corner is where the recursion buys the most precision on the rates
     (median ratio 1.64), so a fill that reads as "nothing to see here" is backwards. The two
     textures do not collide: the veil is white, wide and on the right-hand columns, the hatch is
     black, tight and in the left-hand corner, and no cell carries both. Reverting is one token in
     the manuscript's includegraphics plus "violet" in the caption. -->

```{r map, fig.width = 7, fig.height = 5.9}
PT <- function(pt) pt * 0.3527
TH <- log10(2)

ir   <- IR %>% distinct(what, par, nch, noise, D, .keep_all = TRUE)
irse <- ir %>% filter(what == "var")  %>% transmute(par, nch, noise, D, se = sqrt(pmax(v, 0)))
irdi <- ir %>% filter(what == "dist") %>% transmute(par, nch, noise, D, d  = abs(log(v)))
lsed <- ILSE %>% filter(what == "dist", par == "koff") %>% distinct(nch, noise, D, .keep_all = TRUE) %>%
        transmute(nch, noise, D, l = abs(log(v)))

CELL <- irse %>% filter(D == DREF) %>% group_by(nch, noise) %>%
  summarise(A = sum(se < TH), .groups = "drop") %>%
  left_join(irdi %>% filter(D == DREF) %>% inner_join(irse %>% filter(D == DREF),
              by = c("par","nch","noise")) %>% group_by(nch, noise) %>%
              summarise(irbad = if (!any(se < TH)) 0 else as.numeric(!all(d[se < TH] < log(1.15))),
                        .groups = "drop"), by = c("nch","noise")) %>%
  left_join(lsed %>% filter(D == DREF) %>% select(nch, noise, l), by = c("nch","noise"))

# a boundary = the noise at which a quantity crosses a level, per N_ch, then a log-log fit.
# lo/hi are the N_ch span over which crossings were actually FOUND; from here on that is carried
# through to the drawing instead of being discarded.
fitline <- function(d, col, lev, rising) {
  cr <- d %>% group_by(nch) %>%
    summarise(c = cross1(log10(noise), .data[[col]], lev, rising), .groups = "drop") %>%
    filter(is.finite(c)) %>% mutate(lx = log10(nch))
  if (nrow(cr) < 2) return(NULL)
  list(co = coef(lm(c ~ lx, cr)), lo = min(cr$lx), hi = max(cr$lx), pts = cr)
}
# ONE BOUNDARY PER PARAMETER, not one per count level: the boundaries are not nested (slopes +1.1 to
# +2.2), so they cross, and counting how many lie above a point is correct whatever they do.
PB <- lapply(c("i","koff","kon","Nch"), function(p)
  fitline(irse %>% filter(D == DREF, par == p) %>% rename(v = se), "v", TH, TRUE))
names(PB) <- c("i","koff","kon","Nch")
BH  <- fitline(CELL %>% filter(!is.na(irbad)), "irbad", 0.5, FALSE)   # below it IR is dishonest
BL  <- fitline(CELL %>% filter(!is.na(l)), "l", log(1.15), FALSE)     # above it least squares is honest
for (n in c("i","koff","kon","Nch")) cat(sprintf("SE(%-4s): slope %+.2f, %d crossings, N_ch %g-%g\n",
  n, PB[[n]]$co[2], nrow(PB[[n]]$pts), 10^PB[[n]]$lo, 10^PB[[n]]$hi))
for (n in c("BH","BL")) cat(sprintf("%-4s     slope %+.2f, %d crossings, N_ch %g-%g\n",
  n, get(n)$co[2], nrow(get(n)$pts), 10^get(n)$lo, 10^get(n)$hi))

XQ   <- c(log10(6), log10(1.2e8))
ev2  <- function(f, x) if (is.null(f)) rep(-Inf, length(x)) else f$co[1] + f$co[2]*x
# The veil does not start ON the last simulated column -- there ARE simulations there, and a rule
# drawn through that column reads as if it were excluded. It starts half a grid step past it, and
# the grid step that matters is NOT the 1-2-5 spacing of the columns: the 1-2-5 columns (20, 50,
# 200, 500, 2000, 5000) all stop at noise 10, so none of them can produce a crossing at the noise
# where these boundaries live. The columns that carry the sweep upward are 10, 100, 1000, 1e4 --
# one per decade, each reaching a decade higher in noise than the last (1e4, 1e5, 1e6, 1e7) -- so
# for every boundary drawn here the step is a DECADE, the column after 1e4 would have been 1e5, and
# the veil begins at 1e4.5 = 3.16e4. Taken from the data rather than written in, so it follows the
# sweep if the sweep is extended.
DEEP  <- (CELL %>% group_by(nch) %>% summarise(hi = max(noise), .groups = "drop") %>%
          filter(hi > 10))$nch
XLAST <- log10(max(CELL$nch))                       # last simulated column
XSTEP <- median(diff(log10(sort(DEEP))))            # spacing of the columns that reach high noise
XSIM  <- XLAST + XSTEP / 2                          # where the veil begins
cat(sprintf("\ncells: N_ch %g-%g, noise label %g-%g; panel N_ch %g-%g -> %.0f%% of the width is extrapolated\n",
  min(CELL$nch), max(CELL$nch), min(CELL$noise), max(CELL$noise), 10^XQ[1], 10^XQ[2],
  100 * (XQ[2] - XLAST) / diff(XQ)))
cat(sprintf("columns reaching past noise 10: %s -> step %.2f decades; veil edge at N_ch %.3g\n",
  paste(sort(DEEP), collapse = ", "), XSTEP, 10^XSIM))

PAL <- c(amp = "#238B45", rate = "#A1D99B", lse = "#FD8D3C",
         none = "#F7F7F7", bad = "#9E9AC8")

# ---- exact regions -----------------------------------------------------------------------------
# Sutherland-Hodgman against a half-plane f <= 0. Every boundary is y = a + b x, so the regions are
# the cells of a line arrangement and can be written down instead of sampled.
clip1 <- function(P, f) {
  if (!nrow(P)) return(P)
  s <- f(P$x, P$y); n <- nrow(P); out <- list()
  for (i in seq_len(n)) {
    j <- if (i == n) 1L else i + 1L
    if (s[i] <= 0) out[[length(out) + 1]] <- P[i, ]
    if (s[i] * s[j] < 0) {
      t <- s[i] / (s[i] - s[j])
      out[[length(out) + 1]] <- data.frame(x = P$x[i] + t * (P$x[j] - P$x[i]),
                                           y = P$y[i] + t * (P$y[j] - P$y[i]))
    }
  }
  if (!length(out)) P[0, ] else do.call(rbind, out)
}
area <- function(P) { n <- nrow(P); if (n < 3) return(0)
  j <- c(2:n, 1); abs(sum(P$x * P$y[j] - P$x[j] * P$y)) / 2 }
RECT  <- data.frame(x = c(XQ[1], XQ[2], XQ[2], XQ[1]), y = c(YLIM[1], YLIM[1], YLIM[2], YLIM[2]))
below <- function(f) function(x, y) y - ev2(f, x)          # <= 0 means below the boundary
LN    <- list(i = PB$i, Nch = PB$Nch, koff = PB$koff, kon = PB$kon, h = BH, l = BL)

# s[k] = 1 -> the region lies BELOW line k, and below a measurability boundary is the side on which
# that parameter IS measurable. The rates survive exactly when k_off does (no cell carries k_on
# without it) and the amplitude survives exactly when i does (none carries N_ch without it), so the
# whole partition is s[1] and s[3]. k_on and s[2] play no part in it.
regions <- function(hatch_bad) do.call(rbind, lapply(0:(2^length(LN) - 1), function(m) {
  s <- as.integer(intToBits(m))[seq_along(LN)]
  lseok <- s[6] == 0
  yld <- if (s[3] == 0) "none" else if (s[1] == 1) "amp" else if (lseok) "lse" else "rate"
  col <- if (s[5] == 1 && !hatch_bad) "bad" else yld
  P <- RECT
  for (k in seq_along(LN)) P <- clip1(P, local({ g <- below(LN[[k]]); sk <- s[k]
    if (sk == 1) g else function(x, y) -g(x, y) }))
  if (area(P) < 1e-8) return(NULL)
  transform(P, grp = m, fill = PAL[[col]])
}))

# the two textures. Both are line screens, so neither costs a colour; they differ in ink, angle and
# pitch, and the only place they could meet is the bottom of the last simulated column.
lines_in <- function(px, py, angle_deg, spacing) {
  th <- angle_deg * pi / 180
  rx <-  cos(th) * px + sin(th) * py; ry <- -sin(th) * px + cos(th) * py
  n <- length(px); out <- list()
  for (L in seq(min(ry), max(ry), by = spacing)) {
    xs <- numeric(0)
    for (i in seq_len(n)) { j <- if (i == n) 1L else i + 1L
      if ((ry[i] - L) * (ry[j] - L) < 0) {
        t <- (L - ry[i]) / (ry[j] - ry[i]); xs <- c(xs, rx[i] + t * (rx[j] - rx[i])) } }
    xs <- sort(xs)
    if (length(xs) >= 2) for (k in seq(1, length(xs) - 1, by = 2))
      out[[length(out) + 1]] <- data.frame(
        x = cos(th)*xs[k] - sin(th)*L, y = sin(th)*xs[k] + cos(th)*L,
        xend = cos(th)*xs[k+1] - sin(th)*L, yend = sin(th)*xs[k+1] + cos(th)*L)
  }
  if (!length(out)) NULL else do.call(rbind, out)
}
BADP  <- clip1(RECT, below(BH))                                    # the dishonest corner
BADH  <- lines_in(BADP$x, BADP$y, 45, 0.155)
EXTRA <- lines_in(c(XSIM, XQ[2], XQ[2], XSIM), c(YLIM[1], YLIM[1], YLIM[2], YLIM[2]), -45, 0.78)

YD <- function(v) log10(v * 10)          # dimensionless noise -> the data's own units
# panel is about 6.30 x 5.25 in inside a 7.0 x 5.9 in device
ASPECT <- (diff(XQ) / diff(YLIM)) * (5.25 / 6.30)
# TWO CLASSES OF INLINE LABEL, drawn differently because they name two different kinds of statement.
# A PARAMETER label names a thin black line and is set the same way: dark text on plain white, no
# edge. An ALGORITHM label names a honesty boundary, which is drawn white cased in dark -- a double
# line -- so the label repeats that: a white bar with a dark edge around it, bold. The label then
# looks like a piece of the line it belongs to, and no reader has to be told which is which.
# (label.size is silently IGNORED by annotate("label") in ggplot2 >= 3.5 -- it warns and applies the
# 0.25 default -- so the edge has to be set through linewidth, which is why AK's labels all carried
# a faint box nobody asked for.)
online <- function(f, x, txt, dy = 0.22, pt = 8) {
  lx <- log10(x)
  annotate("label", x = lx, y = ev2(f, lx) + dy, label = txt, size = PT(pt), colour = "grey5",
           angle = atan(f$co[2] * ASPECT) * 180 / pi, hjust = .5, parse = TRUE,
           fill = "white", alpha = .85, linewidth = 0, label.padding = unit(.9, "pt"))
}
onalgo <- function(f, x, txt, dy = 0.22, pt = 8) {
  lx <- log10(x)
  annotate("label", x = lx, y = ev2(f, lx) + dy, label = txt, size = PT(pt), colour = "grey20",
           angle = atan(f$co[2] * ASPECT) * 180 / pi, hjust = .5, fontface = "bold",
           fill = "white", linewidth = .45, label.r = unit(0, "pt"),
           label.padding = unit(1.5, "pt"))
}

# ---- footprint edges moved off the simulated columns --------------------------------------------
# The two small footprints ended exactly on N_ch 10 and 20, which are also two of the columns the
# sweep was run at, so their vertical edges ran through a line of cells and the cells read as beads
# threaded on the box. Moved to 8 and 18. READ BEFORE KEEPING THIS: whole cell's edges are DERIVED
# in figure_7.Rmd, not chosen -- 20 = 100 pA / (10 pA x 0.5) and 1e4 = 5 nA / (1 pA x 0.5), so they
# encode Imin and Imax rather than a guess -- and nudging them means the drawn box no longer states
# those two currents. The right edge costs a second thing: at 1e4 the footprint ended exactly where
# the simulations end, so whole cell was entirely covered by measured cells; at 2e4 it crosses into
# the veiled half. Set NUDGE <- FALSE to get the derived footprints back. The legibility problem it
# solves can also be solved by giving the cells a white halo, which costs no number at all.
NUDGE <- FALSE
if (NUDGE) {
  NEW <- list("excised outside-out patch" = c(8, 300),
              "whole cell"                = c(18, 2e4),
              "oocyte, two-electrode"     = c(1e5, 5e7))
  POLYC <- do.call(rbind, lapply(seq_len(nrow(PREP)), function(k) {
    p <- PREP[k, ]; e <- NEW[[p$name]]
    xs <- seq(log10(e[1]), log10(e[2]), length.out = 160); N <- 10^xs
    itop <- pmax(I_LO, p$Imin / (N * POPEN))   # few channels have to be big ones
    ibot <- pmin(I_HI, p$Imax / (N * POPEN))   # many channels have to be small ones
    data.frame(name = p$name, x = c(xs, rev(xs)),
               y = pmin(pmax(c(log10(lbl(p$S_hi, p$tau_lo, itop)),
                               rev(log10(lbl(p$S_lo, p$tau_hi, ibot)))), YLIM[1]), YLIM[2]))
  }))
  cat("\nfootprint edges drawn at:", paste(sapply(names(NEW), function(n)
      sprintf("%s %g-%g", n, NEW[[n]][1], NEW[[n]][2])), collapse = "; "), "\n")
}

# LINE CODING. White over a dark casing = a statement about an ALGORITHM's error bar; plain black =
# a statement about whether a PARAMETER can be measured. Solid over the N_ch span where that
# boundary's own crossings were measured, dashed where the fit is carried beyond them.
seg3 <- function(f, cap = NULL) {
  lo <- max(f$lo, XQ[1]); hi <- min(f$hi, XQ[2])
  y <- function(d) { d$y <- ev2(f, d$x); if (!is.null(cap)) d$y <- pmin(d$y, ev2(cap, d$x)); d }
  out <- rbind(if (lo > XQ[1] + 1e-9) data.frame(x = seq(XQ[1], lo, length.out = 60), s = "L") else NULL,
               if (hi < XQ[2] - 1e-9) data.frame(x = seq(hi, XQ[2], length.out = 200), s = "R") else NULL)
  list(ins = y(data.frame(x = seq(lo, hi, length.out = 200))),
       out = if (is.null(out)) NULL else y(out))
}
measur <- function(f, cap = NULL) { p <- seg3(f, cap)
  list(geom_line(data = p$ins, aes(x, y), colour = "grey5", linewidth = .45),
       if (!is.null(p$out)) geom_line(data = p$out, aes(x, y, group = s), colour = "grey5",
                                      linewidth = .45, linetype = "22") else NULL) }
honesty <- function(f) { p <- seg3(f)
  list(geom_line(data = p$ins, aes(x, y), colour = "grey25", linewidth = 1.6),
       geom_line(data = p$ins, aes(x, y), colour = "white",  linewidth = 1.05),
       # thinner outside the fit: the IR-honesty line is fitted over one decade and drawn over
       # seven, so at full weight it is the loudest mark in the panel and the least supported
       if (!is.null(p$out)) list(
         geom_line(data = p$out, aes(x, y, group = s), colour = "grey25", linewidth = 1.0,
                   linetype = "42"),
         geom_line(data = p$out, aes(x, y, group = s), colour = "white", linewidth = .6,
                   linetype = "42")) else NULL) }
CROSS <- do.call(rbind, lapply(names(PB), function(p)
  setNames(transform(PB[[p]]$pts[, c("lx","c")], par = p), c("x","y","par"))))

# where each preparation's tab hangs: an N_ch chosen so the three do not collide, and the y taken
# from that footprint's OWN top edge (POLY is built top-run first, then the floor reversed), so the
# tab is attached to the box rather than floating inside it.
# TEVC = two-electrode voltage clamp, which is what PREP's "oocyte, two-electrode" row is; it is
# standard in the field's literature, so the acronym earns its place on a tab this size, but it has
# to be spelled out in the caption. The patch row is named for its ORIENTATION rather than for being
# excised, because the whole map assumes an agonist concentration jump and only an outside-out patch
# puts the extracellular face in the bath: "excised patch" would take in inside-out, which cannot run
# the experiment being mapped. The tabs hang where the name still fits inside its own footprint --
# N_ch 25 for the patch, whose top edge is flat from N_ch 20 on, so the height does not move. That
# one is set over two lines: on one it is 1.12 decades wide against a 1.48-decade footprint, so it
# arrives within a hair of the box's own right edge whatever N_ch it hangs from.
# hj = 0 hangs the tab from the top edge running rightwards, hj = 1 hangs it from the top RIGHT
# corner running leftwards. The patch tab takes the corner: at 1.12 decades against a 1.48-decade
# footprint it fits on one line only if it is anchored to an edge, and the corner is the one part of
# that box with no boundary line crossing it -- hung on the left it landed among k_on, N_ch and the
# IR label. It still covers the three cells on the noise = 0.2 row inside its own span; that is the
# cheapest thing available to cover there, and cheaper than covering a line.
PTAB <- do.call(rbind, lapply(list(c("excised outside-out patch", "outside-out patch", "3.0e2", "1"),
                                   c("whole cell",                "whole cell",        "3.5e2", "0"),
                                   c("oocyte, two-electrode",     "oocyte, TEVC",      "1.15e5","0")),
  function(z) { d <- POLYC[POLYC$name == z[1], ]; top <- d[seq_len(nrow(d) / 2), ]
    data.frame(lab = z[2], hj = as.numeric(z[4]), lx = log10(as.numeric(z[3])),
               ly = approx(top$x, top$y, xout = log10(as.numeric(z[3])))$y - .04) }))

# Does the DRAWN partition agree with the cells it was fitted from? Every earlier revision drew
# fitted lines over measured cells and never asked. Classify each cell twice, once from its own
# measurements and once from the four fitted lines, and count the disagreements.
CHK <- CELL %>% mutate(lx = log10(nch), ly = log10(noise)) %>%
  left_join(irse %>% filter(D == DREF, par %in% c("i","koff")) %>%
              mutate(ok = se < TH) %>% select(par, nch, noise, ok) %>%
              tidyr::pivot_wider(names_from = par, values_from = ok), by = c("nch","noise")) %>%
  mutate(cell = case_when(irbad == 1 ~ "bad", !koff ~ "none", i ~ "amp",
                          !is.na(l) & l < log(1.15) ~ "lse", TRUE ~ "rate"),
         drawn = case_when(ly < ev2(BH, lx) ~ "bad", ly > ev2(PB$koff, lx) ~ "none",
                           ly < ev2(PB$i, lx) ~ "amp",
                           ly > ev2(BL, lx) ~ "lse", TRUE ~ "rate"))
cat(sprintf("\npartition check: %d of %d cells disagree with the fitted lines\n",
            sum(CHK$cell != CHK$drawn), nrow(CHK)))
print(table(cell = CHK$cell, drawn = CHK$drawn))
print(as.data.frame(CHK %>% filter(cell != drawn) %>% arrange(nch, noise) %>%
                    select(nch, noise, cell, drawn)))

for (HATCH_BAD in c(TRUE, FALSE)) {

# ---- the key. One swatch column, because the tool axis is no longer a second dimension of the
# fill: it is the ORDER of the rows. Laid out on an explicit cursor in text lines rather than on a
# row count, since the blocks are 1, 2 and 5 lines tall and a uniform row pitch either overlaps them
# or wastes half the box. LH is one 7 pt line measured in y units: the panel is 5.25 in over
# diff(YLIM) decades, so a decade is 0.536 in and 9 pt of leading is 0.233 of one.
LH  <- 0.235
KX  <- c(log10(1.05e5), XQ[2] - .05)
kcx <- KX[1] + diff(KX) * .075; ktx <- KX[1] + diff(KX) * .165
kw  <- diff(KX) * .048; kh <- LH * .62
kx0 <- kcx - kw; kx1 <- kcx + kw + diff(KX) * .06
Y <- local({ y <- 0; f <- function(n, gap = 0) { y <<- y - gap - LH * n / 2
                                                 v <- y; y <<- y - LH * n / 2; v }
             list(next_ = f, used = function() -y) })
y_title <- Y$next_(1, .30)
y_rows  <- c(Y$next_(1, .18), Y$next_(1, .10), Y$next_(1, .10), Y$next_(1, .10), Y$next_(1, .10))
y_note  <- Y$next_(1, .16)
y_line1 <- Y$next_(2, .18)
y_line2 <- Y$next_(2, .10)
y_gloss <- Y$next_(1, .16)
# anchored from the FLOOR, not from a chosen top: a box sized from the top runs off the bottom of
# the panel and ggplot then drops the whole rect, silently, leaving the key with no background.
KY  <- c(YLIM[1] + .02, 0); KY[2] <- KY[1] + Y$used() + .12
kr  <- function(v) KY[2] + v
# ROW ORDER IS THE PANEL'S OWN VERTICAL ORDER, top to bottom: the key is a section through the map,
# its first row the region at the top of the panel and its last the one at the bottom. Listing them
# in the order a column is crossed going UP matches the sequence but reverses the direction, so the
# reader has to flip between key and map; this way the two are read the same way round.
FILLS <- c(PAL[["none"]], PAL[["lse"]], PAL[["rate"]], PAL[["amp"]],
           if (HATCH_BAD) PAL[["amp"]] else PAL[["bad"]])
KEY <- list(
  # opaque, not alpha .95: at 95% the white veil stripes behind still print as faint bands
  annotate("rect", xmin = KX[1], xmax = KX[2], ymin = KY[1], ymax = KY[2],
           fill = "white", colour = "grey70", linewidth = .25),
  annotate("text", x = kx0, y = kr(y_title), hjust = 0, size = PT(7.5), colour = "grey30",
                      # No second person. Published prose in the manuscript uses "you" zero times (the two hits
           # in 00_abstract.tex are both inside % comments), and no other caption in the set has it,
           # so a key written at the reader would be the one place in the paper that does.
           fontface = "italic", label = "what to use, and what it recovers"),
  annotate("rect", xmin = kcx - kw, xmax = kcx + kw, ymin = kr(y_rows) - kh,
           ymax = kr(y_rows) + kh, fill = FILLS, colour = "grey70", linewidth = .2),
  # the dishonest corner keeps its hatch in the key when it has one on the map
  if (HATCH_BAD) geom_segment(data = lines_in(
      c(kcx-kw, kcx+kw, kcx+kw, kcx-kw),
      c(kr(y_rows[5])-kh, kr(y_rows[5])-kh, kr(y_rows[5])+kh, kr(y_rows[5])+kh), 45, .075),
      aes(x = x, y = y, xend = xend, yend = yend), colour = "grey15", linewidth = .22,
      inherit.aes = FALSE) else NULL,
  # PRESCRIPTIVE, not descriptive: this is a usage map, so the rows say what to do and the thing
  # that varies is the first word. Three of the five reading "use MacroIR" is the argument and not a
  # collapse of it -- at a glance the likelihood is needed almost everywhere. Two of the rows were
  # arrived at against my own first objection and are worth recording:
  #   AMBER. "use ILSE" is right, not merely permitted. Both methods recover the rates there and both
  #   are calibrated, and the widths are the same: over the 32 cells where both were run the ratio
  #   of the corrected standard error, ILSE over IR, has median 1.04 on k_off and 1.01 on k_on, and
  #   in two of the six rate entries of the three amber cells ILSE is the TIGHTER of the two. Same
  #   answer, same interval, cheaper method. This is Figure 4's corrected covariance and the
  #   information budget: the rates live in the shape of the deterministic transient.
  #   VIOLET. "use MacroIR but do not trust the calibration" tells the reader to use a method and
  #   distrust it without saying what to do instead. What is wrong there is the REPORTED interval,
  #   so that is what the row names. The estimate itself is not what the diagnostic condemns.
  # "often all four": 11 of the 63 cells in the first region carry only k_off and the conductance,
  # so naming the full four would be false over a sixth of it. The two thin lines inside the region
  # are where the other two go, and the caption carries the subset table.
  annotate("text", x = ktx, y = kr(y_rows), hjust = 0, size = PT(7.5), colour = "grey25",
           parse = TRUE,
           # Two of these were written for the old row order and stopped parsing when it changed.
           # "calibrated here TOO" pointed at a MacroIR row that used to sit above it and now sits
           # below, so the antecedent is gone and the row is made self-contained. In exchange the
           # new order earns an anaphora the old one could not have: the conductance row now sits
           # directly under the rates row, so "i as well" has something to be as well AS, which is
           # how the rows were first drafted.
           label = c("'nothing measurable'",
                     "'use ILSE:  calibrated here, and cheaper'",
                     "'use MacroIR:  the rates, calibrated'",
                     "'use MacroIR:'~italic(i)~'as well, often all four'",
                     "'use MacroIR, but not its reported interval'")),
  # WHY the first row belongs to one method only, stated as a fact about the method and nothing
  # more. It used to carry a second clause, "so in the top row it is not a candidate at all", which
  # was cut: "top row" is the key's first row but the MAP's lowest region, and read as the map it
  # says the opposite of the amber region, where least squares is an honest option. The reader who
  # needs the inference has the row above; the caption spells it out.
  annotate("text", x = kx0, y = kr(y_note), hjust = 0, size = PT(7), colour = "grey30",
           fontface = "italic",
           label = "ILSE holds i fixed and cannot separate N_ch from it"),
  # the two line classes, explained once here instead of twice on the map
  annotate("rect", xmin = kx0 - .02, xmax = kx1 + .02, ymin = kr(y_line1) - kh*.9,
           ymax = kr(y_line1) + kh*.9, fill = PAL[["rate"]], colour = NA),
  annotate("segment", x = kx0, xend = kx1, y = kr(y_line1), yend = kr(y_line1),
           colour = "grey25", linewidth = 1.6),
  annotate("segment", x = kx0, xend = kx1, y = kr(y_line1), yend = kr(y_line1),
           colour = "white", linewidth = 1.05),
  annotate("segment", x = kx0, xend = (kx0 + kx1)/2, y = kr(y_line2), yend = kr(y_line2),
           colour = "grey5", linewidth = .45),
  annotate("segment", x = (kx0 + kx1)/2, xend = kx1, y = kr(y_line2), yend = kr(y_line2),
           colour = "grey5", linewidth = .45, linetype = "22"),
  annotate("point", x = (kx0 + kx1)/2, y = kr(y_line2), size = 1.1, shape = 21,
           fill = "white", colour = "grey5", stroke = .35),
  annotate("text", x = kx1 + .10, y = c(kr(y_line1), kr(y_line2)), hjust = 0, size = PT(7),
           colour = "grey25", lineheight = .95,
           label = c("that interval becomes calibrated\n(information distortion = 1.15)",
                     "that parameter stops being measurable\n(standard error = a factor of 2)")),
  annotate("text", x = kx0, y = kr(y_gloss), hjust = 0, size = PT(7), colour = "grey30",
           label = "IR = MacroIR;  ILSE = classical least squares."))

gq <- ggplot() +
  geom_polygon(data = regions(HATCH_BAD), aes(x, y, group = grp, fill = fill, colour = fill),
               linewidth = .12) +
  scale_fill_identity() + scale_colour_identity() +
  (if (HATCH_BAD) geom_segment(data = BADH, aes(x = x, y = y, xend = xend, yend = yend),
                               colour = "grey15", linewidth = .28, inherit.aes = FALSE) else NULL) +
  # past the last simulated column there is nothing but the fits, so the fills are veiled
  geom_segment(data = EXTRA, aes(x = x, y = y, xend = xend, yend = yend),
               colour = "white", linewidth = .55, alpha = .5, inherit.aes = FALSE) +
  measur(PB$i) + measur(PB$koff) + measur(PB$kon) + measur(PB$Nch, cap = PB$i) +
  honesty(BL) + honesty(BH) +
  geom_point(data = CROSS, aes(x, y), size = 1.1, shape = 21, fill = "white",
             colour = "grey5", stroke = .35) +
  # THE THREE PREPARATIONS ARE A DIFFERENT KIND OF OBJECT from everything else in the panel: they
  # are where real experiments sit, not something the sweep measured. In AK they were drawn in the
  # same language as the diagnostics -- thin dark outline, white label box, same size -- so "patch"
  # and "k_off" read as the same class of thing. Here they get their own: a heavier outline, and a
  # solid tab hung from the box's own top edge with the name reversed out of it. Reversed on solid
  # is the exact inverse of the boundary labels (dark on white), which is the cheapest way to say
  # "different layer" without spending a colour or a line style that already means something.
  geom_polygon(data = POLYC, aes(x, y, group = name), fill = NA, colour = "grey5",
               linewidth = .8, inherit.aes = FALSE) +
  lapply(c(0, 1), function(h) geom_label(data = PTAB[PTAB$hj == h, ],
             aes(x = lx, y = ly, label = lab), size = PT(7.5), colour = "white",
             fontface = "bold", inherit.aes = FALSE, fill = "grey5", linewidth = 0,
             label.padding = unit(1.8, "pt"), hjust = h, vjust = 1)) +
  # a thin white ring, so a cell still reads as a cell where a footprint edge runs through its
  # column. That collision is what the footprint edges were nudged off, and the ring solves it
  # without moving a number: two of those edges are DERIVED (whole cell 20 = 100 pA / (10 pA x 0.5)
  # and 1e4 = 5 nA / (1 pA x 0.5)), so a caption cannot state them and the drawing show something
  # else. NUDGE is kept for comparison but is off.
  geom_point(data = CELL %>% mutate(x = log10(nch), y = log10(noise)), aes(x, y),
             size = .95, shape = 21, fill = "grey5", colour = "white", stroke = .28) +
  # THE EDGE OF THE VEIL, drawn last of the map marks so the preparation boxes cannot bury it: the
  # whole-cell footprint happens to end at exactly N_ch 1e4, so its heavy edge was sitting on top of
  # this line for the bottom third of the panel. Pure white, no casing, matching the veil stripes it
  # bounds -- the honesty lines are white CASED IN DARK, so the two never read as the same mark.
  # The caption hangs off the line itself rather than floating in the middle of the veiled half.
  annotate("segment", x = XSIM, xend = XSIM, y = YLIM[1], yend = YLIM[2],
           colour = "white", linewidth = 1.0, linetype = "42") +
  annotate("segment", x = XSIM, xend = XSIM + .60, y = YLIM[2] - .42, yend = YLIM[2] - .42,
           colour = "white", linewidth = .9) +
  # italic grey on white, which is this figure's voice for an editorial note (the key's own notes are
  # set the same way): not bold-on-solid, which now means a preparation, and not a plain white box,
  # which now means a parameter. Reversed white type failed anyway -- half the sentence lands on the
  # mid green and stops being legible.
  annotate("label", x = XSIM + .55, y = YLIM[2] - .42, hjust = 0, size = PT(7.5), colour = "grey20",
           fontface = "italic", label = "no simulations beyond here", fill = "white",
           linewidth = 0, label.padding = unit(1.4, "pt")) +
  # the inline labels last of all, so nothing crosses them
  online(PB$i,    2.2e2, "italic(i)",  .20) +
  online(PB$Nch,  1.5e3, "N[ch]",    -.30) +
  online(PB$koff, 3.0e1, "k[off]",     .20) +
  online(PB$kon,  1.2e2, "k[on]",      .20) +
  onalgo(BL, 6.0e3, "ILSE", .30, 8) +
  onalgo(BH, 45,    "IR",  .34, 8) +
  KEY +
  scale_x_continuous(breaks = log10(10^(1:8)),
                     labels = lapply(1:8, function(e) if (e == 1) "10" else bquote(10^.(e))),
                     limits = XQ, expand = c(0, 0)) +
  scale_y_continuous(breaks = log10(10^(-2:7)),
                     labels = lapply((-2:7) - 1, function(e)
                       if (e >= -2 && e <= 1) format(10^e, scientific = FALSE, trim = TRUE)
                       else bquote(10^.(e))),
                     limits = YLIM, expand = c(0, 0)) +
  labs(x = expression(N[ch]), y = "dimensionless instrumental noise") +
  theme_bw(base_size = 9, base_family = "Helvetica") +
  theme(panel.grid = element_blank(),
        axis.title = element_text(size = 9.5), axis.text = element_text(size = 8),
        plot.margin = margin(4, 5, 2, 4))

# the hatch sibling is the BODY FIGURE since 2026-08-26; the purple override is the alternative
f <- if (HATCH_BAD) "Figure_7_hatch" else "Figure_7"
ggsave(paste0(f, ".pdf"), gq, width = 7.0, height = 5.9, device = cairo_pdf)
ggsave(paste0(f, ".png"), gq, width = 7.0, height = 5.9, dpi = 140)
if (!HATCH_BAD) print(gq)
}
```
