---
title: "Figure 2"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

```{r}

filename=c(
  "data/figure_3_nch_10_nsim_1024_macro_IR_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_100_nsim_1024_macro_IR_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_1000_nsim_1024_macro_IR_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_10000_nsim_1024_macro_IR_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_10_nsim_1024_macro_R_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_100_nsim_1024_macro_R_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_1000_nsim_1024_macro_R_noise_0.1_gs_1_battery_pool",
  "data/figure_3_nch_10000_nsim_1024_macro_R_noise_0.1_gs_1_battery_pool"
  
  )

nsamples=rep(1024,length(filename))

```

```{r}

library(dplyr)

d <- lapply(seq_along(filename), function(i) {
  read.csv(paste0(filename[i], ".csv")) %>% 
    mutate(nsamples = nsamples[i])
}) %>% 
  bind_rows()
```

```{r}
d<- d %>%
  select(where(~ !all(is.na(.))))
```

```{r}
d%>%select(component_path)%>%distinct()%>%pull()

```

```{r}
d<-d%>%mutate(component= component_path %>%
           str_replace("^Probit_statistics_", "") )
   
     
```

```{r}
d%>%select(component)%>%distinct()%>%pull()

```



```{r}
d_mean<-
  d%>%filter(!component%in%c("Probit_statistics","count_r_std","count_dlogL"))%>%
  #select(component_path,component,param_name,algorithm,sample_index,calculus,statistic,param_index,param_col,param_name,simulation_algorithm,noise_in_conductance_tau,Num_ch,interval_in_tau,statistic,probit,quantile_level,value,value_row,value_col)%>%
  pivot_wider(names_from = c(probit,quantile_level), values_from = value)
d_mean
```

```{r}
scalar_components<-d_mean%>%filter(statistic%in%c("mean", "value"), is.na(param_col),is.na(param_index) , !is.na(mean_NA))%>%select(component)%>%distinct()%>%pull()
scalar_components

```
```{r}
d_algo<-
  d%>%filter(!component%in%c("Probit_statistics","count_r_std","count_dlogL"))%>%
  #select(component_path,component,param_name,algorithm,sample_index,calculus,statistic,param_index,param_col,param_name,simulation_algorithm,noise_in_conductance_tau,Num_ch,interval_in_tau,statistic,probit,quantile_level,value,value_row,value_col)%>%
  pivot_wider(names_from = c(algorithm,probit,quantile_level), values_from = value)
d_algo

```
```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.10" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Moment_statistics_Sum_r_std", "Moment_statistics_Sum_r2_std")

for (al in c("macro_IR","macro_R"))
for (comp in components_to_plot) {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      component == comp, 
      statistic == "mean", 
      algorithm == al, 
      simulation_algorithm == "uniformization"
    ) %>%
    # Ensure matching by formatting numbers to match your exact vector keys
    mutate(
      key = sprintf("%.2f", interval_in_tau), # e.g., 0.5 becomes "0.50"
      # If your keys are exactly "1.00", "0.5", "0.20", let's normalize them:
      key = case_when(
        interval_in_tau == 1.00 ~ "1.00",
        interval_in_tau == 0.50 ~ "0.5",
        interval_in_tau == 0.20 ~ "0.20",
        interval_in_tau == 0.10 ~ "0.10",
        interval_in_tau == 0.05 ~ "0.05",
        interval_in_tau == 0.02 ~ "0.02",
        interval_in_tau == 0.01 ~ "0.01",
        TRUE ~ as.character(interval_in_tau)
      ),
      n_samples = num_samples[key],
      
      # Pre-calculate scaled values to keep ggplot code clean
      scaled_mean = mean_NA / n_samples,
      scaled_ymin = quantile_0.025 / n_samples,
      scaled_ymax = quantile_0.975 / n_samples
    )
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = interval_in_tau, color = component)) +
    geom_point(aes(y = scaled_mean,color=factor(nsamples))) +
    geom_errorbar(aes(ymin = scaled_ymin, ymax = scaled_ymax, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
    facet_grid(Num_ch ~ noise_in_conductance_tau, scales = "free") +
    geom_hline(yintercept = ifelse(comp=="Moment_statistics_Sum_r2_std",1,0), color="green")+
    ggtitle(paste0(al," ",comp)) 
  
  print(p)
}


```

 [1] "Moment_statistics_Sum_logL"                                           "Moment_statistics_Sum_elogL"                                          "Moment_statistics_Sum_r_std"                                         
 [4] "Moment_statistics_Sum_r2_std"                                         "Moment_statistics_Sum_trust_coefficient"                              "Moment_statistics_Sum_dlogL"                                         
 [7] "Moment_statistics_Sum_Gaussian_Fisher_Information"                    "Sum_Moment_statistics_r_std_false"                                    "Sum_Moment_statistics_dlogL_true"                                    
[10] "Sum_Moment_statistics_Gaussian_Fisher_Information_false"              "Moment_statistics_evaluation_time"                                    "Likelihood_Numerical_Fisher_Information"                             
[13] "Likelihood_Information_Distortion"                                    "log_Det_Likelihood_Information_Distortion"                            "Eigenvalue_Spectrum_Likelihood_Information_Distortion"               
[16] "Spectrum_Condition_Number_Likelihood_Information_Distortion"          "Max_Eigenvalue_Likelihood_Information_Distortion"                     "Min_Eigenvalue_Likelihood_Information_Distortion"                    
[19] "Mean_Log_Eigenvalue_Likelihood_Information_Distortion"                "Log_Eigenvalue_Variance_Likelihood_Information_Distortion"            "Affine_Invariant_Distance_Likelihood_Information_Distortion"         
[22] "Symmetrized_KL_Distortion_Likelihood_Information_Distortion"          "Likelihood_Gaussian_Fisher_Distortion"                                "log_Det_Likelihood_Gaussian_Fisher_Distortion"                       
[25] "Eigenvalue_Spectrum_Likelihood_Gaussian_Fisher_Distortion"            "Spectrum_Condition_Number_Likelihood_Gaussian_Fisher_Distortion"      "Max_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                
[28] "Min_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                 "Mean_Log_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"            "Log_Eigenvalue_Variance_Likelihood_Gaussian_Fisher_Distortion"       
[31] "Affine_Invariant_Distance_Likelihood_Gaussian_Fisher_Distortion"      "Symmetrized_KL_Distortion_Likelihood_Gaussian_Fisher_Distortion"      "Likelihood_Information_Distortion_Reconstituted"                     
[34] "Likelihood_Sample_Distortion"                                         "log_Det_Likelihood_Sample_Distortion"                                 "Likelihood_Correlation_Distortion"                                   
[37] "log_Det_Likelihood_Correlation_Distortion"                            "Likelihood_Fisher_Covariance"                                         "log_Det_Likelihood_Fisher_Covariance"                                
[40] "Correlation_Of_Likelihood_Fisher_Covariance"                          "Eigenvalue_Spectrum_Likelihood_Fisher_Covariance"                     "Effective_Rank_Likelihood_Fisher_Covariance"                         
[43] "Spectrum_Condition_Number_Likelihood_Fisher_Covariance"               "Null_Space_Projector_Likelihood_Fisher_Covariance"                    "Worst_Subspace_Projector_Likelihood_Fisher_Covariance"               
[46] "Likelihood_Distortion_Corrected_Covariance"                           "log_Det_Likelihood_Distortion_Corrected_Covariance"                   "Correlation_Of_Likelihood_Distortion_Corrected_Covariance"           
[49] "Eigenvalue_Spectrum_Likelihood_Distortion_Corrected_Covariance"       "Effective_Rank_Likelihood_Distortion_Corrected_Covariance"            "Spectrum_Condition_Number_Likelihood_Distortion_Corrected_Covariance"
[52] "Null_Space_Projector_Likelihood_Distortion_Corrected_Covariance"      "Worst_Subspace_Projector_Likelihood_Distortion_Corrected_Covariance"  "Likelihood_Distortion_Induced_Bias"                                  
[55] "Report_integral_logL"                                                 "Report_integral_y_mean"                                               "Report_integral_y_var"                                               
[58] "Report_integral_r_std"                                                "Report_integral_dlogL"              

```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Likelihood_Distortion_Induced_Bias", "Moment_statistics_Sum_dlogL")


for (al in c("macro_IR","macro_R")){
noise_in_conductance_tau_plot =d_mean%>%filter(algorithm == al)%>%select(noise_in_conductance_tau)%>%distinct()%>%pull()
  
for (comp in components_to_plot) 
for (noi in noise_in_conductance_tau_plot)
    {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      component == comp, 
      statistic %in%c( "value","mean"), 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      noise_in_conductance_tau==noi, 
      Num_ch>5)
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x =interval_in_tau, color = factor(nsamples)) )+
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
    #facet_grid(param_name~Num_ch, scales = "free") +
    facet_wrap(param_name~Num_ch, scales = "free", nrow=6) +
    geom_hline(yintercept = 0, color="green")+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  print(p)
}
}

```



[1] "Moment_statistics_Sum_logL"                                           "Moment_statistics_Sum_elogL"                                          "Moment_statistics_Sum_r_std"                                         
 [4] "Moment_statistics_Sum_r2_std"                                         "Moment_statistics_Sum_trust_coefficient"                              "Moment_statistics_Sum_dlogL"                                         
 [7] "Moment_statistics_Sum_Gaussian_Fisher_Information"                    "Sum_Moment_statistics_r_std_false"                                    "Sum_Moment_statistics_dlogL_true"                                    
[10] "Sum_Moment_statistics_Gaussian_Fisher_Information_false"              "Moment_statistics_evaluation_time"                                    "Likelihood_Numerical_Fisher_Information"                             
[13] "Likelihood_Information_Distortion"                                    "log_Det_Likelihood_Information_Distortion"                            "Eigenvalue_Spectrum_Likelihood_Information_Distortion"               
[16] "Spectrum_Condition_Number_Likelihood_Information_Distortion"          "Max_Eigenvalue_Likelihood_Information_Distortion"                     "Min_Eigenvalue_Likelihood_Information_Distortion"                    
[19] "Mean_Log_Eigenvalue_Likelihood_Information_Distortion"                "Log_Eigenvalue_Variance_Likelihood_Information_Distortion"            "Affine_Invariant_Distance_Likelihood_Information_Distortion"         
[22] "Symmetrized_KL_Distortion_Likelihood_Information_Distortion"          "Likelihood_Gaussian_Fisher_Distortion"                                "log_Det_Likelihood_Gaussian_Fisher_Distortion"                       
[25] "Eigenvalue_Spectrum_Likelihood_Gaussian_Fisher_Distortion"            "Spectrum_Condition_Number_Likelihood_Gaussian_Fisher_Distortion"      "Max_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                
[28] "Min_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                 "Mean_Log_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"            "Log_Eigenvalue_Variance_Likelihood_Gaussian_Fisher_Distortion"       
[31] "Affine_Invariant_Distance_Likelihood_Gaussian_Fisher_Distortion"      "Symmetrized_KL_Distortion_Likelihood_Gaussian_Fisher_Distortion"      "Likelihood_Information_Distortion_Reconstituted"                     
[34] "Likelihood_Sample_Distortion"                                         "log_Det_Likelihood_Sample_Distortion"                                 "Likelihood_Correlation_Distortion"                                   
[37] "log_Det_Likelihood_Correlation_Distortion"                            "Likelihood_Fisher_Covariance"                                         "log_Det_Likelihood_Fisher_Covariance"                                
[40] "Correlation_Of_Likelihood_Fisher_Covariance"                          "Eigenvalue_Spectrum_Likelihood_Fisher_Covariance"                     "Effective_Rank_Likelihood_Fisher_Covariance"                         
[43] "Spectrum_Condition_Number_Likelihood_Fisher_Covariance"               "Null_Space_Projector_Likelihood_Fisher_Covariance"                    "Worst_Subspace_Projector_Likelihood_Fisher_Covariance"               
[46] "Likelihood_Distortion_Corrected_Covariance"                           "log_Det_Likelihood_Distortion_Corrected_Covariance"                   "Correlation_Of_Likelihood_Distortion_Corrected_Covariance"           
[49] "Eigenvalue_Spectrum_Likelihood_Distortion_Corrected_Covariance"       "Effective_Rank_Likelihood_Distortion_Corrected_Covariance"            "Spectrum_Condition_Number_Likelihood_Distortion_Corrected_Covariance"
[52] "Null_Space_Projector_Likelihood_Distortion_Corrected_Covariance"      "Worst_Subspace_Projector_Likelihood_Distortion_Corrected_Covariance"  "Likelihood_Distortion_Induced_Bias"                                  
[55] "Report_integral_logL"                                                 "Report_integral_y_mean"                                               "Report_integral_y_var"                                               
[58] "Report_integral_r_std"                                                "Report_integral_dlogL"              


```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Likelihood_Information_Distortion", "Likelihood_Gaussian_Fisher_Distortion", "Likelihood_Correlation_Distortion", "Likelihood_Sample_Distortion")

noise_in_conductance_tau_plot =unique(d_mean$noise_in_conductance_tau)

for (al in c("macro_IR","macro_R")){
noise_in_conductance_tau_plot =d_mean%>%filter(algorithm == al)%>%select(noise_in_conductance_tau)%>%distinct()%>%pull()
  
for (comp in components_to_plot) 
for (noi in noise_in_conductance_tau_plot)
    {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      Num_ch>5,
      component == comp, 
      statistic == "value", 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      noise_in_conductance_tau==noi, 
      interval_in_tau>0.001,
      param_col==param_index)
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = Num_ch, color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(interval_in_tau~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  p2 <- ggplot(plot_data, aes(x = interval_in_tau , color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(Num_ch~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  
  print(p)
  print(p2)
}
}

```




"Moment_statistics_Sum_logL"                                           "Moment_statistics_Sum_elogL"                                         
 [3] "Moment_statistics_Sum_r_std"                                          "Moment_statistics_Sum_r2_std"                                        
 [5] "Moment_statistics_Sum_trust_coefficient"                              "Moment_statistics_Sum_dlogL"                                         
 [7] "Moment_statistics_Sum_Gaussian_Fisher_Information"                    "Sum_Moment_statistics_r_std_false"                                   
 [9] "Sum_Moment_statistics_dlogL_true"                                     "Sum_Moment_statistics_Gaussian_Fisher_Information_false"             
[11] "Moment_statistics_evaluation_time"                                    "Likelihood_Numerical_Fisher_Information"                             
[13] "Likelihood_Information_Distortion"                                    "log_Det_Likelihood_Information_Distortion"                           
[15] "Eigenvalue_Spectrum_Likelihood_Information_Distortion"                "Spectrum_Condition_Number_Likelihood_Information_Distortion"         
[17] "Max_Eigenvalue_Likelihood_Information_Distortion"                     "Min_Eigenvalue_Likelihood_Information_Distortion"                    
[19] "Mean_Log_Eigenvalue_Likelihood_Information_Distortion"                "Log_Eigenvalue_Variance_Likelihood_Information_Distortion"           
[21] "Affine_Invariant_Distance_Likelihood_Information_Distortion"          "Symmetrized_KL_Distortion_Likelihood_Information_Distortion"         
[23] "Likelihood_Gaussian_Fisher_Distortion"                                "log_Det_Likelihood_Gaussian_Fisher_Distortion"                       
[25] "Eigenvalue_Spectrum_Likelihood_Gaussian_Fisher_Distortion"            "Spectrum_Condition_Number_Likelihood_Gaussian_Fisher_Distortion"     
[27] "Max_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                 "Min_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                
[29] "Mean_Log_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"            "Log_Eigenvalue_Variance_Likelihood_Gaussian_Fisher_Distortion"       
[31] "Affine_Invariant_Distance_Likelihood_Gaussian_Fisher_Distortion"      "Symmetrized_KL_Distortion_Likelihood_Gaussian_Fisher_Distortion"     
[33] "Likelihood_Information_Distortion_Reconstituted"                      "Likelihood_Sample_Distortion"                                        
[35] "log_Det_Likelihood_Sample_Distortion"                                 "Likelihood_Correlation_Distortion"                                   
[37] "log_Det_Likelihood_Correlation_Distortion"                            "Likelihood_Fisher_Covariance"                                        
[39] "log_Det_Likelihood_Fisher_Covariance"                                 "Correlation_Of_Likelihood_Fisher_Covariance"                         
[41] "Eigenvalue_Spectrum_Likelihood_Fisher_Covariance"                     "Effective_Rank_Likelihood_Fisher_Covariance"                         
[43] "Spectrum_Condition_Number_Likelihood_Fisher_Covariance"               "Null_Space_Projector_Likelihood_Fisher_Covariance"                   
[45] "Worst_Subspace_Projector_Likelihood_Fisher_Covariance"                "Likelihood_Distortion_Corrected_Covariance"                          
[47] "log_Det_Likelihood_Distortion_Corrected_Covariance"                   "Correlation_Of_Likelihood_Distortion_Corrected_Covariance"           
[49] "Eigenvalue_Spectrum_Likelihood_Distortion_Corrected_Covariance"       "Effective_Rank_Likelihood_Distortion_Corrected_Covariance"           
[51] "Spectrum_Condition_Number_Likelihood_Distortion_Corrected_Covariance" "Null_Space_Projector_Likelihood_Distortion_Corrected_Covariance"     
[53] "Worst_Subspace_Projector_Likelihood_Distortion_Corrected_Covariance"  "Likelihood_Distortion_Induced_Bias"                                  
[55] "Report_integral_logL"                                                 "Report_integral_y_mean"                                              
[57] "Report_integral_y_var"                                                "Report_integral_r_std"                                               
[59] "Report_integral_dlogL"                                               

```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Likelihood_Numerical_Fisher_Information")

noise_in_conductance_tau_plot =unique(d_mean$noise_in_conductance_tau)

for (al in c("macro_IR","macro_R")){
noise_in_conductance_tau_plot =d_mean%>%filter(algorithm == al)%>%select(noise_in_conductance_tau)%>%distinct()%>%pull()
  
for (comp in components_to_plot) 
for (noi in noise_in_conductance_tau_plot)
    {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      Num_ch>5,
      component == comp, 
      statistic == "value", 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      noise_in_conductance_tau==noi, 
      interval_in_tau>0.001,
      param_col==param_index)
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = Num_ch, color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(interval_in_tau~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  p2 <- ggplot(plot_data, aes(x = interval_in_tau , color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(Num_ch~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  
  print(p)
  print(p2)
}
}

```

[1] "Moment_statistics_Sum_logL"                                           "Moment_statistics_Sum_elogL"                                         
 [3] "Moment_statistics_Sum_r_std"                                          "Moment_statistics_Sum_r2_std"                                        
 [5] "Moment_statistics_Sum_trust_coefficient"                              "Moment_statistics_Sum_dlogL"                                         
 [7] "Moment_statistics_Sum_Gaussian_Fisher_Information"                    "Sum_Moment_statistics_r_std_false"                                   
 [9] "Sum_Moment_statistics_dlogL_true"                                     "Sum_Moment_statistics_Gaussian_Fisher_Information_false"             
[11] "Moment_statistics_evaluation_time"                                    "Likelihood_Numerical_Fisher_Information"                             
[13] "Likelihood_Information_Distortion"                                    "log_Det_Likelihood_Information_Distortion"                           
[15] "Eigenvalue_Spectrum_Likelihood_Information_Distortion"                "Spectrum_Condition_Number_Likelihood_Information_Distortion"         
[17] "Max_Eigenvalue_Likelihood_Information_Distortion"                     "Min_Eigenvalue_Likelihood_Information_Distortion"                    
[19] "Mean_Log_Eigenvalue_Likelihood_Information_Distortion"                "Log_Eigenvalue_Variance_Likelihood_Information_Distortion"           
[21] "Affine_Invariant_Distance_Likelihood_Information_Distortion"          "Symmetrized_KL_Distortion_Likelihood_Information_Distortion"         
[23] "Likelihood_Gaussian_Fisher_Distortion"                                "log_Det_Likelihood_Gaussian_Fisher_Distortion"                       
[25] "Eigenvalue_Spectrum_Likelihood_Gaussian_Fisher_Distortion"            "Spectrum_Condition_Number_Likelihood_Gaussian_Fisher_Distortion"     
[27] "Max_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                 "Min_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"                
[29] "Mean_Log_Eigenvalue_Likelihood_Gaussian_Fisher_Distortion"            "Log_Eigenvalue_Variance_Likelihood_Gaussian_Fisher_Distortion"       
[31] "Affine_Invariant_Distance_Likelihood_Gaussian_Fisher_Distortion"      "Symmetrized_KL_Distortion_Likelihood_Gaussian_Fisher_Distortion"     
[33] "Likelihood_Information_Distortion_Reconstituted"                      "Likelihood_Sample_Distortion"                                        
[35] "log_Det_Likelihood_Sample_Distortion"                                 "Likelihood_Correlation_Distortion"                                   
[37] "log_Det_Likelihood_Correlation_Distortion"                            "Likelihood_Fisher_Covariance"                                        
[39] "log_Det_Likelihood_Fisher_Covariance"                                 "Correlation_Of_Likelihood_Fisher_Covariance"                         
[41] "Eigenvalue_Spectrum_Likelihood_Fisher_Covariance"                     "Effective_Rank_Likelihood_Fisher_Covariance"                         
[43] "Spectrum_Condition_Number_Likelihood_Fisher_Covariance"               "Null_Space_Projector_Likelihood_Fisher_Covariance"                   
[45] "Worst_Subspace_Projector_Likelihood_Fisher_Covariance"                "Likelihood_Distortion_Corrected_Covariance"                          
[47] "log_Det_Likelihood_Distortion_Corrected_Covariance"                   "Correlation_Of_Likelihood_Distortion_Corrected_Covariance"           
[49] "Eigenvalue_Spectrum_Likelihood_Distortion_Corrected_Covariance"       "Effective_Rank_Likelihood_Distortion_Corrected_Covariance"           
[51] "Spectrum_Condition_Number_Likelihood_Distortion_Corrected_Covariance" "Null_Space_Projector_Likelihood_Distortion_Corrected_Covariance"     
[53] "Worst_Subspace_Projector_Likelihood_Distortion_Corrected_Covariance"  "Likelihood_Distortion_Induced_Bias"                                  
[55] "Report_integral_logL"                                                 "Report_integral_y_mean"                                              
[57] "Report_integral_y_var"                                                "Report_integral_r_std"                                               
[59] "Report_integral_dlogL"                                               

```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Likelihood_Numerical_Fisher_Information")

noise_in_conductance_tau_plot =unique(d_mean$noise_in_conductance_tau)

for (al in c("macro_IR","macro_R")){
noise_in_conductance_tau_plot =d_mean%>%filter(algorithm == al)%>%select(noise_in_conductance_tau)%>%distinct()%>%pull()
  
for (comp in components_to_plot) 
for (noi in noise_in_conductance_tau_plot)
    {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      Num_ch>5,
      component == comp, 
      statistic == "value", 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      noise_in_conductance_tau==noi, 
      interval_in_tau>0.001,
      param_col==param_index)
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = Num_ch, color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(interval_in_tau~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  p2 <- ggplot(plot_data, aes(x = interval_in_tau , color = factor(nsamples))) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin =quantile_0.025 , ymax = quantile_0.975, color=param_name), width = 0.05) +
    geom_errorbar(aes(ymin =quantile_0.16 , ymax = quantile_0.84, color=factor(nsamples)), width = 0.05) +
    scale_x_log10() +
  #  facet_grid(interval_in_tau~param_name, scales = "free") +
    facet_wrap(Num_ch~param_name+nsamples, scales = "free", ncol=6) +
    geom_hline(yintercept = 1, color="green")+
    scale_y_log10()+
    ggtitle(paste0(al," ",comp, " noise:", noi)) 
  
  
  print(p)
  print(p2)
}
}

```




```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c("Max_Eigenvalue_Likelihood_Information_Distortion",
                        
"Min_Eigenvalue_Likelihood_Information_Distortion"                )


noise_in_conductance_tau_plot =unique(d_mean$noise_in_conductance_tau)

for (al in c("macro_IR","macro_R"))
  
for (noi in noise_in_conductance_tau_plot)

 {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      noise_in_conductance_tau == noi, 
      statistic == "value", 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      component%in% components_to_plot
    )
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = interval_in_tau, color = component)) +
    geom_point(aes(y = quantile_0.5)) +
    geom_errorbar(aes(ymin = quantile_0.16, ymax = quantile_0.84, color=component), width = 0.05) +
    scale_x_log10() +
    facet_grid(Num_ch ~ component, scales = "free") +
    geom_hline(yintercept = 1.0, color="green")+
    ggtitle(paste0(al," ",noi)) 
  
  print(p)
}


```



```{r,fig.width=20,fig.height=20}
  
  library(dplyr)
library(ggplot2)

num_samples <- c("1.00" = 10, "0.5" = 20, "0.20" = 50,"0.1" = 100, "0.05" = 200, "0.02" = 500, "0.01" = 1000 )
components_to_plot <- c(                   "Log_Eigenvalue_Variance_Likelihood_Information_Distortion",    
"Affine_Invariant_Distance_Likelihood_Information_Distortion",   "Symmetrized_KL_Distortion_Likelihood_Information_Distortion"   )


noise_in_conductance_tau_plot =unique(d_mean$noise_in_conductance_tau)

for (al in c("macro_IR","macro_R"))
  
for (noi in noise_in_conductance_tau_plot)

 {
  
  # 1. Filter and compute scaled values safely beforehand
  plot_data <- d_mean %>%
    filter(
      noise_in_conductance_tau == noi, 
      statistic == "value", 
      algorithm == al, 
      simulation_algorithm == "uniformization", 
      component%in% components_to_plot
    )
  
  # 2. Build and print the plot
  p <- ggplot(plot_data, aes(x = interval_in_tau, color = component)) +
    geom_point(aes(y = mean_NA)) +
    geom_errorbar(aes(ymin = quantile_0.025, ymax = quantile_0.975, color=component), width = 0.05) +
    scale_x_log10() +
    facet_grid(Num_ch ~ component, scales = "free") +
    geom_hline(yintercept = ifelse(comp=="Moment_statistics_Sum_r2_std",1,0), color="green")+
    ggtitle(paste0(al," ",noi)) 
  
  print(p)
}


```




```{r,fig.width=40,fig.height=80}

pars=d_mean%>%filter(component=="Likelihood_Information_Distortion", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Likelihood_Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()


#for (p in pars )
 for (a in c("macro_IR","macro_R"))
{print(
d_mean%>%filter(component%in%c("Numerical_Fisher_Information","Moment_statistics_Sum_dlogL"), statistic %in% c("value","covariance"), algorithm==a , param_index==param_col, simulation_algorithm!="lsubsteps_1000")%>%ggplot()+geom_point(aes(x = component, y = mean_NA, color=algorithm))+geom_errorbar(aes(x=component, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+scale_y_log10()+facet_wrap(Num_ch~interval_in_tau+param_name, scales="free", ncol=6) 
)
}
```

