---
title: "Figure 2"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

```{r}

filename=c(
  #"data/figure_2_basic_diagnostics__0420_1",
  #"data/figure_2_basic_diagnostics__0420_2_001", 
  #"data/figure_2_basic_diagnostics__0421_3_001",
  #"data/figure_2_basic_diagnostics__0422_4_001", 
  #"data/figure_2_basic_diagnostics__0513_004",
  "data/figure_2_basic_diagnostics__0515_001",
  "data/figure_2_basic_diagnostics__0516_001",
  "data/figure_2_basic_diagnostics__0517_001"
  
  
  
  )
  
algorithms_to_exclude=list(c("macro_rIRT","macro_rIR"),
                           c("macro_rIRT","macro_rIR"),
                           c("macro_rMR","macro_rIR")  )


```

```{r}
d <- map2_dfr(filename, algorithms_to_exclude, function(file, exclude) {
  
  read.csv(paste0(file, ".csv")) %>% 
    filter(!algorithm %in% exclude)
    
})%>%bind_rows()

```

```{r}
d<- d %>%
  select(where(~ !all(is.na(.))))
```

```{r}
d%>%select(component_path)%>%distinct()%>%pull()

```

```{r}
d<-d%>%mutate(component= component_path %>%
           str_replace("^Probit_statistics_", "") )
   
     
```

```{r}
d%>%select(component)%>%distinct()%>%pull()

```



```{r}
d_mean<-
  d%>%filter(!component%in%c("Probit_statistics","count_r_std","count_dlogL"))%>%
  #select(component_path,component,param_name,algorithm,sample_index,calculus,statistic,param_index,param_col,param_name,simulation_algorithm,noise_in_conductance_tau,Num_ch,interval_in_tau,statistic,probit,quantile_level,value,value_row,value_col)%>%
  pivot_wider(names_from = c(probit,quantile_level), values_from = value)
d_mean
```

```{r}
scalar_components<-d_mean%>%filter(statistic%in%c("mean", "value"), is.na(param_col),is.na(param_index) , !is.na(mean_NA))%>%select(component)%>%distinct()%>%pull()
scalar_components

```
```{r}
d_algo<-
  d%>%filter(!component%in%c("Probit_statistics","count_r_std","count_dlogL"))%>%
  #select(component_path,component,param_name,algorithm,sample_index,calculus,statistic,param_index,param_col,param_name,simulation_algorithm,noise_in_conductance_tau,Num_ch,interval_in_tau,statistic,probit,quantile_level,value,value_row,value_col)%>%
  pivot_wider(names_from = c(algorithm,probit,quantile_level), values_from = value)
d_algo

```
```{r,fig.width=20,fig.height=15}

#for (nch in unique(d_mean$Num_ch))
{print(
d_mean%>%filter(component=="Moment_statistics_Sum_logL", statistic=="mean", !algorithm%in%c("macro_MNR", "macro_NR"))%>%ggplot()+geom_point(aes(x =algorithm , y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+facet_wrap(Num_ch~interval_in_tau+noise_in_conductance_tau,nrow = length(unique(d_mean$Num_ch)), scales="free", )+ggtitle(paste0("Moment_statistics_Sum_logL for Nch=")) 
)
}
```
```{r,fig.width=20,fig.height=20}
for (com in scalar_components)

{print(
d_mean%>%filter(component==com, statistic%in%c("mean","value"), !algorithm%in%c("macro_MNR", "macro_NR"),  statistic%in%c("mean","value"))%>%ggplot()+geom_point(aes(x =noise_in_conductance_tau , y = mean_NA, color=algorithm))+geom_errorbar(aes(x=noise_in_conductance_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+facet_wrap(Num_ch~interval_in_tau+algorithm, ncol=14,scales="free", )+ggtitle(paste0(com)) 
)
}
```

"Moment_statistics_evaluation_time"                         "Numerical_Fisher_Information"                             
[13] "Information_Distortion_Matrix"                             "log_Det_Information_Distortion_Matrix"                     "Gaussian_Fisher_Distortion"                               
[16] "log_Det_Gaussian_Fisher_Distortion"                        "Information_Distortion_Reconstituted"                      "Sample_Distortion_Matrix"                                 
[19] "log_Det_Sample_Distortion_Matrix"                          "Correlation_Distortion_Matrix"                             "log_Det_Correlation_Distortion_Matrix"                    
[22] "Fisher_Covariance"                                         "log_Det_Fisher_Covariance"                                 "Correlation_Of_Fisher_Covariance"                         
[25] "Eigenvalue_Spectrum_Fisher_Covariance"                     "Effective_Rank_Fisher_Covariance"                          "Spectrum_Condition_Number_Fisher_Covariance"              
[28] "Null_Space_Projector_Fisher_Covariance"                    "Worst_Subspace_Projector_Fisher_Covariance"                "Distortion_Corrected_Covariance"                          
[31] "log_Det_Distortion_Corrected_Covariance"                   "Correlation_Of_Distortion_Corrected_Covariance"            "Eigenvalue_Spectrum_Distortion_Corrected_Covariance"      
[34] "Effective_Rank_Distortion_Corrected_Covariance"            "Spectrum_Condition_Number_Distortion_Corrected_Covariance" "Null_Space_Projector_Distortion_Corrected_Covariance"     
[37] "Worst_Subspace_Projector_Distortion_Corrected_Covariance"  "Distortion_Induced_Bias"                                   "Report_integral_logL"                                     
[40] "Report_integral_y_mean"                                    "Report_integral_y_var"                                     "Report_integral_r_std"                                    
[43] "Report_integral_dlogL"  


```{r,fig.width=20,fig.height=20}

print(
d_mean%>%filter(component%in% c("Spectrum_Condition_Number_Fisher_Covariance","Spectrum_Condition_Number_Distortion_Corrected_Covariance"), statistic=="value", !algorithm%in%c("macro_MNR", "macro_NR"), simulation_algorithm!="funiformization")%>%ggplot()+geom_point(aes(x =simulation_algorithm , y = mean_NA, color=component))+geom_errorbar(aes(x=simulation_algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=component))+facet_wrap(algorithm~interval_in_tau, ncol=4,scales="free", )+ggtitle(paste0("Spectrum_Condition_Number_Fisher_Covariance "))
)


```



Affine_Invariant_Distance_Information_Distortion_Matrix
Moment_statistics_evaluation_time
```{r,fig.width=20,fig.height=20}

print(
d_mean%>%filter(component%in% c("Affine_Invariant_Distance_Information_Distortion_Matrix"), statistic=="value", !algorithm%in%c("macro_MtNR", "macro_NtR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x =interval_in_tau , y = mean_NA, color=component))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=component))+facet_wrap(Num_ch~algorithm, ncol=5,scales="free", )+ggtitle(paste0("Affine_Invariant_Distance_Information_Distortion_Matrix "))+scale_x_log10()
)


```

```{r,fig.width=20,fig.height=20}

for (nch in unique(d_mean$Num_ch))
{print(
d_mean%>%filter(component=="Moment_statistics_evaluation_time", statistic=="mean",  !algorithm%in%c("macro_rMNR", "macro_rNR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x =algorithm , y = mean_NA, color=algorithm))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+facet_wrap(Num_ch~interval_in_tau,ncol = 7, scales="free", )+ggtitle(paste0("Moment_statistics_evaluation_time for Nch=",nch)) 
)
}
```

log_Det_Information_Distortion_Matrix

```{r,fig.width=20,fig.height=12}

for (nch in unique(d_mean$Num_ch))
{print(
d_mean%>%filter(component=="log_Det_Information_Distortion_Matrix", statistic=="value", Num_ch==nch, !algorithm%in%c("macro_MNR", "macro_NR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x =algorithm , y = mean_NA, color=algorithm))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+geom_hline(yintercept = 0,color="green")+facet_wrap(.~interval_in_tau,nrow = 5, scales="free", )+ggtitle(paste0("Moment_statistics_Sum_logL for Nch=",nch)) 
)
}
```

```{r,fig.width=20,fig.height=4}

for (nch in unique(d_mean$Num_ch))
{print(
d_mean%>%filter(component=="log_Det_Distortion_Corrected_Covariance", statistic=="value", Num_ch==nch, !algorithm%in%c("macro_MNR", "macro_NR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x =algorithm , y = mean_NA, color=algorithm))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+facet_wrap(.~interval_in_tau,nrow = 2, scales="free", )+ggtitle(paste0("Moment_statistics_Sum_logL for Nch=",nch)) 
)
}
```

```{r,fig.width=20,fig.height=4}

for (nch in unique(d_mean$Num_ch))
{print(
d_mean%>%filter(component=="Moment_statistics_Sum_trust_coefficient", statistic=="mean", Num_ch==nch, !algorithm%in%c("macro_MNR", "macro_NR"))%>%ggplot()+geom_point(aes(x = algorithm, y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+facet_wrap(.~interval_in_tau,nrow = 2, scales="free", )+ggtitle(paste0("Moment_statistics_Sum_trust_coefficient for Nch=",nch)) 
)
}
```

Gaussian_Fisher_Distortion


```{r,fig.width=20,fig.height=10}

pars=d_mean%>%filter(component=="Information_Distortion_Matrix", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
#  for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component%in%c("Information_Distortion_Matrix"), statistic=="value", param_name==p, algorithm%in%c("macro_IR","macro_MR","macro_MNR","macro_R","macro_NR"))%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=component))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=component))+geom_hline(yintercept = 1, color="green")+scale_y_log10()+scale_x_log10()+facet_wrap(Num_ch~algorithm, scales="free", ncol=5)+ggtitle(paste0("Information_Distortion_Matrix for ",p)) 
)
}
```


Numerical_Fisher_Information
Moment_statistics_Sum_dlogL

```{r,fig.width=20,fig.height=20}

pars=d_mean%>%filter(component=="Information_Distortion_Matrix", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
#  for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component%in%c("Numerical_Fisher_Information","Moment_statistics_Sum_dlogL"), statistic %in% c("value","covariance"), param_name==p, algorithm%in%c("macro_IR","macro_IRT","macro_MR"), simulation_algorithm!="lsubsteps_1000")%>%ggplot()+geom_point(aes(x = component, y = mean_NA, color=algorithm))+geom_errorbar(aes(x=component, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+scale_y_log10()+facet_wrap(Num_ch~interval_in_tau, scales="free", ncol=7)+ggtitle(paste0("Information_Distortion_Matrix for ",p)) 
)
}
```
log_Det_Fisher_Covariance"                                 "Correlation_Of_Fisher_Covariance"                         
[25] "Eigenvalue_Spectrum_Fisher_Covariance"                     "Effective_Rank_Fisher_Covariance"                          "Spectrum_Condition_Number_Fisher_Covariance"              
[28] "Null_Space_Projector_Fisher_Covariance"                    "Worst_Subspace_Projector_Fisher_Covariance"                "Distortion_Corrected_Covariance"                          
[31] "log_Det_Distortion_Corrected_Covariance"                   "Correlation_Of_Distortion_Corrected_Covariance"            "Eigenvalue_Spectrum_Distortion_Corrected_Covariance"

```{r,fig.width=20,fig.height=20}

pars=d_mean%>%filter(component=="Information_Distortion_Matrix", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
 for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component%in%c("Numerical_Fisher_Information","Moment_statistics_Sum_dlogL"), statistic %in% c("value","covariance"), param_name==p, algorithm%in%c("macro_IR","macro_IRT","macro_MR"), simulation_algorithm!="lsubsteps_1000")%>%ggplot()+geom_point(aes(x = component, y = mean_NA, color=algorithm))+geom_errorbar(aes(x=component, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=algorithm))+scale_y_log10()+facet_wrap(Num_ch~interval_in_tau, scales="free", )+ggtitle(paste0("Information_Distortion_Matrix for ",p, " ", a)) 
)
}
```


Gaussian_Fisher_Distortion

```{r,fig.width=20,fig.height=4}

pars=d_mean%>%filter(component=="Gaussian_Fisher_Distortion", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
#  for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component=="Gaussian_Fisher_Distortion", statistic=="value", param_name==p, algorithm%in%c("macro_IR","macro_IRT","macro_MR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+geom_hline(yintercept = 1, color="green")+scale_y_log10()+facet_wrap(Num_ch~algorithm,nrow = 5, scales="free", )+ggtitle(paste0("Gaussian_Fisher_Distortion for ",p)) 
)
}
```

Sample_Distortion_Matrix
Correlation_Distortion_Matrix

```{r,fig.width=20,fig.height=4}

pars=d_mean%>%filter(component=="Sample_Distortion_Matrix", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
#  for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component%in%c("Sample_Distortion_Matrix", "Correlation_Distortion_Matrix","Information_Distortion_Matrix","Report_integral_dlogL"),statistic%in%c("value","integral_correlation_lag2"), param_name==p, algorithm%in%c("macro_IR","macro_IRT","macro_MR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=component))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=component))+geom_hline(yintercept = 1, color="green")+scale_x_log10()+scale_y_log10()+facet_wrap(algorithm~Num_ch,nrow = 2, scales="free", )+ggtitle(paste0("Sample_Distortion_Matrix for ",p)) 
)
}
```

```{r,fig.width=20,fig.height=4}

pars=d_mean%>%filter(component=="Correlation_Distortion_Matrix", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
#  for (a in c("macro_IR","macro_MR"))
{print(
d_mean%>%filter(component=="Correlation_Distortion_Matrix", statistic=="value", param_name==p, algorithm%in%c("macro_IR","macro_IRT","macro_MR"), simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+geom_hline(yintercept = 1, color="green")+scale_y_log10()+facet_wrap(Num_ch~algorithm,nrow = 2, scales="free", )+ggtitle(paste0("Correlation_Distortion_Matrix for ",p)) 
)
}
```
```{r,fig.width=20,fig.height=20}

pars=d_mean%>%filter(component=="Distortion_Corrected_Covariance", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()

for (p in pars )
{print(
d_mean%>%filter(component%in%c("Distortion_Corrected_Covariance"),simulation_algorithm=="uniformization", statistic=="value", param_name==p, !algorithm%in%c("macro_MNR", "macro_NR"))%>%ggplot()+geom_point(aes(x = algorithm, y = mean_NA, color=component))+geom_errorbar(aes(x=algorithm, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=component))+scale_y_log10()+facet_wrap(Num_ch~interval_in_tau,ncol = 7, scales="free", )+ggtitle(paste0("Moment_statistics_Sum_logL for ",p)) 
)
}
```


```{r,fig.width=20,fig.height=20}

pars=d_mean%>%filter(component=="Distortion_Corrected_Covariance", param_index!="", param_col==param_index)%>%select(param_name)%>%distinct()%>%pull()

al=d_mean%>%filter(component=="Distortion_Corrected_Covariance", param_index!="", param_col==param_index)%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
{print(
d_mean%>%filter(component%in%c("Fisher_Covariance"),simulation_algorithm=="uniformization", statistic=="value", param_name==p, algorithm%in%c("macro_IR", "macro_NR_"))%>%mutate(noise_in_conductance_tau=factor(noise_in_conductance_tau))%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=noise_in_conductance_tau))+geom_errorbar(aes(x=interval_in_tau, y= (mean_NA),ymin = quantile_0.025,ymax = quantile_0.975 ,color=noise_in_conductance_tau))+scale_y_log10()+scale_x_log10()+facet_wrap(Num_ch~algorithm+noise_in_conductance_tau,ncol = 5, scales="free", )+ggtitle(paste0("Moment_statistics_Sum_logL for ",p)) 
)
}
```






```{r,fig.width=8,fig.height=5}

pars=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
  for (a in c("macro_IR","macro_IRT"))
{print(
d_mean%>%filter(component=="Distortion_Induced_Bias", statistic=="value", param_name==p, algorithm==a,simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_line(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=interval_in_tau, y= mean_NA,ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+geom_hline(yintercept = 0, color="green")+facet_wrap(Num_ch~simulation_algorithm,ncol = 3, scales="free", )+ggtitle(paste0(a," Distortion_Induced_Bias for ",p)) 
)
}
```



```{r,fig.width=8,fig.height=5}

pars=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(param_name)%>%distinct()%>%pull()
al=d_mean%>%filter(component=="Distortion_Induced_Bias", param_index!="")%>%select(algorithm)%>%distinct()%>%pull()

for (p in pars )
  for (a in c("macro_IR","macro_IRT"))
{print(
d_mean%>%filter(component=="Moment_statistics_Sum_dlogL", statistic=="mean", param_name==p, algorithm==a,simulation_algorithm=="uniformization")%>%ggplot()+geom_point(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_line(aes(x = interval_in_tau, y = mean_NA, color=simulation_algorithm))+geom_errorbar(aes(x=interval_in_tau, y= mean_NA,ymin = quantile_0.025,ymax = quantile_0.975 ,color=simulation_algorithm))+geom_hline(yintercept = 0, color="green")+facet_wrap(Num_ch~simulation_algorithm,ncol = 3, scales="free", )+ggtitle(paste0(a," Moment_statistics_Sum_dlogL for ",p)) 
)
}
```





