---
title: "Figure 1 — the interval-likelihood ladder, one filter step (R, MR, VR, IR)"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

The paper's Figure 1. Four columns, the **recursive** ladder only: `R` → `MR` → `VR` → `IR`, in the
order in which each rung adds structure to what the likelihood conditions on. The non-recursive
members are not measured in paper 1 (`papers/1_method/decisions.md`, A-strict), so they are not shown
here; `figure_1_all.Rmd` renders the same figure with every algorithm for exploration.

Everything below the roster lives in `figure_1_panels.R`, which both notebooks share. Edit the panels
there, not here, or the two versions drift.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r panels, include=FALSE}
source("figure_1_panels.R")
```

```{r roster}
COLS <- c("R", "MR", "VR", "IR")
```

```{r figures}
build_figure(0:4,     "Figure_S1.pdf", COLS)   # samples 0-4 (0 to 10 ms of the 12 ms record)
build_figure(c(3, 4), "Figure_1.pdf",  COLS)   # the 6 to 10 ms zoom
```
