\documentclass[11pt]{article}

\usepackage{amsmath, amssymb, amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=1in}

\title{Supplementary Information:\\
Per-Replicate Safety Categorization\\
in the Posterior Information Distortion Framework}
\author{}
\date{}

\begin{document}

\maketitle

\begin{abstract}
\noindent We formalize a per-replicate safety categorization policy that
operationalizes the four-test algorithm validation framework of
\texttt{supplement\_algorithm\_validation.tex} into a four-category
classification: SAFE, MARGINAL, UNRELIABLE, BIASED. The classification
gates which diagnostics are reportable for each bootstrap replicate
and each cell of the figure\_2 sweep. The taxonomy is built on two
independent axes --- spectral health of the likelihood Hessian
$\mathbf H_{\mathrm{lik}}$ and pattern of the four validation tests
$(\mathcal D(\mathbf C_{\mathrm{IDM}}^{\mathrm{post}}),
\mathcal D(\mathbf{GFD}^{\mathrm{post}}),
\mathcal D(\mathbf C_{\mathrm{match}}), T^2)$ --- with thresholds
derived from existing numerical tolerances in the codebase
($\tau_{\mathrm{safe}}$, $\tau_{\mathrm{singular}}$ from
\texttt{lapack\_headers.h}) and the universal noise scale
$2\sqrt{p/(B-1)}$ for the matrix diagnostics. We specify per-category
reporting policies, cross-replicate aggregation rules at the
observable level, and the binary-output column schema for downstream
analysis scripts.
\end{abstract}

\section*{1. Scope and Position in the PID Framework}
\label{sec:sc-scope}

This document operationalizes the algorithm-validation framework
([[supplement-algorithm-validation]]) as a per-replicate per-cell
safety categorization. It does not introduce new diagnostics; it
specifies the decision policy by which the existing diagnostics gate
the validity of downstream inference outputs (evidence, credible
intervals, information gain reports).

The categorization is per-replicate $b \in 1,\ldots,B$ within each
cell of the figure\_2 sweep, with cross-replicate aggregation to the
per-cell level for table and figure reporting. It is operationally
distinct from the $\theta_{\mathrm{MAP}}$-level matrix diagnostics:
those produce numbers; this produces decisions.

\paragraph{Disambiguation: ``BIASED'' here is not the DIB matrix.}
``BIASED'' as a safety category refers to estimator-level structural
failure of the algorithm, NOT to the Distortion-Induced Bias matrix
$\mathbf b_{\mathrm{DIB},\mathrm{post}}$ of Section 7 of
[[supplement-posterior-main]]. The DIB measures bias of the posterior
mean estimate; the BIASED category here flags algorithm structural
failure of the kind documented in Section 7 of
[[supplement-algorithm-validation]] (the fourth regime). The two
share a word but not a referent.

\paragraph{Operational scope.}
The four-category classification is the unit of safety information
exposed to users:
\begin{itemize}
\item As an integer enum column in the bootstrap-samples binary
  (\texttt{safety\_category} $\in \{0,1,2,3\}$)
\item As a per-cell summary in the figure\_2 analysis CSV
\item As a categorical gate in downstream plotting (which replicates
  contribute to quantiles, which observable cells are reportable)
\item As a textual flag in published tables and figures
\end{itemize}

\section*{2. The Two Decision Axes}
\label{sec:sc-two-axes}

The categorization combines two independent axes:

\subsection*{2.1 Spectral health of $\mathbf H_{\mathrm{lik}}$}
\label{sec:sc-spectral-axis}

The first axis is the position of the minimum eigenvalue of the
likelihood Hessian relative to two numerical tolerances:
\begin{equation}
\tau_{\mathrm{singular}}
\;=\;
\mathrm{atol}_{\mathrm{psd}} \;+\; \mathrm{rtol}_{\mathrm{psd}} \cdot \lambda_{\max}(\mathbf H_{\mathrm{lik}}),
\qquad
\tau_{\mathrm{safe}}
\;=\;
10^4 \cdot \tau_{\mathrm{singular}}.
\label{eq:sc-tolerances}
\end{equation}
The constants $\mathrm{atol}_{\mathrm{psd}}$ and $\mathrm{rtol}_{\mathrm{psd}}$
are inherited from the existing
\texttt{subspace\_tol} pattern in
\texttt{legacy/lapack\_headers.h} (currently $\mathrm{atol} = 0$,
$\mathrm{rtol} = 10^{-10}$); the factor of $10^4$ separating
$\tau_{\mathrm{safe}}$ from $\tau_{\mathrm{singular}}$ reflects four
orders of magnitude between ``numerically reliable as positive'' and
``numerically indistinguishable from zero''.

The four spectral states are:
\begin{itemize}
\item $\lambda_{\min}(\mathbf H_{\mathrm{lik}}) > \tau_{\mathrm{safe}}$:
  \textbf{safely PD}. Likelihood Hessian is unambiguously positive
  definite. Likelihood IDM is computable without retained-eigenspace
  machinery.
\item $\tau_{\mathrm{singular}} < \lambda_{\min}(\mathbf H_{\mathrm{lik}}) \leq \tau_{\mathrm{safe}}$:
  \textbf{marginally PD}. Positive but within numerical reach of zero.
  Likelihood IDM is technically defined but the smallest eigenvalue
  estimate is unreliable. Posterior IDM is the diagnostic of record.
\item $|\lambda_{\min}(\mathbf H_{\mathrm{lik}})| \leq \tau_{\mathrm{singular}}$:
  \textbf{singular}. Numerically indistinguishable from zero.
  Likelihood IDM requires retained-eigenspace projection
  ([[implementation-subspace-information-distortion]]). Posterior IDM
  is well-defined under the dominance condition of
  [[supplement-posterior-main]] Section 2.1.
\item $\lambda_{\min}(\mathbf H_{\mathrm{lik}}) < -\tau_{\mathrm{singular}}$:
  \textbf{indefinite}. The likelihood Hessian has substantively
  negative eigenvalues. This is the regime where the dominance
  condition is non-trivial and where the fourth-regime signature of
  [[supplement-algorithm-validation]] Section 7 manifests when
  coincident with $\mathrm{rank}(\mathbf G_{\mathrm{lik}}) < p$.
\end{itemize}

The four states are exhaustive and mutually exclusive. The spectral
axis can be summarized by an integer code
\texttt{spectral\_state} $\in \{0, 1, 2, 3\}$ for safely PD, marginally
PD, singular, indefinite respectively.

\subsection*{2.2 Pattern of validation tests}
\label{sec:sc-validation-axis}

The second axis is the joint pattern of the four canonical tests of
[[supplement-algorithm-validation]] Section 4. For each replicate
the four scalar statistics are computed:
\begin{equation}
\bigl(
\mathcal D(\mathbf C_{\mathrm{IDM}}^{\mathrm{post}}),\;
\mathcal D(\mathbf{GFD}^{\mathrm{post}}),\;
\mathcal D(\mathbf C_{\mathrm{match}}),\;
T^2
\bigr)
\label{eq:sc-four-tests}
\end{equation}
and compared against their thresholds:
$\tau_{\mathcal D} = 2\sqrt{p/(B-1)}$ for the three matrix tests and
$\chi^2_{p,\,0.95}$ for the Wald statistic. The number of failed
tests is an integer between 0 and 4; the identity of the failed
tests further classifies the failure mode.

\paragraph{Production scope.}
In production use, $\theta_{\mathrm{sim}}$ is unknown and $T^2$ is
unavailable. The validation axis reduces to three tests. The safety
categorization defined in Section 4 below remains well-defined; only
the BIASED-vs-UNRELIABLE distinction loses some discriminating power
in production, as detailed there.

\section*{3. The Combined Categorization Rule}
\label{sec:sc-combined-rule}

The two axes combine into the four-category SAFE/MARGINAL/UNRELIABLE/BIASED
classification:

\begin{equation}
\boxed{\quad
\begin{array}{l|l}
\text{Spectral state} & \text{Worst-case category given validation pattern} \\
\hline
\text{safely PD} & \text{SAFE if 0--1 validation tests fail, MARGINAL if 2+} \\
\text{marginally PD} & \text{MARGINAL if 0--1 tests fail, UNRELIABLE if 2+} \\
\text{singular} & \text{UNRELIABLE if 0--2 tests fail, BIASED if 3+} \\
\text{indefinite} & \text{BIASED (always)}
\end{array}
\quad}
\label{eq:sc-combined-rule}
\end{equation}

The promotion rules (worse spectral state $\Longrightarrow$ at least
one category worse than the validation alone) reflect the
operational principle: spectral indeterminacy is always a deeper
problem than statistical-test failure, because it propagates to
\emph{all} downstream inversions and not just one diagnostic.

\subsection*{3.1 Operational definitions}

\paragraph{SAFE.} Likelihood Hessian is safely PD AND at most one
validation test marginally exceeds threshold. The Likelihood IDM,
Posterior IDM, Distortion-Corrected Covariance, Distortion-Induced
Bias, GFD, and all spectral scalars (\texttt{idm\_min\_eigenvalue},
\texttt{gfd\_log\_det}, etc.) are computed and reported without
caveat. The replicate participates in all cross-replicate aggregates
including tail quantiles ($q_{0.025}$, $q_{0.975}$). The integer
code is \texttt{safety\_category} $= 0$.

\paragraph{MARGINAL.} Either (a) spectral state is marginally PD with
at most one test failure, or (b) spectral state is safely PD with two
or more test failures. The Posterior IDM and DCC are the diagnostics
of record; the Likelihood IDM is reported with a flag column or
NaN-filled depending on the specific failure pattern. The replicate
participates in median-based aggregates but is excluded from tail
quantiles. The integer code is $= 1$.

\paragraph{UNRELIABLE.} Either (a) spectral state is singular with at
most two test failures, or (b) spectral state is marginally PD with
two or more failures. The Posterior IDM is reported only with caveat;
Likelihood IDM is NaN-filled (or reported on a reduced retained
subspace if available). The replicate is excluded from tail quantiles
and from any median quantile that would otherwise be visualized at
publication-grade.
The integer code is $= 2$.

\paragraph{BIASED.} Either (a) spectral state is indefinite (any
$\lambda_{\min}(\mathbf H_{\mathrm{lik}}) < -\tau_{\mathrm{singular}}$),
or (b) spectral state is singular with three or four test failures.
All distortion shapes are NaN-filled in the binary output; raw
spectral diagnostics (the bare eigenvalue lists of
$\mathbf H_{\mathrm{lik}}$ and $\mathbf G_{\mathrm{lik}}$) are
preserved for forensic inspection following the
\texttt{compute\_psd\_decomp} spectrum-only fallback pattern at
\texttt{legacy/lapack\_headers.h}. The replicate is excluded entirely
from cross-replicate quantiles and medians. The integer code is $= 3$.

\subsection*{3.2 Why two axes rather than one}

A spectral-only categorization fails to detect classical
misspecification (regime 1 of [[supplement-algorithm-validation]]
Section 7) because the likelihood Hessian can be perfectly
well-conditioned while the Bartlett identity fails. A
validation-only categorization fails to detect rank-deficiency or
indefiniteness that may not manifest in the dimensionless tests when
the prior already dominates the affected subspace.

The combined rule \eqref{eq:sc-combined-rule} captures both failure
modes. Section 7 below specifies the per-direction decomposition that
refines this categorization when the failure pattern is direction-
specific.

\section*{4. Per-Category Reporting Policy}
\label{sec:sc-reporting-policy}

Per replicate, each category $c$ specifies an output policy
$\pi(c)$ for the binary samples and the analysis CSV:

\begin{itemize}
\item[$c = 0$ (SAFE):] All matrix observables computed and emitted
  (\texttt{idm}, \texttt{gfd}, \texttt{sdm}, \texttt{cdm},
  \texttt{fisher\_covariance}, \texttt{dcc}, \texttt{numerical\_fim},
  \texttt{idr}). All scalar summaries
  (\texttt{idm\_log\_det}, \texttt{gfd\_min\_eigenvalue}, etc.) emitted.
  Replicate index in the cross-replicate \texttt{quantile\_mask} is set
  to 1.

\item[$c = 1$ (MARGINAL):] All Posterior matrix observables emitted;
  Likelihood matrix observables emitted with a sister
  \texttt{idm\_flag} column equal to ``marginal''. Scalar summaries
  emitted for Posterior, NaN-filled for Likelihood. Replicate index
  in \texttt{tail\_quantile\_mask} is set to 0 but in
  \texttt{median\_mask} is 1.

\item[$c = 2$ (UNRELIABLE):] Posterior matrix observables emitted with
  caveat tag. Likelihood matrix observables NaN-filled. Replicate
  index in \texttt{median\_mask} is set to 0; only the raw spectral
  diagnostic columns
  (\texttt{idm\_spectrum}, \texttt{gfd\_spectrum}) are populated.

\item[$c = 3$ (BIASED):] All distortion-shape observables NaN-filled.
  Bare spectral observables of $\mathbf H_{\mathrm{lik}}$ and
  $\mathbf G_{\mathrm{lik}}$ emitted for forensic inspection.
  Replicate excluded from all cross-replicate aggregates.
\end{itemize}

This policy is encoded as a static table consulted by
\texttt{write\_bootstrap\_samples\_binary} at the point of per-
replicate observable extraction. The \texttt{safety\_category} column
itself is computed inside
\texttt{calculate\_Likelihood\_diagnostics\_preset\_f} after
$\mathbf F_b$ and $\mathbf{JT}$ are built and before the distortion
matrices are extracted, so a BIASED replicate short-circuits and
avoids the costly distortion-shape computations.

\section*{5. Cross-Replicate Aggregation to the Cell Level}
\label{sec:sc-aggregation}

Per-replicate categories aggregate to a per-cell category for table
and figure reporting. The cell-level safety category is determined by
the fraction of replicates in each per-replicate category:
\begin{equation}
\boxed{\quad
\mathrm{cell\_category}(c)
\;=\;
\begin{cases}
\mathrm{SAFE} & \text{if } f_{\mathrm{SAFE}}(c) \geq f_{\mathrm{safe}} \\
\mathrm{MARGINAL} & \text{if } f_{\mathrm{SAFE}}(c) + f_{\mathrm{MARGINAL}}(c) \geq f_{\mathrm{marginal}} \\
\mathrm{UNRELIABLE} & \text{if } f_{\mathrm{BIASED}}(c) < f_{\mathrm{biased,unrel}} \\
\mathrm{BIASED} & \text{otherwise}
\end{cases}
\quad}
\label{eq:sc-cell-aggregation}
\end{equation}
with default fractions:
\begin{itemize}
\item $f_{\mathrm{safe}} = 0.90$
\item $f_{\mathrm{marginal}} = 0.75$
\item $f_{\mathrm{biased,unrel}} = 0.10$
\end{itemize}

These fractions reflect the practical reality that at $B = 100$
replicates per cell, the granularity is 1\%: requiring all 100
replicates be SAFE before declaring a cell SAFE would demote
operationally good cells based on single noisy replicates. The 90/75/10
defaults give the framework a tolerant interpretation of replicate
noise while preserving categorical signal.

\paragraph{Cell-level operational scope.}
A SAFE cell is the publication-grade benchmark: all replicates
contribute to all aggregates without caveat; tail quantiles are
meaningful estimators of bootstrap uncertainty. A MARGINAL cell is
reportable in supplementary tables with a footnote. An UNRELIABLE
cell is reported with caveat but only median-grade aggregates are
published. A BIASED cell is the categorical failure: the cell is
reported as ``algorithm structurally fails at this $\Delta t$'' with
the per-direction decomposition of Section 7 indicating which
parameter combinations are affected.

\subsection*{5.1 Sensitivity to fraction choices}

The fraction-based aggregation has a known sensitivity: at the
boundary (e.g., $f_{\mathrm{SAFE}} = 0.89$ vs.\ $0.90$), a cell may
flip categories. This is addressed by reporting alongside the
category the actual fraction $f_{\mathrm{SAFE}}(c)$ for each
algorithm$\times$cell, allowing inspection of borderline cases.

\subsection*{5.2 Per-direction refinement at the cell level}

When a cell is categorized as MARGINAL, UNRELIABLE, or BIASED, the
underlying failure may be confined to a single parameter direction.
Section 7 specifies the per-direction reporting that refines the
cell-level summary to identify which parameters are operationally
recoverable and which are not.

\section*{6. The Special Case of Coincident Singularity and Indefiniteness}
\label{sec:sc-coincident}

The fourth regime of algorithm failure documented in
[[supplement-algorithm-validation]] Section 7.4 ---
$\mathrm{rank}(\mathbf G_{\mathrm{lik}}) < p$ coincident with
$\lambda_{\min}(\mathbf H_{\mathrm{lik}}) < 0$ in the same direction
--- requires explicit categorization beyond what
\eqref{eq:sc-combined-rule} delivers.

The signature is:
\begin{enumerate}
\item Spectral state is indefinite ($\lambda_{\min}(\mathbf H_{\mathrm{lik}}) < -\tau_{\mathrm{singular}}$).
  By rule \eqref{eq:sc-combined-rule}, category is BIASED.
\item The eigenvector $\mathbf u_-$ of the negative eigenvalue of
  $\mathbf H_{\mathrm{lik}}$ satisfies
  $\mathbf u_-^\top \mathbf G_{\mathrm{lik}} \mathbf u_- < \tau_{\mathrm{singular}}$.
  The Gaussian moment-matched FIM is rank-deficient in the same direction.
\end{enumerate}

Both conditions together constitute the fourth-regime signature.
Replicates exhibiting this signature are classified as
$\mathrm{BIASED}_{\mathrm{IV}}$ (a sub-code of the BIASED category,
emitted as
$\mathrm{safety\_category} = 4$ rather than $3$ in the binary output)
to distinguish from generic indefiniteness due to off-MLE evaluation
or numerical noise. In the aggregation \eqref{eq:sc-cell-aggregation}
both sub-codes contribute to the BIASED fraction.

The distinction matters for the failure mode reporting:
$\mathrm{BIASED}_{\mathrm{IV}}$ is the
\emph{algorithm-structural failure} of the kind documented for
macro\_R at $\Delta t = 1\tau$. Plain $\mathrm{BIASED}$ may be
recoverable by refitting to a tighter MAP tolerance or by adjusting
the prior; $\mathrm{BIASED}_{\mathrm{IV}}$ is not recoverable by any
post-hoc remediation --- the algorithm is producing a likelihood
whose moment-matched approximation has structurally lost rank.

\section*{7. Per-Direction Reporting Within Non-SAFE Categories}
\label{sec:sc-per-direction}

For a replicate categorized as MARGINAL, UNRELIABLE, or BIASED, the
failure may be confined to a subspace of the full parameter space.
Section 4 of [[supplement-posterior-main]] specifies the
identification of \emph{active} and \emph{ineffective} directions via
spectral decomposition of $\mathbf C_{\mathrm{post}}$. Here we extend
that analysis to the safety report.

\paragraph{Per-direction safety category.}
Each parameter eigendirection $\mathbf v_k$ of
$\mathbf C_{\mathrm{post}}$ has its own per-direction safety category:
\begin{itemize}
\item SAFE: $\mathbf v_k^\top \mathbf H_{\mathrm{lik}} \mathbf v_k > \tau_{\mathrm{safe}}$
  AND the projection of the failed validation test onto $\mathbf v_k$
  is below threshold.
\item MARGINAL: scaled-down version
\item UNRELIABLE: $\mathbf v_k^\top \mathbf H_{\mathrm{lik}} \mathbf v_k$ near zero
\item BIASED: $\mathbf v_k^\top \mathbf H_{\mathrm{lik}} \mathbf v_k < -\tau_{\mathrm{singular}}$
\end{itemize}
The replicate-level safety category is the worst per-direction
category, with the per-direction breakdown available for forensic
inspection.

\paragraph{Reporting in published tables.}
For each algorithm$\times$cell that is categorized as MARGINAL or
worse at the cell level, the published supplementary table lists the
parameter combination eigenvectors of $\mathbf C_{\mathrm{post}}$
that are responsible for the categorization. This makes the
operational failure mode concrete: ``the algorithm fails to
identify the linear combination $\mathbf v = (0.71, -0.71, 0, 0, 0, 0)$,
corresponding to the differential rate $k_{\mathrm{on}}^{(1)} - k_{\mathrm{on}}^{(2)}$
--- in this combination the likelihood is structurally non-convex''
(hypothetical example).

\section*{8. Implementation Hooks}
\label{sec:sc-implementation}

\subsection*{8.1 New binary output columns}

The bootstrap-samples binary writer
(\texttt{write\_bootstrap\_samples\_binary} in
\texttt{src/core/likelihood.cpp}) gains the following scalar columns:

\begin{itemize}
\item \texttt{safety\_category} (integer $\in \{0,1,2,3,4\}$):
  the per-replicate category code
\item \texttt{spectral\_state} (integer $\in \{0,1,2,3\}$):
  the spectral axis code
\item \texttt{n\_validation\_failures} (integer $\in \{0,1,2,3,4\}$):
  the number of failed validation tests
\item \texttt{H\_lik\_lambda\_min} (double): the actual
  $\lambda_{\min}(\mathbf H_{\mathrm{lik}})$ value, for sanity-check
  inspection
\item \texttt{G\_lik\_lambda\_min} (double): the corresponding
  $\lambda_{\min}(\mathbf G_{\mathrm{lik}})$, for fourth-regime
  signature detection
\item \texttt{D\_C\_match} (double): the affine-invariant distance
  of $\mathbf C_{\mathrm{match}}$ (= the new diagnostic of
  [[supplement-algorithm-validation]])
\end{itemize}

\subsection*{8.2 Computation site}

The category is computed inside
\texttt{calculate\_Likelihood\_diagnostics\_preset\_f}
(\texttt{src/core/likelihood.cpp}) after $\mathbf F_b$ and
$\mathbf{JT}$ are built and before the distortion-shape matrices are
computed. This allows BIASED replicates to short-circuit the costly
matrix extractions.

\subsection*{8.3 Downstream consumption in R}

The \texttt{figure\_2\_samples.Rmd} analysis script consumes the
\texttt{safety\_category} column as a factor and uses it for:
\begin{itemize}
\item Filtering: \texttt{boot\$samples \%>\% filter(safety\_category $\leq$ 1)}
  for SAFE+MARGINAL only aggregates
\item Faceting: per-category boxplots of distortion scalars to
  diagnose failure-mode patterns
\item Cell-level rollup: per-cell category from the fraction-based
  rule \eqref{eq:sc-cell-aggregation}, computed in R rather than C++
  for ease of fraction-threshold tuning
\end{itemize}

\subsection*{8.4 Manifest entries}

For each new column, the manifest.csv emits a corresponding row
(\texttt{observable, rows, cols, n\_replicates}). The
\texttt{reader.R} auto-generated reader propagates the new columns
into the 4D-array list automatically.

\section*{9. Worked Example: macro\_R vs.\ macro\_IR at figure\_2}
\label{sec:sc-worked-example}

For the eLife 2025 figure\_2 sweep with macro\_R and macro\_IR across
$\Delta t \in \{1, 0.5, 0.2, 0.1, 0.05, 0.02, 0.01\}\tau$ at
$N_{\mathrm{sim}} = 16384$, $B = 100$, $\sigma^2_{\mathrm{prior}} = 2$:

\subsection*{9.1 Predicted per-cell categorization}

Based on the empirical findings of
[[supplement-algorithm-validation]] Section 6 and the rule
\eqref{eq:sc-combined-rule}:

\begin{center}
\begin{tabular}{lcc}
\hline
$\Delta t$ ($\tau$) & macro\_R & macro\_IR \\
\hline
1.0 & $\mathrm{BIASED}_{\mathrm{IV}}$ & SAFE \\
0.5 & SAFE & SAFE \\
0.2 & SAFE & SAFE \\
0.1 & SAFE & SAFE \\
0.05 & SAFE & SAFE \\
0.02 & SAFE & SAFE \\
0.01 & SAFE & SAFE (with 2/100 MARGINAL flagged outliers) \\
\hline
\end{tabular}
\end{center}

The macro\_R column shows the categorical failure at $\Delta t = 1\tau$
(all 100 replicates exhibit the fourth-regime signature, so
$f_{\mathrm{BIASED}}(c) = 1.0 \gg 0.10$). The macro\_IR column is
SAFE everywhere; the 2/100 MARGINAL outliers at $\Delta t = 0.01\tau$
are reported as a footnote but do not change the cell category
($f_{\mathrm{SAFE}} = 0.98 \geq 0.90$).

\subsection*{9.2 Per-direction breakdown for macro\_R at $\Delta t = 1\tau$}

The fourth-regime failure is confined to a single parameter
combination eigendirection (the direction of the negative eigenvalue
of $\mathbf H_{\mathrm{lik}}$ coincident with the singular direction
of $\mathbf G_{\mathrm{lik}}$). The other five parameter directions
are SAFE.

For the eLife paper this translates to: macro\_R at $\Delta t = 1\tau$
fails to identify a specific linear combination of kinetic parameters.
The other identifiable combinations (the bulk of the parameter
information) are operationally recoverable from the macro\_R analysis;
the failed combination is recoverable only from macro\_IR.

The figure\_2 paper-facing tables list explicitly which parameter
combination is affected, with the eigenvector projected onto the
original kinetic parameter coordinates.

\subsection*{9.3 Reporting impact}

\begin{itemize}
\item Evidence (\texttt{log\_Z}) for macro\_R at $\Delta t = 1\tau$ is
  computed using the prior-dominated subspace decomposition of
  Section 8 of [[supplement-posterior-main]] but reported with a
  $\mathrm{BIASED}_{\mathrm{IV}}$ flag and a direct cross-reference
  to the macro\_IR analysis at the same $\Delta t$ for the operationally
  valid evidence.
\item Information gain ([[supplement-information-gain]]) for macro\_R
  at $\Delta t = 1\tau$ is decomposed per-direction, with the failed
  direction reported as ``algorithm-undetermined'' and the other
  five as data-informed.
\item Posterior intervals on the kinetic rate constants individually
  remain computable for macro\_R at $\Delta t = 1\tau$, but the
  joint interval on the failed parameter combination is reported as
  the prior interval (since the data carries no information in that
  direction under macro\_R).
\end{itemize}

\section*{10. Summary}
\label{sec:sc-summary}

\begin{itemize}
\item \textbf{Two decision axes.} The per-replicate safety
  categorization combines (i) spectral health of
  $\mathbf H_{\mathrm{lik}}$ via tolerances $\tau_{\mathrm{safe}}$
  and $\tau_{\mathrm{singular}}$ inherited from
  \texttt{lapack\_headers.h}, and (ii) the four-test pattern from
  [[supplement-algorithm-validation]] with thresholds
  $\tau_{\mathcal D} = 2\sqrt{p/(B-1)}$ and $\chi^2_{p,\,0.95}$.
\item \textbf{Four categories.} SAFE / MARGINAL / UNRELIABLE /
  BIASED, with the combined rule \eqref{eq:sc-combined-rule}. The
  promotion principle: worse spectral state cannot be redeemed by
  better test pattern.
\item \textbf{Fourth-regime sub-code.} The coincident failure mode
  $\mathrm{rank}(\mathbf G_{\mathrm{lik}}) < p$ and
  $\lambda_{\min}(\mathbf H_{\mathrm{lik}}) < 0$ in the same direction
  is emitted as $\mathrm{BIASED}_{\mathrm{IV}}$
  (\texttt{safety\_category} $= 4$) to distinguish from generic
  indefiniteness.
\item \textbf{Per-category reporting.} Each category maps to a
  reporting policy that gates which diagnostics are emitted, NaN-filled,
  or flagged in the bootstrap-samples binary output, and which
  replicates contribute to which cross-replicate aggregates.
\item \textbf{Cross-replicate aggregation.} Fraction-based rule
  \eqref{eq:sc-cell-aggregation} promotes per-replicate categories to
  cell-level categories. Default fractions $f_{\mathrm{safe}} = 0.90$,
  $f_{\mathrm{marginal}} = 0.75$, $f_{\mathrm{biased,unrel}} = 0.10$.
\item \textbf{Per-direction refinement.} Non-SAFE replicates are
  decomposed by parameter eigendirection of $\mathbf C_{\mathrm{post}}$,
  allowing the supplement table to identify which kinetic-parameter
  combinations are operationally recoverable and which are not.
\item \textbf{Empirical case.} macro\_R at $\Delta t = 1\tau$ is
  categorized as $\mathrm{BIASED}_{\mathrm{IV}}$ at the cell level
  (100\% of replicates exhibit the fourth-regime signature in one
  parameter direction). macro\_IR is SAFE everywhere.
\item \textbf{Production scope.} In production use where
  $\theta_{\mathrm{sim}}$ is unknown and the Wald test $T^2$ is
  unavailable, the BIASED-vs-UNRELIABLE distinction at the validation
  axis loses one piece of evidence, but the spectral axis and the
  three matrix tests remain operational. The categorization framework
  applies to real experimental data, not only to simulation studies.
\end{itemize}

\end{document}
