---
title: "Figure 2"
author: "Luciano Moffatt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

```{r helpers, include=FALSE}
# Generic mean ± 95% CI plot for d_mean (column: mean_NA, quantile_0.025, quantile_0.975).
# - x, color: column names (strings)
# - log_y: y, ymin, ymax become 10^column (then optionally scale_y_log10)
# - facets: a facet_*() expression
ci_plot <- function(d, x = "interval_in_tau", color = "component",
                    log_y = FALSE, log_x = TRUE, log_y_scale = FALSE,
                    hline = NULL, title = "", facets = NULL) {
  if (nrow(d) == 0) return(ggplot() + ggtitle(paste(title, "(no data)")))
  d2 <- d %>% mutate(
    .x    = .data[[x]],
    .col  = factor(.data[[color]]),
    .y    = if (log_y) 10^mean_NA else mean_NA,
    .ymin = if (log_y) 10^quantile_0.025 else quantile_0.025,
    .ymax = if (log_y) 10^quantile_0.975 else quantile_0.975
  )
  p <- ggplot(d2, aes(x = .x, y = .y, color = .col)) +
    geom_line() + geom_point() +
    geom_errorbar(aes(ymin = .ymin, ymax = .ymax)) +
    labs(x = x, color = color, title = title)
  if (log_x)       p <- p + scale_x_log10()
  if (log_y_scale) p <- p + scale_y_log10()
  if (!is.null(hline))  p <- p + geom_hline(yintercept = hline, color = "darkgreen")
  if (!is.null(facets)) p <- p + facets
  p
}

# Ratio plot from d_algo: <num_alg>_mean_NA / <den_alg>_mean_NA, with CI on numerator.
ratio_plot <- function(d, num_alg = "macro_MNR", den_alg = "macro_IR",
                       x = "interval_in_tau", color = "component",
                       facets = NULL, title = "", log_y = TRUE, hline = 1) {
  if (nrow(d) == 0) return(ggplot() + ggtitle(paste(title, "(no data)")))
  d2 <- d %>% mutate(
    .x   = .data[[x]],
    .col = factor(.data[[color]]),
    .num    = .data[[paste0(num_alg, "_mean_NA")]],
    .num_lo = .data[[paste0(num_alg, "_quantile_0.025")]],
    .num_hi = .data[[paste0(num_alg, "_quantile_0.975")]],
    .den    = .data[[paste0(den_alg, "_mean_NA")]]
  )
  p <- ggplot(d2, aes(x = .x, y = .num/.den, color = .col)) +
    geom_line() + geom_point() +
    geom_errorbar(aes(ymin = .num_lo/.den, ymax = .num_hi/.den)) +
    geom_hline(yintercept = hline, color = "darkgreen") +
    scale_x_log10() + labs(x = x, title = title)
  if (log_y)            p <- p + scale_y_log10()
  if (!is.null(facets)) p <- p + facets
  p
}

# Diag-only param names for a (component, algorithm)
params_for <- function(d, comp, alg = "macro_IR", diag_only = TRUE) {
  q <- d %>% filter(component == comp, algorithm == alg, param_name != "")
  if (diag_only) q <- q %>% filter(param_index == param_col)
  q %>% pull(param_name) %>% unique()
}

# Safe replacements for figure_2A helpers: tolerate missing mean_NA_bootstrap_count
.has_bc       <- function(d) "mean_NA_bootstrap_count" %in% names(d)
.bc_nonempty  <- function(d, p) {
  if (!.has_bc(d)) return(FALSE)
  nrow(drop_na(filter(d, param_name == p), mean_NA_bootstrap_count)) > 0
}

make_figure <- function(d_mean, select, type = "linear") {
  intercept <- c(linear = 0, log = 1)
  for (c in select) {
    dtemp <- d_mean %>% filter(component == c,
                                is.na(value_row) | is.na(value_col) | value_row == value_col)
    for (p in unique(dtemp$param_name)) {
      if (p == "" && .bc_nonempty(dtemp, p)) {
        print(
          dtemp %>% filter(param_name == p) %>% ggplot() +
            geom_point(aes(x = algorithm, y = mean_NA_bootstrap_count, color = factor(algorithm))) +
            facet_grid(Num_ch ~ noise_in_conductance_tau + interval_in_tau, scales = "free")
        )
      } else {
        fig <- dtemp %>% filter(param_name == p) %>% ggplot() +
          geom_point(aes(x = algorithm, y = mean_NA, color = factor(algorithm))) +
          geom_errorbar(aes(x = algorithm, ymin = quantile_0.025, ymax = quantile_0.975,
                            color = factor(algorithm))) +
          geom_hline(yintercept = intercept[type], color = "darkgreen") +
          ggtitle(paste(c, p)) +
          facet_grid(Num_ch ~ noise_in_conductance_tau + interval_in_tau, scales = "free")
        if (type != "linear") fig <- fig + scale_y_log10()
        print(fig)
      }
    }
  }
}

make_figure_per_Nch <- function(d_mean, select, type = "linear") {
  intercept <- c(linear = 0, log = 1)
  Nch <- unique(d_mean$Num_ch)
  for (c in unique(d_mean$component)) {
    dtemp <- d_mean %>% filter(component == c,
                                is.na(value_row) | is.na(value_col) | value_row == value_col)
    params <- unique(dtemp$param_name)
    for (num_ch in Nch) {
      figs <- list()
      for (p in params) {
        if (p == "" && .bc_nonempty(dtemp, p)) {
          figs[[p]] <- dtemp %>% filter(param_name == p) %>% ggplot() +
            geom_point(aes(x = algorithm, y = mean_NA_bootstrap_count, color = factor(algorithm))) +
            facet_grid(Num_ch ~ noise_in_conductance_tau + interval_in_tau, scales = "free") +
            ggtitle(paste(c, p, "Nch =", num_ch))
        } else {
          fig <- dtemp %>% filter(Num_ch == num_ch, param_name == p) %>% ggplot() +
            geom_point(aes(x = algorithm, y = mean_NA, color = factor(algorithm))) +
            geom_errorbar(aes(x = algorithm, ymin = quantile_0.025, ymax = quantile_0.975,
                              color = factor(algorithm))) +
            geom_hline(yintercept = intercept[type], color = "darkgreen") +
            ggtitle(paste(c, p, "Nch =", num_ch)) +
            facet_grid(interval_in_tau ~ noise_in_conductance_tau + param_name, scales = "free")
          if (type != "linear") fig <- fig + scale_y_log10()
          figs[[p]] <- fig
        }
      }
      if (length(figs)) print(wrap_plots(figs) + plot_layout(guides = "collect"))
    }
  }
}
```

```{r}

filename=c(
  #"data/figure_2_basic_diagnostics__0420_1",
  #"data/figure_2_basic_diagnostics__0420_2_001",
  #"data/figure_2_basic_diagnostics__0421_3_001",
  #"data/figure_2_basic_diagnostics__0422_4_001",
  "data/figure_2_basic_diagnostics__0509_005"


  )


```


```{r}
d<- lapply(paste0(filename, ".csv"), read.csv) %>%
  bind_rows()
```


```{r}
d<- d %>%
  select(where(~ !all(is.na(.))))
```

```{r}
d%>%select(component_path)%>%distinct()%>%pull()

```


```{r}
d<-d%>%mutate(component= component_path %>%
           str_replace("^Probit_statistics_", "") )


```

```{r}
d%>%select(component)%>%distinct()%>%pull()

```

```{r}
drop_components <- c("Probit_statistics", "count_r_std", "count_dlogL")

d_mean <- d %>% filter(!component %in% drop_components) %>%
  pivot_wider(names_from = c(probit, quantile_level), values_from = value)

d_algo <- d %>% filter(!component %in% drop_components) %>%
  pivot_wider(names_from = c(algorithm, probit, quantile_level), values_from = value)
```

# Report_integral_r_std

```{r, fig.width=16, fig.height=16}
ci_plot(
  d_mean %>% filter(component == "Report_integral_r_std",
                    statistic == "integral_correlation_lag",
                    algorithm %in% c("macro_IR", "macro_IRT", "macro_MR", "macro_R")),
  color = "Num_ch", hline = 1, title = "Report_integral_r_std",
  facets = facet_grid(Num_ch ~ algorithm + noise_in_conductance_tau, scales = "free")
)
```

# Distortion_Induced_Bias by algorithm (1/Num_ch/interval_in_tau)

```{r, fig.width=8, fig.height=8}
for (a in unique(d_mean$algorithm)) {
  dd <- d_mean %>% filter(component == "Distortion_Induced_Bias",
                          algorithm == a, param_name != "") %>%
        mutate(x_inv = 1 / interval_in_tau / Num_ch)
  if (nrow(dd) == 0) next
  print(ci_plot(dd, x = "x_inv", color = "Num_ch", log_y = TRUE, hline = 1,
                title = a,
                facets = facet_wrap(~ param_name, scales = "free")))
}
```

# Report_integral_dlogL ratio MNR/IR by noise level (3-panel)

```{r, fig.width=16, fig.height=8}
mk_ratio <- function(noise) {
  ratio_plot(
    d_algo %>% filter(noise_in_conductance_tau == noise,
                      component == "Report_integral_dlogL",
                      param_name %in% c("on_on", "off_off"),
                      Num_ch * interval_in_tau > 10,
                      param_index == param_col),
    facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
    title = paste("noise =", noise)
  )
}
noise_levels <- sort(unique(d_algo$noise_in_conductance_tau))
Reduce(`|`, lapply(noise_levels, mk_ratio)) + plot_layout(guides = "collect")
```

# Distortion_Induced_Bias and Distortion_Corrected_Covariance ratios

```{r, fig.width=16, fig.height=8}
panel_bias <- function(noise) {
  p <- ratio_plot(
    d_algo %>% filter(noise_in_conductance_tau == noise,
                      component == "Distortion_Induced_Bias",
                      param_name %in% c("on", "off"),
                      Num_ch * interval_in_tau > 10),
    facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
    title = paste("Bias diff noise =", noise), log_y = FALSE, hline = 0
  )
  if ("data" %in% names(p) && nrow(p$data) > 0) p <- p + aes(y = abs(.num) - abs(.den))
  p
}
panel_cov <- function(noise) ratio_plot(
  d_algo %>% filter(noise_in_conductance_tau == noise,
                    component == "Distortion_Corrected_Covariance",
                    param_name %in% c("on_on", "off_off"),
                    Num_ch * interval_in_tau > 10,
                    param_index == param_col),
  facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
  title = paste("Cov ratio noise =", noise)
)
(Reduce(`|`, lapply(noise_levels, function(n) panel_bias(n) | panel_cov(n)))) +
  plot_layout(guides = "collect")
```

# Fisher_Covariance for macro_MNR (diag, on_on/off_off, etc.)

```{r, fig.width=16, fig.height=8}
fc_panel <- function(params) ci_plot(
  d_mean %>% filter(component %in% c("Distortion_Corrected_Covariance", "Fisher_Covariance"),
                    algorithm == "macro_MNR",
                    param_name %in% params,
                    Num_ch * interval_in_tau > 10,
                    param_index == param_col),
  log_y = FALSE, log_y_scale = TRUE,
  facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
  title = "macro_MNR"
)
(fc_panel(c("on_on", "off_off"))
  | fc_panel(c("Num_ch_mean_Num_ch_mean", "unitary_current_unitary_current"))
  | fc_panel("Current_Noise_Current_Noise")) + plot_layout(guides = "collect")
```

# Same panels for macro_IR

```{r, fig.width=20, fig.height=8}
fc_panel_ir <- function(params, title = "macro_IR") ci_plot(
  d_mean %>% filter(component %in% c("Distortion_Corrected_Covariance", "Fisher_Covariance"),
                    algorithm == "macro_IR",
                    param_name %in% params,
                    Num_ch * interval_in_tau > 10,
                    param_index == param_col),
  log_y_scale = TRUE,
  facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
  title = title
)
(fc_panel_ir(c("on_on", "off_off"))
  | fc_panel_ir(c("Num_ch_mean_Num_ch_mean", "unitary_current_unitary_current"))
  | fc_panel_ir("Current_Noise_Current_Noise", "macro_NR")
) + plot_layout(guides = "collect")
```

# Report_integral_dlogL and Correlation_Distortion_Matrix for macro_IR

```{r, fig.width=20, fig.height=8}
rid_panel <- function(params, title = "macro_IR") ci_plot(
  d_mean %>% filter(component %in% c("Report_integral_dlogL", "Correlation_Distortion_Matrix"),
                    algorithm == "macro_IR",
                    param_name %in% params,
                    Num_ch * interval_in_tau > 10,
                    param_index == param_col),
  log_y_scale = TRUE,
  facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
  title = title
)
(rid_panel(c("on_on", "off_off"))
  | rid_panel(c("Num_ch_mean_Num_ch_mean", "unitary_current_unitary_current"))
  | rid_panel("Current_Noise_Current_Noise", "macro_NR")
) + plot_layout(guides = "collect")
```

# Distortion_Corrected_Covariance + Fisher_Covariance combined IR/MNR

```{r, fig.width=20, fig.height=8}
combo_panel <- function(params, title = "macro_IR/MNR") {
  d <- d_mean %>% filter(component %in% c("Distortion_Corrected_Covariance", "Fisher_Covariance"),
                          algorithm %in% c("macro_IR", "macro_MNR"),
                          param_name %in% params,
                          Num_ch * interval_in_tau > 10,
                          param_index == param_col) %>%
        mutate(combo = interaction(algorithm, component))
  ci_plot(d, color = "combo", log_y_scale = TRUE,
          facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
          title = title)
}
(combo_panel(c("on_on", "off_off"))
  | combo_panel(c("Num_ch_mean_Num_ch_mean", "unitary_current_unitary_current"))
  | combo_panel("Current_Noise_Current_Noise", "macro_NR")
) + plot_layout(guides = "collect")
```

# Distortion_Induced_Bias scans

```{r, fig.width=8, fig.height=8}
dib_scan <- function(exclude_alg, hi = TRUE) {
  thresh <- if (hi) quote(Num_ch * interval_in_tau > 10) else quote(Num_ch * interval_in_tau <= 10)
  d_mean %>% filter(component == "Distortion_Induced_Bias",
                    !algorithm %in% exclude_alg,
                    !param_name %in% c("Current_Baseline", "Current_Noise"),
                    param_name != "",
                    !!thresh) %>%
    ci_plot(color = "algorithm", log_y = TRUE, hline = 1,
            facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
            title = paste0("excl=", paste(exclude_alg, collapse = ",")))
}
print(dib_scan(c("macro_NR", "macro_MNR", "macro_IRT", "macro_R")))
print(dib_scan(c("macro_NR", "macro_MNR", "macro_IRT")))

for (n in unique(d_mean$noise_in_conductance_tau)) {
  dd <- d_mean %>% filter(noise_in_conductance_tau == n,
                          component == "Distortion_Induced_Bias",
                          !algorithm %in% c("macro_NR", "macro_MNR", "macro_IRT", "macro_R"),
                          !param_name %in% c("Current_Baseline", "Current_Noise"),
                          param_name != "",
                          Num_ch * interval_in_tau <= 10)
  if (nrow(dd) == 0) next
  print(ci_plot(dd, color = "algorithm", log_y = TRUE, hline = 1,
                facets = facet_grid(Num_ch ~ param_name, scales = "free"),
                title = paste("noise:", n)))
}
```

# Information_Distortion_Matrix scans

```{r, fig.width=8, fig.height=8}
idm_scan <- function(exclude_alg, hi = TRUE) {
  thresh <- if (hi) quote(Num_ch * interval_in_tau > 10) else quote(Num_ch * interval_in_tau <= 10)
  d_mean %>% filter(component == "Information_Distortion_Matrix",
                    !algorithm %in% exclude_alg,
                    !param_name %in% c("Current_Baseline_Current_Baseline",
                                       "Current_Noise_Current_Noise"),
                    param_name != "",
                    param_index == param_col,
                    !!thresh) %>%
    ci_plot(color = "algorithm", hline = 1,
            facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
            title = paste0("excl=", paste(exclude_alg, collapse = ",")))
}
print(idm_scan(c("macro_NR", "macro_MNR", "macro_MR", "macro_R")))
print(idm_scan(c("macro_NR", "macro_MNR", "macro_IRT", "macro_R")))
print(idm_scan(c("macro_NR", "macro_MNR", "macro_IRT")))
print(idm_scan(c("macro_NR", "macro_MNR", "macro_IRT", "macro_R"), hi = FALSE))
```

# Distortion_Corrected_Covariance scans

```{r, fig.width=16, fig.height=8}
dcc_scan <- function(exclude_alg, hi = TRUE,
                     exclude_param = c("Current_Baseline_Current_Baseline",
                                       "Current_Noise_Current_Noise")) {
  thresh <- if (hi) quote(Num_ch * interval_in_tau > 10) else quote(Num_ch * interval_in_tau <= 10)
  d_mean %>% filter(component == "Distortion_Corrected_Covariance",
                    !algorithm %in% exclude_alg,
                    !param_name %in% exclude_param,
                    param_name != "",
                    param_index == param_col,
                    !!thresh) %>%
    ci_plot(color = "algorithm",
            facets = facet_grid(Num_ch ~ param_name + noise_in_conductance_tau, scales = "free"),
            title = paste0("excl=", paste(exclude_alg, collapse = ",")))
}
print(dcc_scan(c("macro_NR", "macro_MNR", "macro_MR", "macro_R")))
print(dcc_scan(c("macro_NR", "macro_MNR", "macro_IRT", "macro_R")))
print(dcc_scan(c("macro_NR", "macro_MNR", "macro_IRT", "macro_MR")))
print(dcc_scan(c("macro_NR", "macro_MNR", "macro_IRT"), hi = FALSE))
```

# log_Det_Distortion_Corrected_Covariance

```{r}
ci_plot(
  d_mean %>% filter(component == "log_Det_Distortion_Corrected_Covariance",
                    statistic == "value") %>%
             mutate(inv_T = 1 / interval_in_tau),
  x = "inv_T", color = "Num_ch",
  facets = facet_wrap(algorithm ~ noise_in_conductance_tau, scales = "free")
)
```

# Distortion_Induced_Bias by param

```{r, fig.width=20, fig.height=20}
for (p in params_for(d_mean, "Distortion_Induced_Bias", diag_only = FALSE)) {
  dd <- d_mean %>% filter(component == "Distortion_Induced_Bias", param_name == p) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "noise_in_conductance_tau",
                log_y = TRUE, hline = 1,
                facets = facet_wrap(algorithm ~ Num_ch, scales = "free"),
                title = p))
}
```

# Information_Distortion_Matrix by param (two scalings)

```{r, fig.width=8, fig.height=8}
idm_params <- params_for(d_mean, "Information_Distortion_Matrix")
for (p in idm_params) {
  dd <- d_mean %>% filter(component == "Information_Distortion_Matrix", param_name == p) %>%
                   mutate(inv_T_sN = 1 / interval_in_tau / sqrt(Num_ch))
  print(ci_plot(dd, x = "inv_T_sN", color = "Num_ch", hline = 1,
                facets = facet_wrap(~ algorithm, scales = "free"),
                title = p))
}
for (p in idm_params) {
  dd <- d_mean %>% filter(component == "Information_Distortion_Matrix", param_name == p) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "Num_ch", log_y_scale = TRUE, hline = 1,
                facets = facet_wrap(algorithm ~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Report_integral_dlogL by param

```{r}
for (p in idm_params) {
  dd <- d_mean %>% filter(component == "Report_integral_dlogL", param_name == p) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "Num_ch", hline = 1,
                facets = facet_wrap(algorithm ~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Distortion_Corrected_Covariance + Fisher_Covariance by (algo, param)

```{r, fig.width=8, fig.height=8}
dcc_params <- params_for(d_mean, "Distortion_Corrected_Covariance")
for (a in unique(d_mean$algorithm)) {
  for (p in dcc_params) {
    dd <- d_mean %>% filter(component %in% c("Distortion_Corrected_Covariance", "Fisher_Covariance"),
                            param_name == p, algorithm == a) %>%
                     mutate(inv_T = 1 / interval_in_tau)
    print(ci_plot(dd, x = "inv_T", log_y_scale = TRUE,
                  facets = facet_wrap(~ Num_ch, scales = "free"),
                  title = paste(a, p)))
  }
}
```

# Correlation_Distortion_Matrix by param

```{r}
for (p in idm_params) {
  dd <- d_mean %>% filter(component == "Correlation_Distortion_Matrix", param_name == p) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "Num_ch", log_y_scale = TRUE, hline = 1,
                facets = facet_wrap(algorithm ~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Information_Distortion_Reconstituted

```{r}
recon_params <- d_mean %>%
  filter(component == "Information_Distortion_Reconstituted",
         algorithm == "macro_MR",
         Num_ch * interval_in_tau > 9,
         is.na(value_row) | is.na(value_col) | value_row == value_col) %>%
  pull(param_name) %>% unique()
for (p in recon_params) {
  dd <- d_mean %>% filter(component == "Information_Distortion_Reconstituted",
                          param_name == p, algorithm == "macro_IR",
                          Num_ch * interval_in_tau > 10) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "Num_ch", log_y_scale = TRUE, hline = 1,
                facets = facet_wrap(~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Distortion_Corrected_Covariance + Fisher_Covariance, all algos

```{r, fig.width=20, fig.height=20}
dcc_params_irv <- d_mean %>%
  filter(component == "Distortion_Corrected_Covariance",
         algorithm == "macro_IRV",
         Num_ch * interval_in_tau > 9,
         is.na(value_row) | is.na(value_col) | value_row == value_col) %>%
  pull(param_name) %>% unique()
for (p in dcc_params_irv) {
  dd <- d_mean %>% filter(component %in% c("Distortion_Corrected_Covariance", "Fisher_Covariance"),
                          param_name == p) %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", log_y_scale = TRUE,
                facets = facet_grid(Num_ch ~ algorithm, scales = "free"),
                title = p))
}
```

# Fisher_Covariance by param (macro_IRV)

```{r}
fc_params <- d_mean %>%
  filter(component == "Fisher_Covariance", algorithm == "macro_IRV",
         Num_ch * interval_in_tau > 9,
         is.na(value_row) | is.na(value_col) | value_row == value_col) %>%
  pull(param_name) %>% unique()
for (p in fc_params) {
  dd <- d_mean %>% filter(component == "Fisher_Covariance",
                          param_name == p, algorithm == "macro_IRV") %>%
                   mutate(inv_T = 1 / interval_in_tau)
  print(ci_plot(dd, x = "inv_T", color = "Num_ch", log_y_scale = TRUE, hline = 1,
                facets = facet_wrap(~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Distortion_Corrected_Covariance for MRV/IRV/IRTV

```{r, fig.height=10, fig.width=10}
mrv_params <- d_mean %>%
  filter(component == "Distortion_Corrected_Covariance", algorithm == "macro_MRV",
         Num_ch * interval_in_tau > 9,
         is.na(value_row) | is.na(value_col) | value_row == value_col) %>%
  pull(param_name) %>% unique()
for (p in mrv_params) {
  dd <- d_mean %>% filter(component == "Distortion_Corrected_Covariance",
                          param_name == p,
                          algorithm %in% c("macro_MRV", "macro_IRV", "macro_IRTV"),
                          Num_ch * interval_in_tau > 9) %>%
                   mutate(inv_NchT = 1 / Num_ch / interval_in_tau)
  print(ci_plot(dd, x = "inv_NchT", color = "algorithm", log_y_scale = TRUE,
                facets = facet_wrap(Num_ch ~ noise_in_conductance_tau, scales = "free"),
                title = p))
}
```

# Cross-algorithm summaries via make_figure

```{r, fig.width=20, fig.height=10}
make_figure(d_mean %>% filter(!algorithm %in% c("macro_MNR", "macro_NR")),
            "Distortion_Induced_Bias")
```

```{r, fig.width=20, fig.height=10}
make_figure_per_Nch(d_mean %>% filter(algorithm == "macro_IR"),
                    "Distortion_Induced_Bias")
```

```{r, fig.width=20, fig.height=10}
make_figure(d_mean %>% filter(algorithm != "macro_MNRV"),
            "Moment_statistics_Sum_dlogL")
```

```{r, fig.width=20, fig.height=20}
make_figure(d_mean, "Moment_statistics_Sum_trust_coefficient")
```

```{r, fig.width=10, fig.height=10}
make_figure(d_mean %>% filter(algorithm != "macro_MNRV"),
            "Information_Distortion_Matrix", type = "log")
```

```{r, fig.width=20, fig.height=10}
make_figure(d_mean, "Distortion_Corrected_Covariance", type = "log")
```

```{r, fig.width=20, fig.height=10}
make_figure(d_mean %>% filter(algorithm != "macro_MNRV"),
            c("Moment_statistics_Sum_r_std", "Moment_statistics_Sum_logL"))
```

```{r}
make_figure(d_mean, "Moment_statistics_Sum_r2_std", type = "log")
```

```{r}
make_figure(
  d_mean %>% filter(noise_in_conductance_tau == 0.1, Num_ch == 1, interval_in_tau == 0.01),
  d_mean %>% distinct(component) %>% pull()
)
```
