\documentclass[9pt]{elife}

\usepackage{amsmath,amssymb,mathtools}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage{bm}

% TITLE — candidate, in the "characterize validity" frame (replaces the old "exact/unbiased" titles).
% Alternative: "Where time-averaged likelihoods hold and where they break: macroscopic kinetic inference"
\title{Validity limits of time-averaged likelihoods for macroscopic kinetic inference}

\author[1*]{Luciano Moffatt}
\affil[1]{Instituto de Qu\'imica F\'isica de los Materiales, Medio Ambiente y Energ\'ia (INQUIMAE), CONICET; Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires, Ciudad de Buenos Aires, C1428EHA, Argentina}
\corr{lmoffatt@qi.fcen.uba.ar}

\begin{document}
\maketitle

\begin{abstract}
Time-averaged measurements pervade biology, and the likelihoods used to interpret them are approximations whose reliability is rarely tested. We ask where such an approximation can be trusted and how it fails when it cannot. For macroscopic ion-channel currents, we simulate a two-state channel under a concentration jump and use the exact simulation as ground truth to judge a family of Gaussian likelihood approximations (non-recursive, recursive, interval-averaged, and their combinations). We measure faithfulness with likelihood-level diagnostics: standardized residuals, the score, and an information-distortion matrix that compares the covariance of the score with the model's own Fisher information. We then map how each approximation deforms inference as a function of the number of channels, the acquisition interval, and the instrumental noise. Only the interval-averaged recursive likelihood (MacroIR) stays calibrated across the practical regime; the others lose or misallocate information in ways that trace to the two Gaussian approximations they make, one on channel occupancy and one on the single-interval signal. We locate the multinomial, telegraphic, and Gaussian regimes where the approximations degrade, and we show the distortion carries a concrete consequence: it maps to a correction of the Bayesian evidence that ranks competing mechanisms. The result is a validity map for time-averaged likelihoods and a calibrated default, MacroIR, for macroscopic recordings.
\end{abstract}

% Impact statement (eLife requires one, <= 40 words):
% A validity map shows where time-averaged likelihood approximations for ion-channel currents can be trusted and where they distort inference, and identifies MacroIR as the one that stays calibrated across the practical regime.

% ================= SKELETON: titles only below. Fill-hints in % comments. =================
% One file per top-level section under sections/, so that one section = one owner = no merge
% conflicts (01_writing_plan.md, T-1). The fill-hints live in the section files, with the
% section they brief. Add nothing here but \input lines.

\input{sections/01_introduction}
\input{sections/02_theory}
\input{sections/03_diagnostics}
\input{sections/04_results}
\input{sections/05_discussion}
\input{sections/06_methods}
\input{sections/07_backmatter}

% \nocite{*}
\bibliography{biblio}

\end{document}
