\section*{The macroscopic interval likelihood and its approximations}
% Owner: theory_plan.md. Presents the five algorithms as one object with two knobs
% (recursion x window), with the endpoint ladder as the presentation. Does NOT restate:
% scope (02_decision_log), the novelty claim (D-3, unsettled), diagnostic definitions
% (diagnostics_plan / 03_metrics_diagnostics), or run values and the noise parameter
% (methods_plan). The index-level kernels for the boundary moments (E_2, E_3, the
% Kronecker-sum construction) live in Methods and the Supplement; this section names them.

A macroscopic current is the summed signal of a population of ion channels, each of
which behaves as an independent continuous-time Markov chain over a small set of
conformational states. We write $N_{\mathrm{ch}}$ for the number of channels and $K$ for
the number of kinetic states of one channel ($K=2$ throughout this paper, carried % src: 02_decision_log.md (two-state scope)
symbolically where it costs nothing). Each channel is governed by a generator (rate)
matrix $\mathbf{Q}\in\mathbb{R}^{K\times K}$, whose transition matrix over an elapsed time
$t$ is the row-stochastic $\mathbf{P}(t)=e^{\mathbf{Q}t}$ with entries $P_{i\to j}(t)$. A
channel in state $k$ carries a conductance $\gamma_k$, and we collect these in the vector
$\bm{\gamma}\in\mathbb{R}^{K}$.

The likelihood we need is the probability of a recorded sample given the kinetic
parameters and the data seen so far. Its exact form would require the distribution of the
interval-averaged conductance of the whole population, which has no closed expression.
Every practical likelihood replaces that distribution with a Gaussian, and the members of
the family studied here differ only in what the Gaussian is conditioned on. The rest of
this section unpacks that sentence.

\subsection*{The observable is an interval average}

A recorded sample is the current averaged over the acquisition window rather than sampled at an instant. The acquisition
electronics integrate the current across the sampling window $\Delta$ and pass it through
an anti-aliasing filter before it is written down, so the quantity a likelihood must
describe is the windowed average
\begin{equation}
  \bar{y}_t \;=\; \frac{1}{\Delta}\int_{t-\Delta}^{t} y(\tau)\,\mathrm{d}\tau \;+\; \eta_t,
  \label{eq:obs-avg}
\end{equation}
where $y(\tau)$ is the instantaneous population current and $\eta_t$ is the instrumental
noise accumulated over the window. This interval average is the physical observable. When
$\Delta$ is much shorter than the fastest relaxation of $\mathbf{Q}$ the average sits close
to the instantaneous current and the distinction is negligible; when $\Delta$ is comparable
to or longer than that relaxation the two diverge, and a likelihood built on instantaneous
sampling describes the wrong random variable.

\subsection*{Two Gaussian approximations}

Each member of the family makes two closures, on two different objects.

\begin{tcolorbox}[title={The two Gaussian approximations}, colback=gray!5, colframe=gray!55]
\textbf{Occupancy closure.} The joint distribution of the channel occupancies is exactly
multinomial when the channels are independent, which is also the maximum-entropy
distribution consistent with the mean occupancies. It is replaced by a multivariate
Gaussian with mean $\bm{\mu}$ and covariance $\bm{\Sigma}$, valid for large
$N_{\mathrm{ch}}$ by the central limit theorem and degrading as the channel count falls.

\smallskip
\textbf{Interval-signal closure.} The distribution of the interval-averaged conductance
over one window is a telegraph-like object with no closed form. It is replaced by a
Gaussian, valid when the window spans many transitions or when instrumental noise
dominates, and degrading as the window shrinks below the relaxation time.
\end{tcolorbox}

The higher moments these closures discard do not vanish. They reappear in the data as
temporal correlation that the likelihood does not predict, which is what the correlation
term of the information-distortion decomposition measures (Diagnostics). Three regimes
follow, and we name them here so the maps in the Results have vocabulary: the
\emph{multinomial} regime (few channels, where the occupancy closure fails), the
\emph{telegraphic} regime (windows far shorter than the relaxation time, where the
interval-signal closure fails), and the \emph{Gaussian} regime, where both closures hold
because the channel count is high enough for the occupancy Gaussian while the window is long
enough, or the instrumental noise large enough, for the interval-signal Gaussian. Any Gaussian likelihood of this
kind must fail somewhere. The aim of this paper is to show that the failures are the
predicted degradation of these two closures rather than defects of a particular algorithm.

\subsection*{The family as a coordinate system}

The family has two knobs. One axis is the window: how much of the acquisition interval the
conductance is conditioned on. The other is recursion: whether the occupancy distribution
is conditioned on the data as the recording proceeds.

Along the recursion axis, a recursive (prequential) member factorizes the likelihood as
$\prod_t q_t(\bar{y}_t \mid \bar{y}_{1:t-1})$, so the occupancy distribution carried into
each window is the posterior left by the previous windows and the hidden state is
conditioned on the data. A non-recursive (open-loop) member factorizes as
$\prod_t m_t(\bar{y}_t)$, with each window's predictive propagated from the initial
condition and never conditioned on the observed trace.

Along the window axis the members form a ladder, indexed by how many of the interval's
endpoints the conductance is conditioned on (Table~\ref{tab:ladder}).

\begin{table}[htbp]
\centering
\begin{tabular}{@{}lll@{}}
\toprule
Conditioned on & Members & Conductance model \\
\midrule
no endpoints & \texttt{NR}, \texttt{R} & instantaneous; the acquisition averaging is ignored \\
one endpoint (the start) & \texttt{NMR}, \texttt{MR} & interval-mean conductance given the start state \\
two endpoints (the boundary) & \texttt{IR} & interval-mean conductance given both boundary states; interior marginalized \\
the full trajectory & (exact) & intractable; supplied by the stochastic simulation as ground truth \\
\bottomrule
\end{tabular}
\caption{The window axis as an endpoint ladder. Crossing it with the recursion axis gives
the implemented grid: \texttt{NR} (non-recursive, $\mathrm{av}=0$), \texttt{R} (recursive,
$\mathrm{av}=0$), \texttt{NMR} (non-recursive, $\mathrm{av}=1$), \texttt{MR} (recursive,
$\mathrm{av}=1$), \texttt{IR} (recursive, $\mathrm{av}=2$), where $\mathrm{av}$ counts the
conditioned endpoints. MacroIR is \texttt{IR}.} % src: nomenclature.md (family grid, av flag)
\label{tab:ladder}
\end{table}

Two things follow from the ladder. First, MacroIR occupies the top rung below the
intractable exact likelihood, so its standing in the family is structural and set before
any measurement. Second, the lower rungs are established likelihoods rather than
constructed foils: the non-recursive instantaneous member is the independent-interval
likelihood of \citet{milescu2005maximum}, and the recursive instantaneous member is the
macroscopic recursive filter of \citet{moffatt2007estimation} and the generalized filter
of \citet{munch2022bayesian}. The family is one coordinate system that contains these
methods as special cases.

\subsection*{The boundary state}

A \emph{boundary state} is the pair $(i_0, i_t)$ of a channel's states at the two ends of
an acquisition interval. Chaining intervals, the end state of one interval is the start
state of the next, so there is a single state variable per junction: the filter conditions
on the states at the interval boundaries and marginalizes the trajectory between them
analytically. This is the same device as static condensation, or the spatial Markov
property, applied on the time axis. It is not a transition state in the mechanistic sense,
and we keep the word \emph{transition} away from it. Because the joint dynamics of the two
endpoints are generated by the Kronecker sum
$\mathbf{Q}\oplus\mathbf{Q}=\mathbf{Q}\otimes\mathbf{I}+\mathbf{I}\otimes\mathbf{Q}$, whose
propagator factorizes as $e^{(\mathbf{Q}\oplus\mathbf{Q})t}=e^{\mathbf{Q}t}\otimes
e^{\mathbf{Q}t}$, the moments over the $K^2$ boundary pairs collapse to bilinear forms in
the $K$-dimensional state space and no explicit $K^2$ object is formed (Methods).

The ladder needs two conductance objects, both built from the boundary-conditioned
single-channel moments $\overline{\Gamma}_{i_0\to i_t}$ (the mean interval-averaged
conductance of a channel that starts in $i_0$ and ends in $i_t$) and
$\overline{V}_{i_0\to i_t}$ (its variance), which are computed from $\mathbf{Q}$ and the
recording filter (Methods). The start-conditioned mean conductance is the $K$-vector
\begin{equation}
  (\bar{\gamma}_0)_{i_0} \;=\; \sum_{i_t} P_{i_0\to i_t}(\Delta)\,\overline{\Gamma}_{i_0\to i_t},
  \label{eq:gammabar}
\end{equation}
which conditions on the state at the interval's start alone and is the object used at
$\mathrm{av}=1$. The boundary-conditioned mean conductance is the same average retained on
both endpoints, so it stays a $K\times K$ object rather than collapsing to a vector, and is
the object used at $\mathrm{av}=2$.

\subsection*{The interval update}

Over a window of length $\Delta$, the occupancy mean and covariance propagate as
\begin{align}
  \bm{\mu}^{\mathrm{prop}} &= \mathbf{P}(\Delta)^\top \bm{\mu}_0, \label{eq:mu-prop}\\
  \bm{\Sigma}^{\mathrm{prop}} &= \mathbf{P}(\Delta)^\top\bigl(\bm{\Sigma}_0-\mathrm{diag}(\bm{\mu}_0)\bigr)\mathbf{P}(\Delta) + \mathrm{diag}(\bm{\mu}^{\mathrm{prop}}), \label{eq:sig-prop}
\end{align}
where $(\bm{\mu}_0, \bm{\Sigma}_0)$ is the occupancy Gaussian at the start of the window.
The predicted mean of the interval-averaged current is
\begin{equation}
  \bar{y}^{\mathrm{pred}} \;=\; N_{\mathrm{ch}}\,\bm{\mu}_0^\top \bar{\bm{\gamma}}_0,
  \label{eq:ypred}
\end{equation}
and its predicted variance decomposes into three independent contributions,
\begin{equation}
  \sigma^2_{\mathrm{pred}} \;=\; \epsilon^2_{\Delta}
  \;+\; N_{\mathrm{ch}}\,\widetilde{\bm{\gamma}^\top\bm{\Sigma}\bm{\gamma}}
  \;+\; N_{\mathrm{ch}}\,\bm{\mu}_0^\top\bm{\sigma}^2_{\bar{\gamma}_0}.
  \label{eq:pred-var}
\end{equation}
The first term $\epsilon^2_{\Delta}$ is the instrumental noise over the window, which
carries the acquisition averaging and decreases with $\Delta$; its parametric form is given
in Methods. The second term is the population gating variance, the fluctuation of the window
current transmitted through the occupancy covariance: $\widetilde{\bm{\gamma}^\top\bm{\Sigma}\bm{\gamma}}$
is the boundary covariance of the window collapsed onto the conductance, computed without
ever forming the $K^2$ boundary matrix. The third term is the within-window per-channel
variance, with $\bm{\sigma}^2_{\bar{\gamma}_0}$ the start-conditioned interval variance
built from $\overline{V}_{i_0\to i_t}$.

Writing $\mathbf{g}=\mathrm{Cov}(\text{end-of-window occupancy}, \bar{y})/N_{\mathrm{ch}}$
for the per-channel cross-covariance between the interval current and the occupancy at the
end of the window, and $\delta = \bar{y}^{\mathrm{obs}} - \bar{y}^{\mathrm{pred}}$ for the
innovation, the Gaussian update of the occupancy is the scalar Kalman correction
\begin{align}
  \bm{\mu}^{\mathrm{post}} &= \bm{\mu}^{\mathrm{prop}} + \frac{\mathbf{g}}{\sigma^2_{\mathrm{pred}}}\,\delta, \label{eq:mu-post}\\
  \bm{\Sigma}^{\mathrm{post}} &= \bm{\Sigma}^{\mathrm{prop}} - \frac{N_{\mathrm{ch}}\,\mathbf{g}\mathbf{g}^\top}{\sigma^2_{\mathrm{pred}}}. \label{eq:sig-post}
\end{align}
The factor $N_{\mathrm{ch}}$ enters the covariance downdate but not the mean correction
because $\bm{\Sigma}$ is carried on the population scale, the same $N_{\mathrm{ch}}$ that
multiplies the gating variance in Eq.~\ref{eq:pred-var}, whereas $\bm{\mu}$ and the gain
$\mathbf{g}$ are per-channel objects, so the rank-one correction to $\bm{\Sigma}$ must be
rescaled by $N_{\mathrm{ch}}$ to sit on its scale while the correction to $\bm{\mu}$ needs
no such factor.
In a recursive member the posterior pair $(\bm{\mu}^{\mathrm{post}}, \bm{\Sigma}^{\mathrm{post}})$
becomes the prior for the next window; in a non-recursive member it is not carried forward,
and each window's predictive is propagated from the initial condition alone. Reusing the
posterior as the next prior is itself a closure: it is exact only while the end marginal is
a sufficient summary of the boundary posterior, which the Gaussian projection does not
guarantee. For this reason the recursive members' Gaussian information is a model-based
quantity and is not claimed to be the exact information of the process; how far it departs
from the truth is what the Diagnostics measure.

The single algebraic difference within the family sits in the conductance object of
Eq.~\ref{eq:pred-var}. MacroIR conditions the interval-averaged conductance on both
boundary states; MacroMR conditions it on the start state alone, using $\bar{\bm{\gamma}}_0$
of Eq.~\ref{eq:gammabar} in its place. This drops the boundary cross-covariance term
\begin{equation}
  N_{\mathrm{ch}}\,\widetilde{\bm{\gamma}^\top\bm{\Sigma}\bm{\gamma}}
  \label{eq:mr-ir}
\end{equation}
that MacroIR retains, and it is the only place the two members diverge. Its consequence for
the predicted variance across the regime grid is characterized in the Results.

\subsection*{The top of the ladder}

The exact likelihood conditions on the full trajectory inside each window and is
intractable, which is why the two closures above are needed at all. The same stochastic
model that makes the exact likelihood intractable can be simulated forward exactly, one
trajectory at a time, so the simulation realizes the object at the top of the ladder and
serves as the ground truth against which every rung below it is judged.
