\section*{Materials and methods}

% Owner: methods_plan.md (M1--M11) + docs/figure_provenance.md (run manifest).
% Numbers are annotated with `% src:` per LINT-SRC. Definitions owned by the Theory
% and Diagnostics sections are cited here, not restated (00_master_list.md).

\subsection*{Minimal model and simulation}

We work with a two-state channel scheme, closed ($C$) and open ($O$), the smallest
scheme in which the acquisition-averaging effect we study is present.
% src: papers/macroir-elife-2025/methods_plan.md#M1 (scheme_CO; legacy/models_simple.h:10-66)
The closed-to-open transition is agonist dependent with association constant
$k_{\mathrm{on}}$ multiplying the agonist concentration; the open-to-closed
transition is agonist independent with dissociation rate $k_{\mathrm{off}}$. Only
the open state conducts, and the code applies a sign flip so the single-channel
current is inward. At the start of every recording all channels are closed, so the
occupancy is $(1,0)$.
% src: papers/macroir-elife-2025/methods_plan.md#M1 (v_g_formula; P_initial=[1,0]; sign flip)

The scheme carries six parameters, all estimated in base-ten logarithmic
coordinates. The values used to generate the data reported here, read from the
production run scripts rather than from the parameter tables in the repository
(the tables hold stale values the figures never used), are
$k_{\mathrm{on}} = 10\;\mu\mathrm{M}^{-1}\,\mathrm{s}^{-1}$,
$k_{\mathrm{off}} = 100\;\mathrm{s}^{-1}$,
a unitary conductance $g$ set to $1$, and a current baseline $b$ set to $1$
(both in units of $g$),
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:42-50 (on=10, off=100, unitary=1, baseline=1)
% src: papers/macroir-elife-2025/decisions/D-2_parameter_units.md (settled units; CSV != run values)
while the white-noise level and the mean channel number were swept (below). Current
is expressed throughout in units of the unitary conductance $g$, which corresponds
physically to a single-channel current of roughly $1\;\mathrm{pA}$.
% src: papers/macroir-elife-2025/decisions/D-2_parameter_units.md (current in units of g; g ~ 1 pA)
The dissociation time $\tau = 1/k_{\mathrm{off}} = 10\;\mathrm{ms}$ sets the natural
time scale that normalises the acquisition-interval axis. The relaxation at the
agonist condition is faster, with rate $k_{\mathrm{on}}[A] + k_{\mathrm{off}} =
200\;\mathrm{s}^{-1}$; the axis uses $\tau = 1/k_{\mathrm{off}}$ by convention.
% src: papers/macroir-elife-2025/decisions/D-2_parameter_units.md (tau = 1/k_off = 10 ms; lambda = k_on[A]+k_off)

The emission model is white instrumental noise only. Both the pink (one-over-$f$)
and the proportional noise channels are set to zero, so the term the reader may
expect from flicker noise is genuinely absent.
% src: papers/macroir-elife-2025/methods_plan.md#M1 (pink and proportional = 0)
For a measurement interval $t$ of duration $\Delta_t$, the predicted current has
mean $\bar y_t = N_{\mathrm{ch}}\,\mu_t\!\cdot\!\gamma_t + b$ and variance
\[
  \sigma_t^2 \;=\; e_t \;+\; N_{\mathrm{ch}}\,v_t ,
  \qquad
  e_t \;=\; \mathrm{Current\_Noise}\,\frac{f_s}{n_{\mathrm{samp},t}}
        \;=\; \frac{\mathrm{Current\_Noise}}{\Delta_t},
\]
where $N_{\mathrm{ch}}$ is the channel number, $\mu_t$ the occupancy mean, $\gamma_t$
the interval-averaged conductance, $f_s$ the raw sampling rate, $n_{\mathrm{samp},t}$
the number of raw samples averaged into the interval, $e_t$ the white instrumental
variance over the interval, and $N_{\mathrm{ch}}\,v_t$ the channel-gating variance.
% src: papers/macroir-elife-2025/decisions/D-2_parameter_units.md (e = Current_Noise*fs/n_samp; legacy/qmodel.h:3730,3735)
The white variance $e_t$ falls as the interval lengthens, because more raw samples
are averaged into each recorded value. The gating variance $v_t$ is the
single-interval conductance variance, whose form differs across the algorithm
family and is derived in the Theory section.

The stimulus is a single concentration jump in three segments, a pre-pulse at
$0\;\mu\mathrm{M}$, an agonist step at $10\;\mu\mathrm{M}$, and a post-pulse at
$0\;\mu\mathrm{M}$, with raw sampling at $f_s = 50\;\mathrm{kHz}$ throughout.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:54-61 (three segments; 0/10/0 uM; fs=50e3)
Each recorded measurement is the mean of $n_{\mathrm{samp},t}$ consecutive raw
samples, so the averaging window is fixed at acquisition. That window is the
acquisition interval the whole study is about, and its length is set by the
number of raw samples entering each recorded value.
% src: papers/macroir-elife-2025/methods_plan.md#M2 (interval = averaging window, not a filter)

Ground truth is generated by exact simulation of the continuous-time Markov chain
(CTMC). We use uniformization with $1{,}000$ substeps per measurement interval, so
the interval average of the conductance is realised from an actual sampled
trajectory rather than from any moment-closure approximation.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:69-74 (uniformization; number_of_substeps=1000)
This is what licenses treating the simulation as the reference against which the
Gaussian likelihood approximations are judged: the approximations are the object
under test, and the simulator supplies the exact answer they are compared to. The
two maximum-likelihood batteries used $10{,}000$ recordings per grid cell; the
time-resolved run used $1{,}000$ recordings; the single-recording illustration
(Figure~1) simulated one $20$-channel trace with $250$ within-interval substeps for
visualisation and does not enter the quantitative comparison.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#5 (n_sim=10000, 433ed13); #6 (1000 recordings); #4 (Figure 1: 20 ch, 250 substeps)

The simulation seed was set to $0$, which in this engine is the sentinel for a
random seed drawn from \texttt{std::random\_device}, and the resolved seed was
never logged. The deposited ensembles are therefore statistically equivalent to,
but not bit-for-bit reproducible from, a rerun of the same script. Because each
ensemble is large ($1{,}000$ to $10{,}000$ recordings), the reported statistics are
stable; the Monte-Carlo standard here is statistical equivalence, not bit identity.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#9 (seed=0 -> random_device; legacy/mcmc.h:37-45)
% src: papers/macroir-elife-2025/decisions/D-0_freeze_and_rerun_scope.md (seed never logged; not bit-reproducible; corrects methods_plan.md M5)

\subsection*{The macroscopic likelihood approximations}

The five likelihoods compared here are one filter run under two flags. The first flag, \emph{recursive approximation}, sets whether the
occupancy covariance is propagated between intervals. The second,
\emph{averaging approximation} $\mathrm{av}\in\{0,1,2\}$, sets how the
single-channel conductance is treated within an interval, and it literally counts
the number of interval endpoints the hidden state is conditioned on.
% src: papers/macroir-elife-2025/nomenclature.md (two flags; av counts conditioned endpoints)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:149-160 (algorithm -> (recursive, averaging) flags)

\begin{table}[htbp]
\centering
\begin{tabular}{lccl}
\toprule
Member & Recursive & $\mathrm{av}$ & Conditioning of the conductance \\
\midrule
NR  & no  & 0 & instantaneous; the acquisition averaging is ignored \\
NMR & no  & 1 & interval mean given the start state \\
R   & yes & 0 & instantaneous; the acquisition averaging is ignored \\
MR  & yes & 1 & interval mean given the start state \\
IR  & yes & 2 & interval mean given both boundary states, interior marginalised \\
\bottomrule
\end{tabular}
\caption{The likelihood family as two flags. The members form an endpoint ladder:
no endpoints (NR, R), one endpoint (NMR, MR), both boundary states (IR). The data
key for NMR is \texttt{macro\_NMR}.}
\label{tab:family}
\end{table}

Table~\ref{tab:family} lists the members. Reading the endpoint flag as a ladder,
the members condition the conductance on progressively more of the interval: no
endpoints, one endpoint, then both boundary states, with the exact trajectory (what
the simulator supplies) as the unreachable top rung.
% src: papers/macroir-elife-2025/nomenclature.md (endpoint ladder)
MacroIR is the top implemented rung: it conditions on the pair of channel states at
the two ends of the interval, which we call the boundary state, and marginalises the
interior analytically. It is the likelihood introduced for the P2X2 receptor
\citep{moffatt2025bayesian}, here placed as one member of the completed grid, and it
is verified in the Theory section to coincide with a time-integrated Kalman filter.
The construction of the interval-averaged mean and variance, the boundary-state
covariance, and the Kalman-like update are derived in the Theory section and its
supplement; we do not restate them here.

Two reductions are worth stating because they anchor the family to prior work.
Setting the boundary-conditioned conductance variance to zero in the recursive,
no-endpoint member (R) recovers the macroscopic-fluctuation update of
\citet{moffatt2007estimation}, so R is the recursive ancestor already in print.
% src: papers/macroir-elife-2025/methods_plan.md#M6 (R reduces to Moffatt 2007; macromrt_macromrt_paper_section.md)
The single structural difference between the M and I families is that the
one-endpoint members drop the boundary cross-covariance term
$N_{\mathrm{ch}}\,\gamma^{\mathsf T}\!\Sigma\gamma$ that IR retains.
% src: papers/macroir-elife-2025/nomenclature.md (MR drops N*gamma^T Sigma gamma that IR keeps)

Throughout the reported runs the remaining build flags were held fixed:
\emph{taylor variance correction} off, \emph{micro approximation} off, and the
\emph{variance approximation} on.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:81-89 (taylor=false, micro=false, variance=true)
Because the Taylor variance correction is off, the recursive one-endpoint member
that runs is MacroMR and not MacroMRT; the Taylor-corrected variants are not part
of this paper.
% src: papers/macroir-elife-2025/methods_plan.md#M6 (taylor=false -> MacroMR, not MRT)

The filter carries two numerical safeguards, present because the code runs them and
because anyone reimplementing the filter meets the same constraint on day one. The
occupancy mean is held on the probability simplex, and the occupancy covariance is
kept positive semidefinite through a trust-region shrink of the Kalman-like mean
update. Two constants enter, a binomial-count floor set to $5$ and a minimum
occupancy probability set to $10^{-12}$.
% src: papers/macroir-elife-2025/methods_plan.md#M11 (simplex + PSD trust-region; Binomial_magical_number=5, min_P=1e-12)
% src: theory/macroir/docs/Macro_IR/macro_ir_variance_inflation_correction.tex

\subsection*{Parameter estimation and the two anchors}

Inference in each grid cell is two-stage. First, the ensemble is partitioned into
groups and a separate maximum-likelihood estimate (MLE) is computed per group,
producing a cloud of estimates whose scatter is the empirical parameter covariance.
The optimiser is Gauss--Newton with Levenberg damping (initial damping $10$, growth
factor $3$, maximum damping $10^{10}$, at most $100$ iterations, relative gradient
tolerance $10^{-6}$, value tolerance $10^{-10}$), warm-started at the simulation
truth.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:93-95 (lambda_kickoff=10, lambda_factor=3, lambda_max=1e10, grad_rtol=1e-6, dvalue_tol=1e-10; max_iter=gn_max_iter is injected, not literal here) ; :104 (theta_warmstart=par_tr)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:107 (GN_MAX_ITER default 100 -> injected as gn_max_iter; the "at most 100 iterations")
Two group sizes were used, $10$ and $100$; with $10{,}000$ recordings these give
$1{,}000$ groups of $10$ and $100$ groups of $100$.
% src: papers/macroir-elife-2025/methods_plan.md#M7 (group_size 10 and 100; 1000 groups of 10, 100 of 100)
Second, a single joint MLE is computed over all replicates at once (one group of
size $n_{\mathrm{sim}}$), giving the pooled estimate $\theta_{\mathrm{pool}}$; its
near-zero gradient certifies that it is a stationary point of the pooled likelihood.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:131-148 (joint MLE = theta_pool)

Warm-starting the per-group fits at the truth is deliberate. It isolates the error
of the likelihood from the error of the optimiser, which is what this paper studies;
it would be the wrong choice for a study of an estimator, which this is not.
% src: papers/macroir-elife-2025/methods_plan.md#M7 (warm-start caveat: study of the likelihood, not the estimator)

The two anchors, the simulation truth $\theta_{\mathrm{sim}}$ and the pooled fit
$\theta_{\mathrm{pool}}$, are distinct points because the Gaussian macroscopic
likelihood is misspecified with respect to the exact simulator. Their difference
$\theta_{\mathrm{pool}} - \theta_{\mathrm{sim}}$ is the misspecification bias. Every
diagnostic below is computed at both anchors, and the two answer different
questions: at $\theta_{\mathrm{pool}}$ the score is zero by construction so bias
washes out, whereas at $\theta_{\mathrm{sim}}$ the bias is visible but the Fisher
information can be indefinite away from the maximum. Each figure states which anchor
it uses.
% src: papers/macroir-elife-2025/methods_plan.md#M7 (two anchors; state the anchor per figure)
% src: papers/macroir-elife-2025/docs/figure_provenance.md#5 (theta_pool score ~ 0; theta_sim Fisher can be indefinite)

\subsection*{Diagnostics and the Gaussian-Fisher anchor}

Faithfulness of each approximate likelihood to the simulator is measured with
likelihood-level diagnostics: standardized residuals, the score (the gradient of
the log-likelihood), and an information-distortion matrix that compares the
covariance $J$ of the score across replicates with the model's own Fisher
information. The definitions, sign conventions, and thresholds of these
diagnostics are owned by the Diagnostics section and its metric registry; here we
describe only how they were computed.
% src: papers/macroir-elife-2025/03_metrics_diagnostics.md (diagnostic definitions owned there)

The Fisher information was formed two ways. The finite-difference construction, which
we write $F_{\mathrm{num}}$, is a central difference of the analytic score, with
per-coordinate step
$h_i = h_{\mathrm{rel}}\max(|\theta_i|,1)$ at $h_{\mathrm{rel}} = 10^{-5}$, and it is
symmetrised as $(F_{\mathrm{num}} + F_{\mathrm{num}}^{\mathsf T})/2$. Because the score
itself is obtained by automatic differentiation, this is a single numerical derivative,
not two.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:151-157 (numerical Fisher; h_rel injected)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:98 (H_RELS default 1e-5)
% src: papers/macroir-elife-2025/docs/figure_provenance.md#5 (central difference of analytic score; symmetrized)
The analytic construction is the Gaussian Fisher information
$\bar G = \sum_t \mathrm{GFI}_t$, summed over intervals. This analytic Gaussian Fisher
is the anchor the Diagnostics section writes $\mathbf{H}$, that is
$\mathbf{H} = \bar G = \sum_t \mathrm{GFI}_t$. Its per-interval term is
\[
  \mathrm{GFI}_t = \frac{(\partial_\theta \bar y_t)(\partial_\theta \bar y_t)^{\mathsf T}}{\sigma_t^2}
  + \frac{(\partial_\theta \sigma_t^2)(\partial_\theta \sigma_t^2)^{\mathsf T}}{2\,\sigma_t^4},
\]
which follows from the analytic derivative alone and needs no extra likelihood
passes.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#7 (Gaussian Fisher G_b = sum_t GFI_t; legacy/qmodel.h:6096-6098)
Comparing the score covariance $J$ against either Fisher construction is the
sandwich, or information-matrix-equality, diagnostic \citep{white1982maximum}:
equality between the Fisher and the score covariance ($\mathbf{H} = J$ under the
analytic anchor, $F_{\mathrm{num}} = J$ under the finite-difference one) holds only
when the likelihood is correctly specified, and the extent to which they differ is
what the distortion matrix quantifies.

Uncertainty on every reported statistic comes from a nonparametric bootstrap over
groups, reported as percentile intervals at the $2.5$, $16$, $50$, $84$ and
$97.5\%$ quantiles. Despite the internal column names, no probit transform is
applied; the intervals are ordinary empirical percentiles, verified equal to the
R quantile of type~$6$.
% src: papers/macroir-elife-2025/methods_plan.md#M8 (percentile bootstrap over groups; NOT probit; quantiles listed)
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:111 (probit_cis = [0.025,0.16,0.5,0.84,0.975])
The diagnostic battery used $100$ bootstrap replicates and a maximum lag of $10$ for
the score autocorrelation; the empirical-versus-theoretical distortion capstone used
$200$ replicates; and fewer than $10$ groups leaves the interval slots empty
(not-a-number filled), which is one source of empty cells in the maps.
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:174-175,198 (n_bootstrap 100 + max_lag 10; capstone 200)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:106 (MIN_GROUPS=10)
The raw per-group MLE cloud is written unbootstrapped; any resampling of it is done
downstream in the figure code.
% src: papers/macroir-elife-2025/methods_plan.md#M8 (Model_Parameters_Hat not bootstrapped by the program)

Two production batteries were run that differ only in which Fisher construction
anchors the diagnostics. The finite-difference battery covers all five algorithms;
the Gaussian battery covers MacroIR across the full channel-by-noise grid, with the
remaining members being added on the canonical grid by the in-progress fill (below).
% REFRESH-ON-GAUSSIAN
Cross-algorithm comparisons are drawn from the finite-difference battery and
MacroIR-specific mechanism results from the Gaussian battery. The split is a design
choice rather than a completed cross-grid verification: the analytic Gaussian Fisher
and the finite-difference Fisher target the same information, an identity the
time-resolved figures show holds per interval, so anchoring a given panel on either
construction is consistent. The finite-difference battery is retained to exhibit that
agreement, not to anchor the mechanism results. The choice of anchor is recorded in
the project decision log.
% src: papers/macroir-elife-2025/methods_plan.md#M10 (anchor split; per-step F_t = J_t justification)
% src: papers/macroir-elife-2025/decisions/D-0_freeze_and_rerun_scope.md (multi-commit provenance; Gaussian fill scope)
% src: papers/macroir-elife-2025/docs/figure_provenance.md#7 (numerical run = all five; Gaussian run = IR + fill)

\subsection*{Regime grid and reduction}

The acquisition interval was swept over seven levels of $\Delta/\tau$, namely
$1$, $0.5$, $0.2$, $0.1$, $0.05$, $0.02$ and $0.01$, realised as
$500$, $250$, $100$, $50$, $25$, $10$ and $5$ raw samples averaged per interval,
that is step durations of $10$, $5$, $2$, $1$, $0.5$, $0.2$ and $0.1\;\mathrm{ms}$,
giving $10$, $20$, $50$, $100$, $200$, $500$ and $1{,}000$ steps in the record.
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:184-190 (interval_in_tau labels; n_samp; step counts)
The total record duration is held fixed at $100\;\mathrm{ms}$ across the sweep, so
the interval is varied without varying the amount of data. Within each record the
pre-pulse, agonist and post-pulse segments hold a fixed ratio of $20{:}40{:}40$
percent of the steps.
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:185-190 (segment step counts 2:4:4 ... 20:40:40; total 100 ms)

The swept ``noise'' axis is dimensionless. It is
$\mathrm{noise\_in\_conductance\_tau} = 10\,\mathrm{Current\_Noise}\,k_{\mathrm{off}}/g^2$,
the white current variance accumulated over one relaxation time $\tau$ taken relative
to the single-channel signal power $g^2$, with a factor placing the level at which
the white noise equals the single-channel signal near a label of $10$. With the run
values it equals $1000\,\mathrm{Current\_Noise}$, so the figure label $0.1$
corresponds to the engine parameter $\mathrm{Current\_Noise} = 10^{-4}$ (a white-noise
power spectral density in units of $g^2\,\mathrm{s}$).
% src: papers/macroir-elife-2025/decisions/D-2_parameter_units.md (noise_in_conductance_tau = 10*Current_Noise*k_off/g^2; = 1000*Current_Noise; label 0.1 -> 1e-4; crossover at 10)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:162-167,175-176 (Current_Noise = label/1000; axis name)
The figures carry the dimensionless label on their axes; the underlying engine
parameter is not what is plotted.

The mean channel number $N_{\mathrm{ch}}$ and the noise level were the other two map
axes. The finite-difference production grid on disk holds all five algorithms at
$N_{\mathrm{ch}} \in \{10, 100, 1000, 10000\}$ and noise labels $\{0.1, 1, 10\}$ at
$10{,}000$ recordings, with MacroIR additionally at
$N_{\mathrm{ch}} \in \{20, 50, 200, 500\}$; the Gaussian grid holds MacroIR at
$N_{\mathrm{ch}} \in \{10, 20, 50, 100, 200, 500, 1000, 2000, 5000, 10000\}$ and the
six noise labels $\{0.05, 0.1, 0.2, 0.5, 1, 10\}$. Coverage of the four non-IR
members on the Gaussian anchor is supplied by the in-progress D-0 fill, on the
canonical $N_{\mathrm{ch}} \in \{10, 100, 1000, 10000\}$ grid at the three noise
labels $\{0.1, 1, 10\}$ (not MacroIR's six); until that fill lands these members are
not yet deposited across the grid, and any figure that needs them refreshes when it
does.
% REFRESH-ON-GAUSSIAN
% src: papers/macroir-elife-2025/docs/figure_provenance.md#5 (433ed13 grid) ; #7 (1c2ae6f: NMR entirely absent from Gaussian family, NR missing 10000-channel cell; non-IR fill pending)
% src: papers/macroir-elife-2025/decisions/D-0_freeze_and_rerun_scope.md (Gaussian fill IN PROGRESS on Dirac: {NR,NMR,R,MR} x Nch{10,100,1000,10000} x noise{0.1,1,10}; 3-label noise grid, not IR's 6)

Each grid cell is reduced through six stages: the per-group MLE cloud; the pooled
joint fit $\theta_{\mathrm{pool}}$; the Fisher information at both anchors; the score
at both anchors, whose across-replicate covariance is $J$; the diagnostic battery
paired with the Fisher at each anchor; and the empirical-versus-theoretical
distortion capstone, anchored at $\theta_{\mathrm{pool}}$. Each cell writes five
comma-separated-value families (the estimate cloud, the pooled fit, the two
anchor batteries, and the capstone).
% src: projects/eLife_2025/ops/local/figure_3_mle.macroir:97-201 (six stages; five CSV families per cell)

\subsection*{Reproducibility and computational environment}

The engine is \texttt{macro\_dr}, built as C++20, and each analysis is a script in a
small typed domain-specific language run in a single process
\citep{moffatt2025macroir}. Provenance is self-documenting at three levels: on every
invocation the binary snapshots the assembled script and its full argument list into
a run ledger; every analysis output file stamps the build's short git commit hash as
its first row; and the data directories are named by that hash (\texttt{433ed13} for
the finite-difference battery, \texttt{1c2ae6f} for the Gaussian battery), so results
from different code versions can never overwrite one another and multi-commit
provenance is documented per file.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#2 (run ledger; hash-stamped row 1; hash-named data dirs)
% src: papers/macroir-elife-2025/decisions/D-0_freeze_and_rerun_scope.md (each CSV self-stamps its hash; multi-commit provenance accepted)

The production batteries ran under SLURM at $32$ CPUs per task and $48\;\mathrm{GB}$
of memory ($96\;\mathrm{GB}$ for the parameter-recovery lane), with the linear-algebra
back end pinned to a single thread so that OpenMP parallelises the per-simulation and
bootstrap loops. The environment variable \texttt{MACRODR\_AXIS\_SERIAL=1} is
load-bearing: it serialises the internal grid-combination loop, without which memory
grows with the number of concurrent combinations and the jobs exhaust memory.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#2,#9 (SLURM 32 CPU, 48/96 GB, BLAS 1 thread, MACRODR_AXIS_SERIAL=1)
% src: projects/eLife_2025/ops/slurm/dispatch_figure_3.sh:196-205 (CPUS 32, MEM 48G, MACRODR_AXIS_SERIAL=1)

The one reproducibility limit worth stating twice is the simulation seed. Because the
seed was the random sentinel and the resolved value was not recorded, the exact
deposited ensembles cannot be regenerated bit for bit; a rerun yields a statistically
equivalent, numerically different ensemble. All resampling that operates on a fixed
dataset (the bootstraps, the group partitioning) is deterministic and reproducible
given the data. Deposition of the data and code, with pinned versions, is stated in
the Data and code availability section.
% src: papers/macroir-elife-2025/docs/figure_provenance.md#9 (seed not reproducible; bootstrap on fixed data is deterministic)
% src: papers/macroir-elife-2025/decisions/D-0_freeze_and_rerun_scope.md (statistical-equivalence standard; deposition in backmatter)
